---
title: "glm"
author: "Linnea Honeker"
date: "9/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(phyloseq)

library(vegan)

library(ggplot2)

library(rms)

library(reshape2)

library(coefplot)

library(MASS)

library(dplyr)

library(lme4)

library(MASS)

library(leaps)

library(car)

library(relaimpo)

library(nlme)

library(lmerTest)



```

## perform pairwise correlations and scatterplots for gene transcript copies and flux data
# load data tables 1) KO of interest and 2) gene transcript copies. Clean up data times for purposes of this analysis.

```{r, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

KO_interest_subset <- read.csv("../../Data/Deseq2/KO_interest.csv")
metaT <- read.csv("../../Output/WGCNA/vst-metaT.csv")
metadata <- read.csv("../../Data/Deseq2/metaT-metadata.csv")
acetate_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/acetic-acid-for-ggplot.csv")
acetone_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/acetone-for-ggplot.csv")
diacetyl_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/diacetyl-for-ggplot.csv")

#remove "KO:"
metaT$X <- sub("KO:", "", metaT$X)
diacetyl_flux$Sample <- sub("_C1", "", diacetyl_flux$Sample)
diacetyl_flux$Sample <- sub("_C2", "", diacetyl_flux$Sample)

#merge  all VOC
KO_VOC <- KO_interest_subset %>%
  merge(., metaT, by.x = "KO", by.y="X", all.x = FALSE, all.y = FALSE)

#transpose for glm
KO_VOC.t <- t(KO_VOC)

#further format KO_acetate.t for glm. Assign column headers and remove rows containing metadata
colnames(KO_VOC.t) <- KO_VOC.t[1,]
KO_VOC.t.f <- KO_VOC.t[-(1:6),]
KO_VOC.t.f <- as.data.frame(KO_VOC.t.f)

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acectate_flux_proc <- acetate_flux %>%
  filter (Type == "sample") %>%
    filter (Pyruv == "C2") %>%
     mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetate = Flux) %>%
  dplyr::select(-c("Pyruv", "Type", "Flux", "Time", "Time.hr"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetone_flux_proc <- acetone_flux %>%
  filter (Type == "sample") %>%
  filter (Pyruv == "C2") %>%
  mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetone = Flux)%>%
   dplyr::select(-c("Pyruv", "Type", "Flux", "Time", "Time.hr", "Site"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
diacetyl_flux_proc <- diacetyl_flux %>%
  filter (Type == "sample") %>%
  filter (Pyruv == "C2") %>%
  mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.diacetyl = Flux) %>%
   dplyr::select(-c("Pyruv", "Type", "Flux", "Time", "Time.hr", "Site"))

#sample names are different in the flux and expression tables, so need to make them the same
KO_VOC.t.f[,26] <- rownames(KO_VOC.t.f)
KO_VOC.t.f_new_id <- KO_VOC.t.f %>%
  mutate(Sample = case_when(
  startsWith(V26, "X3300044942") ~ "S3P4",
  startsWith(V26, "X3300044943") ~ "S3P4",
  startsWith(V26, "X3300044944") ~ "S3P4",
  startsWith(V26, "X3300044945") ~ "S3P4",
  startsWith(V26, "X3300045458") ~ "S1P5",
  startsWith(V26, "X3300045459") ~ "S1P5",
  startsWith(V26, "X3300045460") ~ "S1P5",
  startsWith(V26, "X3300045461") ~ "S2P4",
  startsWith(V26, "X3300045462") ~ "S2P4",
  startsWith(V26, "X3300045463") ~ "S2P4",
  startsWith(V26, "X3300045464") ~ "S2P4",
  startsWith(V26, "X3300045465") ~ "S2P4",
  startsWith(V26, "X3300045466") ~ "S2P6",
  startsWith(V26, "X3300045467") ~ "S2P6",
  startsWith(V26, "X3300045468") ~ "S2P6",
  startsWith(V26, "X3300045469") ~ "S2P6",
  startsWith(V26, "X3300045470") ~ "S2P6",
  startsWith(V26, "X3300045471") ~ "S3P5",
  startsWith(V26, "X3300045472") ~ "S3P5",
  startsWith(V26, "X3300045473") ~ "S2P6",
  startsWith(V26, "X3300048051") ~ "S1P5",
  startsWith(V26, "X3300048052") ~ "S3P5",
  startsWith(V26, "X3300048053") ~ "S3P5",
  startsWith(V26, "X3300048054") ~ "S3P5",
  startsWith(V26, "X3300048092") ~ "S1P1",
  startsWith(V26, "X3300048093") ~ "S1P1",
  startsWith(V26, "X3300048094") ~ "S1P1",
  startsWith(V26, "X3300048095") ~ "S1P1",
  startsWith(V26, "X3300048096") ~ "S1P1",
  startsWith(V26, "X3300048099") ~ "S1P5",
  startsWith(V26, "X3300048100") ~ "S1P5",
  startsWith(V26, "X3300049511") ~ "S2P4",
  startsWith(V26, "X3300049512") ~ "S3P4",
  startsWith(V26, "X3300049557") ~ "S3P4",
  startsWith(V26, "X3300049628") ~ "S3P5"
  )) %>%
  mutate(Time.d = case_when(
    endsWith(V26, "6hr") ~ "6",
    endsWith(V26, "0hr") ~ "0",
    endsWith(V26, "48hr") ~ "48",
    endsWith(V26, "48_hr") ~ "48"
  )) %>%
  mutate(Condition = case_when(
    grepl("PreDrought", V26) ~ "pre-drought",
    grepl("Drought", V26) ~ "drought",
    grepl("Pre_drought", V26) ~ "pre-drought"
  )) %>%
   dplyr::select(-V26)

#merge VOC and KO tables
merged_subset <- KO_VOC.t.f_new_id %>%
  merge(., acectate_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample","Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate_at(c(4:30), as.numeric) %>%
  mutate(flux.acetate = as.numeric(flux.acetate)) %>%
  mutate(log.flux.acetate = log(flux.acetate+abs(min(flux.acetate))+1, base = exp(2))) %>%
  merge(.,acetone_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.acetone = as.numeric(flux.acetone)) %>%
  mutate(log.flux.acetone = log(flux.acetone+abs(min(flux.acetone))+1, base = exp(2))) %>%
  merge(.,diacetyl_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.diacetyl = as.numeric(flux.diacetyl)) %>%
  mutate(log.flux.diacetyl = log(flux.diacetyl+abs(min(flux.diacetyl))+1, base = exp(2))) 
  
merged_subset
```

### correlations between each gene and acetone
```{r correlation-plots-acetone, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#make subset of numerical data
merged_subset.num <- merged_subset %>%
   dplyr::select(-c(Sample, Condition, Time.d, Site))

sub_col <- c(5:35)
z = merged_subset$Condition
y = merged_subset$flux.acetone
colNames <- names(merged_subset)[sub_col]
for (i in colNames) {
  p <- ggplot(merged_subset, aes_string(x=i, y=y)) +
  geom_point(aes(color=Condition, shape = Time.d)) +
  geom_smooth(method=lm, se=TRUE)
  ggtitle (aes_string(i))
  print(p)
}


```


### correlations between each gene and acetone
```{r correlation-plots-acetone, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#make subset of numerical data
merged_subset.num <- merged_subset %>%
   dplyr::select(-c(Sample, Condition, Time.d, Site))

sub_col <- c(5:35)
z = merged_subset$Condition
y = merged_subset$flux.acetate
colNames <- names(merged_subset)[sub_col]
for (i in colNames) {
  p <- ggplot(merged_subset, aes_string(x=i, y=y)) +
  geom_point(aes(color=Condition, shape = Time.d)) +
  geom_smooth(method=lm, se=TRUE)
  ggtitle (aes_string(i))
  print(p)
}


```

### correlations between each gene and diacetyl
```{r correlation-plots-acetone, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#make subset of numerical data
merged_subset.num <- merged_subset %>%
   dplyr::select(-c(Sample, Condition, Time.d, Site))

sub_col <- c(5:35)
z = merged_subset$Condition
y = merged_subset$flux.diacetyl
colNames <- names(merged_subset)[sub_col]
for (i in colNames) {
  p <- ggplot(merged_subset, aes_string(x=i, y=y)) +
  geom_point(aes(color=Condition, shape = Time.d)) +
  geom_smooth(method=lm, se=TRUE)
  ggtitle (aes_string(i))
  print(p)
}


```


```{r test for normality in VOC flux data}

hist(merged_subset$log.flux.acetate, col='steelblue', main = 'log acetate flux')
hist(merged_subset$flux.acetate, col='steelblue', main = 'acetate flux')
hist(merged_subset$K00158, col='steelblue', main = 'K00158')
hist(merged_subset$K00004, col='steelblue', main = 'K00004')
hist(merged_subset$K01067)
hist(merged_subset$K01574)




```
### correlations between each acetone and physicochemical properties
```{r correlation-plots-acetone, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#make subset of numerical data
merged_subset.num <- merged_subset_physico_full_names %>%
   dplyr::select(-c(Sample, Condition, Time.d, Site))

sub_col <- c(39:41)
z = merged_subset_physico_full_names$Condition
y = merged_subset_physico_full_names$flux.acetone
colNames <- names(merged_subset.num)[sub_col]
for (i in colNames) {
  p <- ggplot(merged_subset_physico_full_names, aes_string(x=i, y=y)) +
  geom_point(aes(color=Condition, shape = Site)) +
  geom_smooth(method=lm, se=TRUE)
  ggtitle (aes_string(i))
  print(p)
}




```
### correlations between each acetate and physicochemical properties
```{r correlation-plots-acetone, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#make subset of numerical data
merged_subset.num <- merged_subset_physico_full_names %>%
   dplyr::select(-c(Sample, Condition, Time.d, Site))

sub_col <- c(39:41)
z = merged_subset_physico_full_names$Condition
y = merged_subset_physico_full_names$flux.acetate
colNames <- names(merged_subset.num)[sub_col]
for (i in colNames) {
  p <- ggplot(merged_subset_physico_full_names, aes_string(x=i, y=y)) +
  geom_point(aes(color=Condition, shape = Site)) +
  geom_smooth(method=lm, se=TRUE)
  ggtitle (aes_string(i))
  print(p)
}




```

### correlations between each acetate and physicochemical properties
```{r correlation-plots-acetone, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

#make subset of numerical data
merged_subset.num <- merged_subset_physico_full_names %>%
   dplyr::select(-c(Sample, Condition, Time.d, Site))

sub_col <- c(39:41)
z = merged_subset_physico_full_names$Condition
y = merged_subset_physico_full_names$flux.diacetyl
colNames <- names(merged_subset.num)[sub_col]
for (i in colNames) {
  p <- ggplot(merged_subset_physico_full_names, aes_string(x=i, y=y)) +
  geom_point(aes(color=Condition, shape = Site)) +
  geom_smooth(method=lm, se=TRUE)
  ggtitle (aes_string(i))
  print(p)
}




```


```{r linear regression statistics}
#cor(merged_subset$K01574, merged_subset$log.flux.acetone)

summary(lm(merged_subset_physico_full_names$SM ~ merged_subset_physico_full_names$flux.acetone))
summary(lm(merged_subset_physico_full_names$SM ~ merged_subset_physico_full_names$flux.acetate))
summary(lm(merged_subset_physico_full_names$SM ~ merged_subset_physico_full_names$flux.diacetyl))
summary(lm(merged_subset_physico_full_names$ST~ merged_subset_physico_full_names$flux.acetone))
summary(lm(merged_subset_physico_full_names$ST ~ merged_subset_physico_full_names$flux.acetate))
summary(lm(merged_subset_physico_full_names$ST ~ merged_subset_physico_full_names$flux.diacetyl))
summary(lm(merged_subset_physico_full_names$pH ~ merged_subset_physico_full_names$flux.acetone))
summary(lm(merged_subset_physico_full_names$pH ~ merged_subset_physico_full_names$flux.acetate))
summary(lm(merged_subset_physico_full_names$pH ~ merged_subset_physico_full_names$flux.diacetyl))
```

