---
title: "PLS"
author: "Linnea Honeker"
date: "11/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(phyloseq)
library(vegan)
library(ggplot2)
library(rms)
library(reshape2)
library(coefplot)
library(MASS)
library(dplyr)
library(lme4)
library(MASS)
library(leaps)
library(car)
library(relaimpo)
library(nlme)
library(lmerTest)
library(plsdepot)
library(pls)
library(caret)

```

# load data tables 1) KO of interest and 2) gene transcript copies. Clean up data times for purposes of this analysis.
## All possible KOs
```{r all-KOs, echo = FALSE}
KO_interest <- read.csv("../../Data/Deseq2/VOC-KO-interest-from-ML.csv")
metaT <- read.csv("../../Output/WGCNA/vst-metaT.csv")
metadata <- read.csv("../../Data/Deseq2/metaT-metadata.csv")
acetate_flux <- df.acetate
acetone_flux <- df.acetone
diacetyl_flux <- df.diacetyl

#remove "KO:"
metaT$X <- sub("KO:", "", metaT$X)

#merge  all VOC
KO_VOC <- KO_interest %>%
  merge(., metaT, by.x = "KO", by.y="X", all.x = FALSE, all.y = FALSE)

#transpose for glm
KO_VOC.t <- t(KO_VOC)

#further format KO_acetate.t for glm. Assign column headers and remove rows containing metadata
colnames(KO_VOC.t) <- KO_VOC.t[1,]
KO_VOC.t.f <- KO_VOC.t[-(1:3),]
KO_VOC.t.f <- as.data.frame(KO_VOC.t.f)

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetate_flux_proc <- acetate_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetate = Flux) %>%
  dplyr::select(-c("Pyr", "Flux", "Time", "Time.hr"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetone_flux_proc <- acetone_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetone = Flux) %>%
  dplyr::select(-c("Pyr", "Flux", "Time", "Time.hr"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
diacetyl_flux_proc <- diacetyl_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.diacetyl = Flux) %>%
  dplyr::select(-c("Pyr", "Flux", "Time", "Time.hr"))

#sample names are different in the flux and expression tables, so need to make them the same
KO_VOC.t.f[,37] <- rownames(KO_VOC.t.f)
KO_VOC.t.f_new_id <- KO_VOC.t.f %>%
  mutate(Sample = case_when(
  startsWith(V37, "X3300044942") ~ "S3P4",
  startsWith(V37, "X3300044943") ~ "S3P4",
  startsWith(V37, "X3300044944") ~ "S3P4",
  startsWith(V37, "X3300044945") ~ "S3P4",
  startsWith(V37, "X3300045458") ~ "S1P5",
  startsWith(V37, "X3300045459") ~ "S1P5",
  startsWith(V37, "X3300045460") ~ "S1P5",
  startsWith(V37, "X3300045461") ~ "S2P4",
  startsWith(V37, "X3300045462") ~ "S2P4",
  startsWith(V37, "X3300045463") ~ "S2P4",
  startsWith(V37, "X3300045464") ~ "S2P4",
  startsWith(V37, "X3300045465") ~ "S2P4",
  startsWith(V37, "X3300045466") ~ "S2P6",
  startsWith(V37, "X3300045467") ~ "S2P6",
  startsWith(V37, "X3300045468") ~ "S2P6",
  startsWith(V37, "X3300045469") ~ "S2P6",
  startsWith(V37, "X3300045470") ~ "S2P6",
  startsWith(V37, "X3300045471") ~ "S3P5",
  startsWith(V37, "X3300045472") ~ "S3P5",
  startsWith(V37, "X3300045473") ~ "S2P6",
  startsWith(V37, "X3300048051") ~ "S1P5",
  startsWith(V37, "X3300048052") ~ "S3P5",
  startsWith(V37, "X3300048053") ~ "S3P5",
  startsWith(V37, "X3300048054") ~ "S3P5",
  startsWith(V37, "X3300048092") ~ "S1P1",
  startsWith(V37, "X3300048093") ~ "S1P1",
  startsWith(V37, "X3300048094") ~ "S1P1",
  startsWith(V37, "X3300048095") ~ "S1P1",
  startsWith(V37, "X3300048096") ~ "S1P1",
  startsWith(V37, "X3300048099") ~ "S1P5",
  startsWith(V37, "X3300048100") ~ "S1P5",
  startsWith(V37, "X3300049511") ~ "S2P4",
  startsWith(V37, "X3300049512") ~ "S3P4",
  startsWith(V37, "X3300049557") ~ "S3P4",
  startsWith(V37, "X3300049628") ~ "S3P5"
  )) %>%
  mutate(Time.d = case_when(
    endsWith(V37, "6hr") ~ "6",
    endsWith(V37, "0hr") ~ "0",
    endsWith(V37, "48hr") ~ "48",
    endsWith(V37, "48_hr") ~ "48"
  )) %>%
  mutate(Condition = case_when(
    grepl("PreDrought", V37) ~ "Pre-Drought",
    grepl("Drought", V37) ~ "Drought",
    grepl("Pre_drought", V37) ~ "Pre-Drought"
  )) %>%
   dplyr::select(-V37)

#merge VOC and KO tables
merged <- KO_VOC.t.f_new_id %>%
  merge(., acectate_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Label","Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate_at(c(4:39), as.numeric) %>%
  mutate(flux.acetate = as.numeric(flux.acetate)) %>%
  mutate(log.flux.acetate = log(flux.acetate+abs(min(flux.acetate))+1, base = exp(2))) %>%
  merge(.,acetone_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Label", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.acetone = as.numeric(flux.acetone)) %>%
  mutate(log.flux.acetone = log(flux.acetone+abs(min(flux.acetone))+1, base = exp(2))) %>%
  merge(.,diacetyl_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Label", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.diacetyl = as.numeric(flux.diacetyl)) %>%
  mutate(log.flux.diacetyl = log(flux.diacetyl+abs(min(flux.diacetyl))+1, base = exp(2))) 
  
merged
```

## Subset of KOs originally included in the first manuscript submission

```{r KOs-subset, echo=FALSE}
KO_interest_subset <- read.csv("../../Data/Deseq2/KO_interest.csv")
metaT <- read.csv("../../Output/WGCNA/vst-metaT.csv")
metadata <- read.csv("../../Data/Deseq2/metaT-metadata.csv")
acetate_flux <- df.acetate
acetone_flux <- df.acetone
diacetyl_flux <- df.diacetyl

#remove "KO:"
metaT$X <- sub("KO:", "", metaT$X)

#merge  all VOC
KO_VOC <- KO_interest_subset %>%
  merge(., metaT, by.x = "KO", by.y="X", all.x = FALSE, all.y = FALSE)

#transpose for glm
KO_VOC.t <- t(KO_VOC)

#further format KO_acetate.t for glm. Assign column headers and remove rows containing metadata
colnames(KO_VOC.t) <- KO_VOC.t[1,]
KO_VOC.t.f <- KO_VOC.t[-(1:6),]
KO_VOC.t.f <- as.data.frame(KO_VOC.t.f)

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetate_flux_proc <- acetate_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetate = Flux) %>%
  dplyr::select(-c("Pyr", "Flux", "Time", "Time.hr"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetone_flux_proc <- acetone_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetone = Flux) %>%
  dplyr::select(-c("Pyr", "Flux", "Time", "Time.hr"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
diacetyl_flux_proc <- diacetyl_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.diacetyl = Flux) %>%
  dplyr::select(-c("Pyr", "Flux", "Time", "Time.hr"))

#sample names are different in the flux and expression tables, so need to make them the same
KO_VOC.t.f[,26] <- rownames(KO_VOC.t.f)
KO_VOC.t.f_new_id <- KO_VOC.t.f %>%
  mutate(Sample = case_when(
  startsWith(V26, "X3300044942") ~ "S3P4",
  startsWith(V26, "X3300044943") ~ "S3P4",
  startsWith(V26, "X3300044944") ~ "S3P4",
  startsWith(V26, "X3300044945") ~ "S3P4",
  startsWith(V26, "X3300045458") ~ "S1P5",
  startsWith(V26, "X3300045459") ~ "S1P5",
  startsWith(V26, "X3300045460") ~ "S1P5",
  startsWith(V26, "X3300045461") ~ "S2P4",
  startsWith(V26, "X3300045462") ~ "S2P4",
  startsWith(V26, "X3300045463") ~ "S2P4",
  startsWith(V26, "X3300045464") ~ "S2P4",
  startsWith(V26, "X3300045465") ~ "S2P4",
  startsWith(V26, "X3300045466") ~ "S2P6",
  startsWith(V26, "X3300045467") ~ "S2P6",
  startsWith(V26, "X3300045468") ~ "S2P6",
  startsWith(V26, "X3300045469") ~ "S2P6",
  startsWith(V26, "X3300045470") ~ "S2P6",
  startsWith(V26, "X3300045471") ~ "S3P5",
  startsWith(V26, "X3300045472") ~ "S3P5",
  startsWith(V26, "X3300045473") ~ "S2P6",
  startsWith(V26, "X3300048051") ~ "S1P5",
  startsWith(V26, "X3300048052") ~ "S3P5",
  startsWith(V26, "X3300048053") ~ "S3P5",
  startsWith(V26, "X3300048054") ~ "S3P5",
  startsWith(V26, "X3300048092") ~ "S1P1",
  startsWith(V26, "X3300048093") ~ "S1P1",
  startsWith(V26, "X3300048094") ~ "S1P1",
  startsWith(V26, "X3300048095") ~ "S1P1",
  startsWith(V26, "X3300048096") ~ "S1P1",
  startsWith(V26, "X3300048099") ~ "S1P5",
  startsWith(V26, "X3300048100") ~ "S1P5",
  startsWith(V26, "X3300049511") ~ "S2P4",
  startsWith(V26, "X3300049512") ~ "S3P4",
  startsWith(V26, "X3300049557") ~ "S3P4",
  startsWith(V26, "X3300049628") ~ "S3P5"
  )) %>%
  mutate(Time.d = case_when(
    endsWith(V26, "6hr") ~ "6",
    endsWith(V26, "0hr") ~ "0",
    endsWith(V26, "48hr") ~ "48",
    endsWith(V26, "48_hr") ~ "48"
  )) %>%
   mutate(Condition = case_when(
    grepl("PreDrought", V26) ~ "Pre-Drought",
    grepl("Drought", V26) ~ "Drought",
    grepl("Pre_drought", V26) ~ "Pre-Drought"
  )) %>%
   dplyr::select(-V26)

#merge VOC and KO tables
merged <- KO_VOC.t.f_new_id %>%
  merge(., acetate_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Label","Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate_at(c(4:30), as.numeric) %>%
  mutate(flux.acetate = as.numeric(flux.acetate)) %>%
  mutate(log.flux.acetate = log(flux.acetate+abs(min(flux.acetate))+1, base = exp(2))) %>%
  merge(.,acetone_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Label", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.acetone = as.numeric(flux.acetone)) %>%
  mutate(log.flux.acetone = log(flux.acetone+abs(min(flux.acetone))+1, base = exp(2))) %>%
  merge(.,diacetyl_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Label", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.diacetyl = as.numeric(flux.diacetyl)) %>%
  mutate(log.flux.diacetyl = log(flux.diacetyl+abs(min(flux.diacetyl))+1, base = exp(2))) 
  
merged
```

## Filter to aceate flux and acetate cycling genes
```{r}
KO_acetate <- KO_interest_subset %>%
  filter(VOC == "Acetate") 

KO_acetate_list <- KO_acetate$KO
flux_list <- c("flux.acetate", "Condition", "Site")

idx <- match(c(KO_acetate_list,flux_list), names(merged_subset))
idx <- idx[!is.na(idx)]

sub_acetate <- merged_subset[,idx]

sub_acetate2 <- sub_acetate %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)

sub_acetate2
```

## Filter to acetate and acetone flux and acetate cycling genes
```{r}
KO_acetate <- KO_interest_subset %>%
  filter(VOC == "Acetate") 

KO_acetate_list <- KO_acetate$KO
flux_list <- c("flux.acetate", "flux.acetone", "Condition", "Site")

idx <- match(c(KO_acetate_list,flux_list), names(merged_subset))
idx <- idx[!is.na(idx)]

sub_acetate_acetone_flux <- merged_subset[,idx]

sub_acetate_acetone_flux2 <- sub_acetate_acetone_flux %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)

sub_acetate_acetone_flux2
```
## Filter to acetate and acetone flux and acetate and acetone cycling genes
```{r}
KO_acetate_acetone <- KO_interest_subset %>%
  filter(VOC == "Acetate" |
           VOC == "Acetone") 

KO_acetate_acetone_list <- KO_acetate_acetone$KO
flux_list <- c("flux.acetate", "flux.acetone", "Condition", "Site")

idx <- match(c(KO_acetate_acetone_list,flux_list), names(merged_subset))
idx <- idx[!is.na(idx)]

sub_acetate_acetone_flux_genes <- merged_subset[,idx]

sub_acetate_acetone_flux_genes_2 <- sub_acetate_acetone_flux_genes %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)

sub_acetate_acetone_flux_genes_2
```

#filter to acetone flux and acetone genes
```{r}
KO_acetone <- KO_interest_subset %>%
  filter(VOC == "Acetone") 

KO_acetone_list <- KO_acetone$KO
flux_list <- c("flux.acetone", "Condition")

idx <- match(c(KO_acetone_list,flux_list), names(merged_subset))
idx <- idx[!is.na(idx)]

sub_acetone <- merged_subset[,idx]

sub_acetone2 <- sub_acetone %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition)

sub_acetone2
```
#filter to all VOC cycling genes with acetate, acetone, and diacetyl fluxes separately, then all three fluxes together
```{r}
KO_all_genes <- KO_interest_subset %>%
  filter(VOC == "Acetone" |
           VOC == "Acetate" |
           VOC == "Diacetyl")

#add genes to log transformed aceate flux data
KO_all_genes_list <- KO_all_genes$KO
flux_list_acetate <- c("log.flux.acetate", "flux.acetate", "Condition", "Time.d")

idx <- match(c(KO_all_genes_list,flux_list_acetate), names(merged))
idx <- idx[!is.na(idx)]

KO_all_genes_acetate_list <- merged[,idx]

KO_all_genes_acetate_list2 <- KO_all_genes_acetate_list %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition)

KO_all_genes_acetate_list2

#add genes to log transformed aceetone flux data
flux_list_acetone <- c("log.flux.acetone", "Condition", "Time.d")

idx <- match(c(KO_all_genes_list,flux_list_acetone), names(merged))
idx <- idx[!is.na(idx)]

KO_all_genes_acetone_list <- merged[,idx]

KO_all_genes_acetone_list2 <- KO_all_genes_acetone_list %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition)

KO_all_genes_acetone_list2

#add genes to log transformed diaceytl flux data
flux_list_diacetyl <- c("log.flux.diacetyl", "Condition", "Time.d")

idx <- match(c(KO_all_genes_list,flux_list_diacetyl), names(merged))
idx <- idx[!is.na(idx)]

KO_all_genes_diacetyl_list <- merged[,idx]

KO_all_genes_diacetyl_list2 <- KO_all_genes_diacetyl_list %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition)

KO_all_genes_diacetyl_list2

#add genes to log transformed diacetyle, acetate, and acetone flux data
flux_list_VOC <- c("log.flux.diacetyl", "log.flux.acetate", "log.flux.acetone", "Condition", "Time.d")

idx <- match(c(KO_all_genes_list,flux_list_VOC), names(merged))
idx <- idx[!is.na(idx)]

KO_all_genes_VOC_list <- merged[,idx]

KO_all_genes_VOC_list2 <- KO_all_genes_VOC_list %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition)

KO_all_genes_VOC_list2
```


```{r pls-single-responsel-acetate, echo = FALSE}
#single flux - acetate with acetete specific genes
pls.reg1 <- plsreg1(sub_acetate2[,1:8], sub_acetate[,9], comps = 4, crosval = TRUE)

pls.reg1$x.scores
pls.reg1$x.loads
pls.reg1$y.scores
pls.reg1$y.loads
pls.reg1$R2
pls.reg1$cor.xyt
pls.reg1$R2Xy
pls.reg1$raw.wgs
pls.reg1$mod.wgs
pls.reg1$y.pred
pls.reg1$std.coefs
pls.reg1$reg.coefs
pls.reg1$resid

plot(pls.reg1)

```
```{r pls-double-response-acetate-acetone-flux, echo = FALSE}
#single flux - acetate with acetete specific genes
res.col <- c(1:8, 11)

pls.reg2 <- plsreg2(sub_acetate_acetone_flux2[,res.col], sub_acetate_acetone_flux2[,9:10], comps=3, crosval = TRUE)

pls.reg2$x.scores
pls.reg2$x.loads
pls.reg2$y.scores
pls.reg2$y.loads
pls.reg2$R2
pls.reg2$cor.xyt
pls.reg2$R2Xy
pls.reg2$raw.wgs
pls.reg2$mod.wgs
pls.reg2$y.pred
pls.reg2$std.coefs
pls.reg2$reg.coefs
pls.reg2$resid
pls.reg2$VIP

plot(pls.reg2)
```

```{r}
#double flux - acetate and acetone with acetete and acetone specific genes
res.col <- c(1:14, 17)

pls.reg2 <- plsreg2(sub_acetate_acetone_flux_genes_2[,res.col], sub_acetate_acetone_flux_genes_2[,15:16], comps=3, crosval = TRUE)

pls.reg2$x.scores
pls.reg2$x.loads
pls.reg2$y.scores
pls.reg2$y.loads
pls.reg2$R2
pls.reg2$cor.xyt
pls.reg2$R2Xy
pls.reg2$raw.wgs
pls.reg2$mod.wgs
pls.reg2$y.pred
pls.reg2$std.coefs
pls.reg2$reg.coefs
pls.reg2$resid
pls.reg2$VIP

plot(pls.reg2)
```

```{r}
res.col <- c(1:6, 8)

pls.reg1 <- plsreg1(sub_acetone2[,res.col], sub_acetate_acetone_flux_genes_2[,9], comps=2, crosval = TRUE)

pls.reg1$x.scores
pls.reg1$x.loads
pls.reg1$y.scores
pls.reg1$y.loads
pls.reg1$R2
pls.reg1$cor.xyt
pls.reg1$R2Xy
pls.reg1$raw.wgs
pls.reg1$mod.wgs
pls.reg1$y.pred
pls.reg1$std.coefs
pls.reg1$reg.coefs
pls.reg1$resid

plot(pls.reg1)

```

```{r}
KO_all_genes_acetate_list2$Time.d <- as.numeric(KO_all_genes_acetate_list2$Time.d)
KO_all_genes_acetate_list2 <- KO_all_genes_acetate_list2 %>%
  mutate(log_scaled = (log.flux.acetate * 1))

res.col <- c(1:17)

pls.reg1 <- plsreg1(KO_all_genes_acetate_list2[,res.col], KO_all_genes_acetate_list2[,22], comps=2, crosval = TRUE)

pls.reg1$x.scores
pls.reg1$x.loads
pls.reg1$y.scores
pls.reg1$y.loads
pls.reg1$R2
pls.reg1$cor.xyt
pls.reg1$R2Xy
pls.reg1$raw.wgs
pls.reg1$mod.wgs
pls.reg1$y.pred
pls.reg1$std.coefs
pls.reg1$reg.coefs
pls.reg1$residn
pls.reg1$Q2

plot(pls.reg1)
```

```{r}
KO_all_genes_acetone_list2$Time.d <- as.numeric(KO_all_genes_acetone_list2$Time.d)
res.col <- c(1:17)


pls.reg1 <- plsreg1(KO_all_genes_acetone_list2[,res.col], KO_all_genes_acetone_list2[,18], comps=2, crosval = TRUE)

pls.reg1$x.scores
pls.reg1$x.loads
pls.reg1$y.scores
pls.reg1$y.loads
pls.reg1$R2
pls.reg1$cor.xyt
pls.reg1$R2Xy
pls.reg1$raw.wgs
pls.reg1$mod.wgs
pls.reg1$y.pred
pls.reg1$std.coefs
pls.reg1$reg.coefs
pls.reg1$resid
pls.reg1$Q2

plot(pls.reg1)
```

```{r}
KO_all_genes_diacetyl_list2$Time.d <- as.numeric(KO_all_genes_diacetyl_list2$Time.d)
res.col <- c(1:17)


pls.reg1 <- plsreg1(KO_all_genes_diacetyl_list2[,res.col], KO_all_genes_diacetyl_list2[,18], comps=2, crosval = TRUE)

pls.reg1$x.scores
pls.reg1$x.loads
pls.reg1$y.scores
pls.reg1$y.loads
pls.reg1$R2
pls.reg1$cor.xyt
pls.reg1$R2Xy
pls.reg1$raw.wgs
pls.reg1$mod.wgs
pls.reg1$y.pred
pls.reg1$std.coefs
pls.reg1$reg.coefs
pls.reg1$resid
pls.reg1$Q2

plot(pls.reg1)
```




```{r}
res.col <- c(1:17)

pls.reg2 <- plsreg2(KO_all_genes_VOC_list2[,res.col], KO_all_genes_VOC_list2[,18:20], comps=2, crosval = TRUE)

pls.reg2$x.scores
pls.reg2$x.loads
pls.reg2$y.scores
pls.reg2$y.loads
pls.reg2$R2
pls.reg2$cor.xyt
pls.reg2$R2Xy
pls.reg2$raw.wgs
pls.reg2$mod.wgs
pls.reg2$y.pred
pls.reg2$std.coefs
pls.reg2$reg.coefs
pls.reg2$resid
pls.reg2$Q2

plot(pls.reg2)
```

# try r base plsr function
```{r}
res.col <- c(1:16, 20)
x.matrix <- as.matrix(KO_all_genes_VOC_list2[,res.col])
y.matrix <- as.matrix(KO_all_genes_VOC_list2[,17:19])

even <- c(2,4,6,8,10,12,14,16)
odd <- c(1,3,5,7,9,11,13,15,17)

x.matrix.test <- x.matrix[even,]
x.matrix.val <- x.matrix[odd,]

y.matrix.test <- y.matrix[even,]
y.matrix.val <- y.matrix[odd,]

my_plsr <- plsr(y.matrix.test ~ x.matrix.test, ncomp=6, scale = TRUE, validation='LOO')
plot(RMSEP(my_plsr))
plot(my_plsr, plottype = "coef", ncomp=c(1:7), legendpos= "bottomleft")


# Loop through possible values for n_comp to optimize R2 on validation data
best_r2 <- 0
best_ncomp <- 0
for (n_comp in c(1:7)){
  model <- plsr(y.matrix.test ~ x.matrix.test, ncomp=n_comp, scale = TRUE, validation='LOO')
  predictions <- as.matrix(data.frame(predict(model, ncomp=n_comp, x.matrix.val)))
  mean_r2 <- mean(diag(cor(predictions, y.matrix.val))**2)

  if (mean_r2 > best_r2){
    best_r2 <- mean_r2
    best_ncomp <- n_comp
  }
}

print(best_ncomp)
print(best_r2)
```


```{r}
res.col <- c(1:16, 18)
x.matrix <- as.matrix(KO_all_genes_acetate_list2[,res.col])
y.matrix <- as.matrix(KO_all_genes_acetate_list2[,17])

even <- c(2,4,6,8,10,12,14,16,18)
odd <- c(1,3,5,7,9,11,13,15,17)

x.matrix.test <- x.matrix[even,]
x.matrix.val <- x.matrix[odd,]

y.matrix.test <- y.matrix[even,]
y.matrix.val <- y.matrix[odd,]

my_plsr <- plsr(y.matrix.test ~ x.matrix.test, ncomp=8, scale = TRUE)
plot(RMSEP(my_plsr))
plot(my_plsr, plottype = "coef", ncomp=c(1:8), legendpos= "bottomleft")


# Loop through possible values for n_comp to optimize R2 on validation data
best_r2 <- 0
best_ncomp <- 0
for (n_comp in c(1:7)){
  model <- plsr(y.matrix.test ~ x.matrix.test, ncomp=n_comp, scale = TRUE, validation='LOO')
  predictions <- as.matrix(data.frame(predict(model, ncomp=n_comp, x.matrix.val)))
  mean_r2 <- mean(diag(cor(predictions, y.matrix.val))**2)

  if (mean_r2 > best_r2){
    best_r2 <- mean_r2
    best_ncomp <- n_comp
  }
}

print(best_ncomp)
print(best_r2)
```


```{r}
res.col <- c(1:16, 18)
x.matrix <- as.matrix(KO_all_genes_acetone_list2[,res.col])
y.matrix <- as.matrix(KO_all_genes_acetone_list2[,17])

even <- c(2,4,6,8,10,12,14,16,18)
odd <- c(1,3,5,7,9,11,13,15,17)

x.matrix.test <- x.matrix[even,]
x.matrix.val <- x.matrix[odd,]

y.matrix.test <- y.matrix[even,]
y.matrix.val <- y.matrix[odd,]

my_plsr <- plsr(y.matrix.test ~ x.matrix.test, ncomp=8, scale = TRUE)
plot(RMSEP(my_plsr))
plot(my_plsr, plottype = "coef", ncomp=c(1:8), legendpos= "bottomleft")


# Loop through possible values for n_comp to optimize R2 on validation data
best_r2 <- 0
best_ncomp <- 0
for (n_comp in c(1:7)){
  model <- plsr(y.matrix.test ~ x.matrix.test, ncomp=n_comp, scale = TRUE, validation='LOO')
  predictions <- as.matrix(data.frame(predict(model, ncomp=n_comp, x.matrix.val)))
  mean_r2 <- mean(diag(cor(predictions, y.matrix.val))**2)

  if (mean_r2 > best_r2){
    best_r2 <- mean_r2
    best_ncomp <- n_comp
  }
}

print(best_ncomp)
print(best_r2)

# predict fit
pls.pred <- predict(my_plsr, x.matrix.val, ncomp=1:2)
summary(pls.pred)

plot(my_plsr, ncomp=4, asp=1, line=TRUE)


pls.pred2 = predict(my_plsr, x.matrix.test, ncomp=4)
plot(y.matrix.val, pls.pred2, ylim=c(0,3), xlim=c(0,3),main="Test Dataset", xlab="observed", ylab="PLS Predicted")
abline(0, 1, col="red")

pls.eval=data.frame(obs=y.matrix.val, pred=pls.pred2[,1,1])
defaultSummary(pls.eval)

```

