---
title: "PLS"
author: "Linnea Honeker"
date: "11/1/2022"
output: html_document
email: linneah@arizona.edu
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(ggplot2)
library(rms)
library(reshape2)
library(coefplot)
library(MASS)
library(dplyr)
library(lme4)
library(MASS)
library(leaps)
library(car)
library(relaimpo)
library(nlme)
library(lmerTest)
library(plsdepot)
library(pls)
library(caret)
library(DESeq2)
library(tibble)
library(metabom8)
library(plsVarSel)


#library(metabom8)

```

# load metaT data and do variance stabilization tranformation
```{r load-metaT, echo = FALSE}
#load metaT gene copy data and column metadata
datExpr0 <-read.csv("../../Data/Deseq2/Drought_vs_predrought_metaT_gene_copy.csv", header = TRUE)
rownames(datExpr0) <- datExpr0$Feature
datExpr0$Feature <- NULL

coldata <- read.csv("../../Data/Deseq2/metaT-metadata.csv", header = TRUE)
rownames(coldata) <- coldata$SampleID
coldata$SampleID <- NULL
row.names(coldata) [7] <- "X3300045460_Site1_Drought_48hr"

#filter features that are not in at least 5 samples
datExpr.f <- datExpr0 %>%
  mutate(count = rowSums(. > 0)) %>%
  filter(count > 5) %>%
  subset(select = -count )

#normalize filtered gene copy data using deseq2, variance stabilization transformation
dds <- DESeqDataSetFromMatrix(countData = datExpr.f,
                              colData = coldata,
                              design= ~ Condition_Time)

#vst transformation#
vstcounts <- vst(dds, blind=TRUE)
vst.export<- assay(vstcounts)
metaT<- as.data.frame(vst.export)
metaT$X <- rownames(metaT)

#log transformed
metaT.log <- log(datExpr.f + 1)
```


```{r load voc flux data, echo=FALSE}
acetate_flux <- read.csv("../../Output/CO2-VOCs/acetate_flux-13C-12C-0-48.csv", header = TRUE)
acetone_flux <- read.csv("../../Output/CO2-VOCs/acetone_flux-13C-12C-0-48.csv", header = TRUE)
diacetyl_flux <- read.csv("../../Output/CO2-VOCs/diacetyl_flux-13C-12C-0-48.csv", header = TRUE)

```


```{r merge KO and metaT and flux data}

KO_interest_subset <- read.csv("../../Data/Deseq2/KO_interest.csv")
metadata <- read.csv("../../Data/Deseq2/metaT-metadata.csv")

#remove "KO:" from metaT and "_C1" and "_C2" from diaceytl flux data
metaT$KO <- rownames(metaT)
metaT$KO <- sub("KO:", "", metaT$KO)
#diacetyl_flux$Sample <- sub("_C1", "", diacetyl_flux$Sample)
#diacetyl_flux$Sample <- sub("_C2", "", diacetyl_flux$Sample)


#merge  all VOC
KO_VOC <- KO_interest_subset %>%
  merge(., metaT, by.x = "KO", by.y="KO", all.x = FALSE, all.y = FALSE)

KO_VOC$X <- NULL

#transpose for plsr
KO_VOC.t <- t(KO_VOC)

#further format KO_acetate.t for glm. Assign column headers and remove rows containing metadata
colnames(KO_VOC.t) <- KO_VOC.t[1,]
KO_VOC.t.f <- KO_VOC.t[-(1:6),]
KO_VOC.t.f <- as.data.frame(KO_VOC.t.f)

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetate_flux_proc <- acetate_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  #mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.h > 0 & Time.h < 2) ~ "0",
    (Time.h > 5 & Time.h < 7) ~ "6",
    (Time.h > 46 & Time.h < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise(mean.flux = mean(Flux)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetate = mean.flux) %>%
  dplyr::select(-"mean.flux")

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetone_flux_proc <- acetone_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  #mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.h > 0 & Time.h < 2) ~ "0",
    (Time.h > 5 & Time.h < 7) ~ "6",
    (Time.h > 46 & Time.h < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise(mean.flux = mean(Flux)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetone = mean.flux) %>%
  dplyr::select(-"mean.flux")

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
diacetyl_flux_proc <- diacetyl_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  #mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.h > 0 & Time.h < 2) ~ "0",
    (Time.h > 5 & Time.h < 7) ~ "6",
    (Time.h > 46 & Time.h < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise(mean.flux = mean(Flux)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.diacetyl = mean.flux) %>%
  dplyr::select(-"mean.flux")

#sample names are different in the flux and expression tables, so need to make them the same
KO_VOC.t.f[,31] <- rownames(KO_VOC.t.f)
KO_VOC.t.f_new_id_w_full_name <- KO_VOC.t.f %>%
  mutate(Sample = case_when(
  startsWith(V31, "X3300044942") ~ "S3P4",
  startsWith(V31, "X3300044943") ~ "S3P4",
  startsWith(V31, "X3300044944") ~ "S3P4",
  startsWith(V31, "X3300044945") ~ "S3P4",
  startsWith(V31, "X3300045458") ~ "S1P5",
  startsWith(V31, "X3300045459") ~ "S1P5",
  startsWith(V31, "X3300045460") ~ "S1P5",
  startsWith(V31, "X3300045461") ~ "S2P4",
  startsWith(V31, "X3300045462") ~ "S2P4",
  startsWith(V31, "X3300045463") ~ "S2P4",
  startsWith(V31, "X3300045464") ~ "S2P4",
  startsWith(V31, "X3300045465") ~ "S2P4",
  startsWith(V31, "X3300045466") ~ "S2P6",
  startsWith(V31, "X3300045467") ~ "S2P6",
  startsWith(V31, "X3300045468") ~ "S2P6",
  startsWith(V31, "X3300045469") ~ "S2P6",
  startsWith(V31, "X3300045470") ~ "S2P6",
  startsWith(V31, "X3300045471") ~ "S3P5",
  startsWith(V31, "X3300045472") ~ "S3P5",
  startsWith(V31, "X3300045473") ~ "S2P6",
  startsWith(V31, "X3300048051") ~ "S1P5",
  startsWith(V31, "X3300048052") ~ "S3P5",
  startsWith(V31, "X3300048053") ~ "S3P5",
  startsWith(V31, "X3300048054") ~ "S3P5",
  startsWith(V31, "X3300048092") ~ "S1P1",
  startsWith(V31, "X3300048093") ~ "S1P1",
  startsWith(V31, "X3300048094") ~ "S1P1",
  startsWith(V31, "X3300048095") ~ "S1P1",
  startsWith(V31, "X3300048096") ~ "S1P1",
  startsWith(V31, "X3300048099") ~ "S1P5",
  startsWith(V31, "X3300048100") ~ "S1P5",
  startsWith(V31, "X3300049511") ~ "S2P4",
  startsWith(V31, "X3300049512") ~ "S3P4",
  startsWith(V31, "X3300049557") ~ "S3P4",
  startsWith(V31, "X3300049628") ~ "S3P5"
  )) %>%
  mutate(Time.d = case_when(
    endsWith(V31, "6hr") ~ "6",
    endsWith(V31, "0hr") ~ "0",
    endsWith(V31, "48hr") ~ "48",
    endsWith(V31, "48_hr") ~ "48"
  )) %>%
  mutate(Condition = case_when(
    grepl("PreDrought", V31) ~ "Pre-Drought",
    grepl("Drought", V31) ~ "Drought",
    grepl("Pre_drought", V31) ~ "Pre-Drought"
  )) %>%
  mutate(Site = case_when(
    grepl("Site1", V31) ~ "Site1",
    grepl("Site2", V31) ~ "Site2",
    grepl("Site3", V31) ~ "Site3"
  )) 

KO_VOC.t.f_new_id <- KO_VOC.t.f_new_id_w_full_name %>%
   dplyr::select(-V31)
  
#merge VOC and KO tables
merged_subset <- KO_VOC.t.f_new_id %>%
  merge(., acetate_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label","Time.d", "Site"),all.x = FALSE, all.y = FALSE) %>%
  mutate_at(c(5:35), as.numeric) %>%
  mutate(flux.acetate = as.numeric(flux.acetate)) %>%
  mutate(log.flux.acetate = log(flux.acetate+abs(min(flux.acetate))+1, base = exp(10))) %>%
  merge(.,acetone_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label", "Time.d", "Site"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.acetone = as.numeric(flux.acetone)) %>%
  mutate(log.flux.acetone = log(flux.acetone+abs(min(flux.acetone))+1, base = exp(10))) %>%
  merge(.,diacetyl_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label", "Time.d", "Site"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.diacetyl = as.numeric(flux.diacetyl)) %>%
  mutate(log.flux.diacetyl = log(flux.diacetyl+abs(min(flux.diacetyl))+1, base = exp(2))) 
  
merged_subset

merged_subset_with_full_names <- KO_VOC.t.f_new_id_w_full_name %>%
  merge(., acetate_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label","Time.d", "Site"),all.x = TRUE, all.y = TRUE) %>%
  mutate_at(c(5:35), as.numeric) %>%
  mutate(flux.acetate = as.numeric(flux.acetate)) %>%
  mutate(log.flux.acetate = log(flux.acetate+abs(min(flux.acetate))+1, base = exp(10))) %>%
  merge(.,acetone_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label", "Time.d", "Site"),all.x = TRUE, all.y = TRUE) %>%
  mutate(flux.acetone = as.numeric(flux.acetone)) %>%
  mutate(log.flux.acetone = log(flux.acetone+abs(min(flux.acetone))+1, base = exp(10))) %>%
  merge(.,diacetyl_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label", "Time.d", "Site"),all.x = TRUE, all.y = TRUE) %>%
  mutate(flux.diacetyl = as.numeric(flux.diacetyl)) %>%
  mutate(log.flux.diacetyl = log(flux.diacetyl+abs(min(flux.diacetyl))+1, base = exp(2))) 

merged_subset_with_full_names


#merge with soil moisture data, which is loaded in the 'soil_moisture.Rmd' script
merged_subset_SM <- merged_subset %>%
  merge(., data.abiotic, by.x = c("Condition", "Site"), by.y = c("Condition", "Site")) 

merged_subset_physico_full_names <- merged_subset_with_full_names %>%
  merge(., data.abiotic, by = c("Condition", "Site"))

#create trait data file for WGCNA
trait_data_for_WGCNA <- merged_subset_physico_full_names[,c(36,37,39,41, 42,43,44)]
trait_data_for_WGCNA$SampleID <- trait_data_for_WGCNA[,1]
trait_data_for_WGCNA$V31 <- NULL
trait_data_for_WGCNA <- trait_data_for_WGCNA[,c(7,1,2,3, 4, 5, 6)]

write.csv(trait_data_for_WGCNA, "../../Output/WGCNA/trait-data-VOC-flux-physicochemical.csv")

```



## Filter to acetate and acetone flux and acetate cycling genes with soil moisture
```{r}
KO_acetate <- KO_interest_subset %>%
  filter(VOC == "Acetate" &
           Priority == "yes") 

KO_acetate_list <- KO_acetate$KO
flux_list.log <- c(  "SM", "Condition", "Site", "log.flux.acetone", "log.flux.acetate")
flux_list <- c( "SM", "Condition", "Site", "flux.acetone","flux.acetate")

idx <- match(c(KO_acetate_list,flux_list), names(merged_subset_SM))
idx <- idx[!is.na(idx)]

sub_acetate_pls2_SM <- merged_subset_SM[,idx]

sub_acetate_pls2_SM <- sub_acetate_pls2_SM %>%
  #mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)


sub_acetate_pls2_SM
```


#####acetate - no pH or ST#####   
# first round:

all variables

p-value = 0.72

total variation explained = 90.7 (5 components)


  K00156       K00158       K01026       K01067       K18118       K01512       K01895       K00925       K18372           SM flux.acetone 
   0.8952097    0.9217706    0.8023164    1.2356767    1.0521327    0.8440945    0.8450783    0.8103762    0.7135039    1.0181347    1.5627854
   
# second round
new model set = c(1,2,3,4,5,6,7,8,10, 11,12) 

p-value = 0.50

total variation explained = 91.04 (6 components) ***BEST MODEL***
   
VIP:
    K00156       K00158       K01026       K01067       K18118       K01512       K01895       K00925           SM flux.acetone 
   0.8445666    0.8664567    0.8199626    1.2078264    1.0774193    0.8583763    0.8247477    0.7266882    1.0047424    1.5130575 
   


# third round
new model set = c(1,2,3,4,5,6,7,10, 11,12) 

total variation explained = 85.01 (4 components)
p-value = 0.83

VIP:
 K00156       K00158       K01026       K01067       K18118       K01512       K01895           SM flux.acetone 
   0.8955157    0.9313750    0.6754161    1.1742540    0.8588176    0.9151972    0.8467789    1.0297782    1.4638654 
   


set up model data frame using Johnny's scripts and the plsr function of the pls package
```{r}
model_data <- sub_acetate_pls2_SM[, c(1,2,3,4,5,6,7,8,9,10, 11) ]

set.seed(123)
m_full <- plsr(flux.acetate ~ ., data = model_data, scale = TRUE, validation = "LOO")
summary(m_full)

par(mfrow = c(3,1))
validationplot(m_full)
validationplot(m_full, val.type = "RMSEP")
validationplot(m_full, val.type = "R2")

```

```{r}

m1 <- plsr(flux.acetate ~ ., data = model_data, 
     scale = TRUE, ncomp = 6) 
summary(m1)
plot(m1)
```

create a plsr and null model to compare and calculate residuals of both
```{r}
model_vals <- data.frame(
  preds = numeric(length = nrow(model_data)), 
  res0 = numeric(length = nrow(model_data)),
  res1 = numeric(length = nrow(model_data))
)


for(i in 1:nrow(model_data)){
  y <- model_data$flux.acetate
  
  m <- plsr(flux.acetate ~ ., data = model_data[-i,], 
     scale = TRUE, ncomp =6)
  model_vals$preds[i] <- predict(m, newdata = model_data[i,], 
                                 ncomp = 6, type = "response")
  # predicted residual of the model
  model_vals$res1[i] <-y[i] - model_vals$preds[i]
  
  # predicted residual using the null model
  y_mean = mean(y[-i])
  model_vals$res0[i] = y[i] - y_mean
}

plot(model_data$flux.acetate, model_vals$preds)

```
Use F-Test to compare to null model

This is done analogues to comparing a regression model to a null model, 
i.e. by comparing the Residual some of squares of model and null model, calculating the F value using the respective degrees of freedom of the two models and getting the two-sided p-value via the cumulative distribution function of the F distribution. 

```{r}
RSS1 = sum((model_vals$res1)^2)
RSS0 = sum((model_vals$res0)^2)

df1 = nrow(model_data) - (ncol(model_data) - 1) - 1
df0 = nrow(model_data) - 1

F_value <- ((RSS0 - RSS1) / (df0 - df1)) / (RSS1 / df1)
c(F = F_value, df0 = df0, df1 = df1, RSS0 = RSS0, RSS1 = RSS1)

p.value <-  pf(F_value, df0, df1, lower.tail = FALSE)
p.value

```
#use plsVarSel to slect combination of variables leading to model with lowest prediction error

```{r}

VIP(m1, opt.comp=6, p=10)

LW(m1, 6)

RC(m1, 6)

plot(m1, "loadings", comps = 1:2)


```


## Filter to acetone and acetate flux and acetone cycling genes with soil moisture
```{r}
KO_acetone <- KO_interest_subset %>%
  filter(VOC == "Acetone") 

KO_acetone_list <- KO_acetone$KO
flux_list.log <- c(  "SM", "Condition", "Site", "log.flux.acetate", "log.flux.acetone")
flux_list <- c( "SM", "Condition", "Site", "flux.acetate","flux.acetone")

idx <- match(c(KO_acetone_list,flux_list), names(merged_subset_SM))
idx <- idx[!is.na(idx)]

sub_acetone_pls2_SM <- merged_subset_SM[,idx]

sub_acetone_pls2_SM <- sub_acetone_pls2_SM %>%
  #mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)


sub_acetone_pls2_SM
```
`

   
###acetone - no pH or ST#######
# first round:

all variables

p-value = 0.32

total variation explained = 75.95 (2 components)



   K01574       K18371       K18372       K10854       K10855       K10856       K18382           SM flux.acetate 
   1.3258560    0.8373220    0.6871190    0.2334001    0.5480074    0.6658359    0.9843501    1.2183880    1.6784891 
   
# second round
new model set = c(1,2,3,5,6,7,8,9, 10)

p-value = 0.19

total variation explained = 76.33 (2 components)
   
  K01574       K18371       K18372       K10855       K10856       K18382           SM flux.acetate 
   1.2543362    0.7972028    0.6503456    0.5320875    0.6454280    0.9257862    1.1463581    1.5802682 

# third round
new model set =  c(1,2,3,6,7,8,9, 10)

p-value = 0.10

total variation explained = 76.25 (2 components)

VIP:
K01574       K18371       K18372       K10856       K18382           SM flux.acetate 
   1.1888973    0.7541031    0.6384944    0.5693243    0.8849511    1.1030446    1.5120186 
   
# fourth round
new model set =  c(1,2,3,7,8,9, 10)


total variation explained = 76.15 (2 components)

p-value = 0.05

VIP:
  K01574       K18371       K18372       K18382           SM           ST flux.acetate 
   1.1297646    0.6857334    0.6067197    0.8661450    1.0527979    1.0150584    1.4129256 
   
# fifth round
new model set = c(1,2,7,8,9, 10)

total variation explained = 76.37(1 components) ***BEST ONE***

p-value = 0.004

VIP:
     K01574       K18371       K18382           SM flux.acetate 
   1.0483239    0.6430058    0.8000718    0.9957126    1.3623517 
   
# sixth round
new model set = c(1,7,8,9, 10)

total variation explained = 74.59 (2 components) 

p-value = 0.008

VIP:
  K01574       K18382           SM flux.acetate 
   0.9905808    0.7456982    0.9181682    1.2726551 
   



set up model data frame using Johnny's scripts and the plsr function of the pls package
```{r}
model_data <- sub_acetone_pls2_SM[, c(1,2,7,8,9, 10)]

set.seed(123)
m_full <- plsr(flux.acetone ~ ., data = model_data, scale = TRUE, validation = "LOO")
summary(m_full)

par(mfrow = c(3,1))
validationplot(m_full)
validationplot(m_full, val.type = "RMSEP")
validationplot(m_full, val.type = "R2")

```

```{r}

m1 <- plsr(flux.acetone ~ ., data = model_data, 
     scale = TRUE, ncomp = 1) 
summary(m1)
plot(m1)
```

create a plsr and null model to compare and calculate residuals of both
```{r}
model_vals <- data.frame(
  preds = numeric(length = nrow(model_data)), 
  res0 = numeric(length = nrow(model_data)),
  res1 = numeric(length = nrow(model_data))
)


for(i in 1:nrow(model_data)){
  y <- model_data$flux.acetone
  
  m <- plsr(flux.acetone ~ ., data = model_data[-i,], 
     scale = TRUE, ncomp = 1)
  model_vals$preds[i] <- predict(m, newdata = model_data[i,], 
                                 ncomp = 1, type = "response")
  # predicted residual of the model
  model_vals$res1[i] <-y[i] - model_vals$preds[i]
  
  # predicted residual using the null model
  y_mean = mean(y[-i])
  model_vals$res0[i] = y[i] - y_mean
}

plot(model_data$flux.acetone, model_vals$preds)

```
Use F-Test to compare to null model

This is done analogues to comparing a regression model to a null model, 
i.e. by comparing the Residual some of squares of model and null model, calculating the F value using the respective degrees of freedom of the two models and getting the two-sided p-value via the cumulative distribution function of the F distribution. 

```{r}
RSS1 = sum((model_vals$res1)^2)
RSS0 = sum((model_vals$res0)^2)

df1 = nrow(model_data) - (ncol(model_data) - 1) - 1
df0 = nrow(model_data) - 1

F_value <- ((RSS0 - RSS1) / (df0 - df1)) / (RSS1 / df1)
c(F = F_value, df0 = df0, df1 = df1, RSS0 = RSS0, RSS1 = RSS1)

p.value <-  pf(F_value, df0, df1, lower.tail = FALSE)
p.value

```
#use plsVarSel to slect combination of variables leading to model with lowest prediction error

```{r}

VIP(m1, opt.comp =1, p =5)

LW(m1, opt.comp = 1)


```



## Filter to acetone and acetone flux and acetone cycling genes with soil moisture
```{r}
KO_diacetyl <- KO_interest_subset %>%
  filter(VOC == "Diacetyl" ) 

KO_diacetyl_list <- KO_diacetyl$KO
flux_list.log <- c(  "SM", "Condition", "Site", "log.flux.diacetyl")
flux_list <- c( "SM", "Condition", "Site","flux.diacetyl")

idx <- match(c(KO_diacetyl_list,flux_list), names(merged_subset_SM))
idx <- idx[!is.na(idx)]

sub_diacetyl_pls2_SM <- merged_subset_SM[,idx]

sub_diacetyl_pls2_SM <- sub_diacetyl_pls2_SM %>%
  #mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)


sub_diacetyl_pls2_SM
```



##without pH or ST###
# first round:

p-value = 0.99
total variation explained = 24.39 (2 components)


  K01652    K03366    K00004        SM 
0.3352935 1.2300897 0.1059071 1.5372837 

# second round:
new model set = c(1,2,4,5)
p-value = 0.99
total variation explained = 24.23 (2 components)


  K01652    K03366        SM 
0.2897927 1.0640465 1.3355992 

# third round:
new model set = c(2,4,5)
p-value = 0.93
total variation explained = 22.82 (2 components) ***BEST MODEL***

  K03366        SM 
0.8573324 1.1247138 


  


```{r}
model_data <- sub_diacetyl_pls2_SM[, c(2,4,5)]

set.seed(123)
m_full <- plsr(flux.diacetyl ~ ., data = model_data, scale = FALSE, validation = "LOO")
summary(m_full)

par(mfrow = c(3,1))
validationplot(m_full)
validationplot(m_full, val.type = "MSEP")
validationplot(m_full, val.type = "R2")

```

```{r}

m1 <- plsr(flux.diacetyl ~ ., data = model_data, 
     scale = FALSE, ncomp = 2) 
summary(m1)
plot(m1)
```
create a plsr and null model to compare and calculate residuals of both
```{r}
model_vals <- data.frame(
  preds = numeric(length = nrow(model_data)), 
  res0 = numeric(length = nrow(model_data)),
  res1 = numeric(length = nrow(model_data))
)


for(i in 1:nrow(model_data)){
  y <- model_data$flux.diacetyl
  
  m <- plsr(flux.diacetyl ~ ., data = model_data[-i,], 
     scale = FALSE, ncomp = 2)
  model_vals$preds[i] <- predict(m, newdata = model_data[i,], 
                                 ncomp = 2, type = "response")
  # predicted residual of the model
  model_vals$res1[i] <-y[i] - model_vals$preds[i]
  
  # predicted residual using the null model
  y_mean = mean(y[-i])
  model_vals$res0[i] = y[i] - y_mean
}

plot(model_data$flux.diacetyl, model_vals$preds)

```

```{r}
RSS1 = sum((model_vals$res1)^2)
RSS0 = sum((model_vals$res0)^2)

df1 = nrow(model_data) - (ncol(model_data) - 1) - 1
df0 = nrow(model_data) - 1

F_value <- ((RSS0 - RSS1) / (df0 - df1)) / (RSS1 / df1)
c(F = F_value, df0 = df0, df1 = df1, RSS0 = RSS0, RSS1 = RSS1)

p.value <- pf(F_value, df0, df1, lower.tail = FALSE)
p.value

```

#use plsVarSel to slect combination of variables leading to model with lowest prediction error

```{r}

VIP(m1, opt.comp =2, p =2)

LW(m1, opt.comp = 2)

#sMC(m1, opt.comp = 6, model_data[,1:11])

```