---
title: "PLS"
author: "Linnea Honeker"
date: "11/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(ggplot2)
library(rms)
library(reshape2)
library(coefplot)
library(MASS)
library(dplyr)
library(lme4)
library(MASS)
library(leaps)
library(car)
library(relaimpo)
library(nlme)
library(lmerTest)
library(plsdepot)
library(pls)
library(caret)
library(DESeq2)
library(tibble)
library(metabom8)

#library(metabom8)

```

# load metaT data and do variance stabilization tranformation
```{r load-metaT, echo = FALSE}
#load metaT gene copy data and column metadata
datExpr0 <-read.csv("../../Data/Deseq2/Drought_vs_predrought_metaT_missing_1_no_feature.csv", header = TRUE)
rownames(datExpr0) <- datExpr0$Feature
datExpr0$Feature <- NULL

coldata <- read.csv("../../Data/Deseq2/metaT-metadata.csv", header = TRUE)
rownames(coldata) <- coldata$SampleID
coldata$SampleID <- NULL
row.names(coldata) [7] <- "X3300045460_Site1_Drought_48hr"

#filter features that are not in at least 5 samples
datExpr.f <- datExpr0 %>%
  mutate(count = rowSums(. > 0)) %>%
  filter(count > 5) %>%
  subset(select = -count )

#normalize filtered gene copy data using deseq2, variance stabilization transformation
dds <- DESeqDataSetFromMatrix(countData = datExpr.f,
                              colData = coldata,
                              design= ~ Condition_Time)

#vst transformation#
vstcounts <- vst(dds, blind=TRUE)
vst.export<- assay(vstcounts)
metaT<- as.data.frame(vst.export)
metaT$X <- rownames(metaT)

#log transformed
metaT.log <- log(datExpr.f + 1)
```


```{r load voc flux data, echo=FALSE}
acetate_flux <- read.csv("../../Output/CO2-VOCs/acetate_flux-13C-12C-0-48.csv", header = TRUE)
acetone_flux <- read.csv("../../Output/CO2-VOCs/acetone_flux-13C-12C-0-48.csv", header = TRUE)
diacetyl_flux <- read.csv("../../Output/CO2-VOCs/diacetyl_flux-13C-12C-0-48.csv", header = TRUE)

```


```{r merge KO and metaT and flux data}

KO_interest_subset <- read.csv("../../Data/Deseq2/KO_interest.csv")
metadata <- read.csv("../../Data/Deseq2/metaT-metadata.csv")

#remove "KO:" from metaT and "_C1" and "_C2" from diaceytl flux data
metaT$KO <- rownames(metaT)
metaT$KO <- sub("KO:", "", metaT$KO)
#diacetyl_flux$Sample <- sub("_C1", "", diacetyl_flux$Sample)
#diacetyl_flux$Sample <- sub("_C2", "", diacetyl_flux$Sample)


#merge  all VOC
KO_VOC <- KO_interest_subset %>%
  merge(., metaT, by.x = "KO", by.y="KO", all.x = FALSE, all.y = FALSE)

KO_VOC$X <- NULL

#transpose for plsr
KO_VOC.t <- t(KO_VOC)

#further format KO_acetate.t for glm. Assign column headers and remove rows containing metadata
colnames(KO_VOC.t) <- KO_VOC.t[1,]
KO_VOC.t.f <- KO_VOC.t[-(1:6),]
KO_VOC.t.f <- as.data.frame(KO_VOC.t.f)

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetate_flux_proc <- acetate_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  #mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.h > 0 & Time.h < 2) ~ "0",
    (Time.h > 5 & Time.h < 7) ~ "6",
    (Time.h > 46 & Time.h < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise(mean.flux = mean(Flux)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetate = mean.flux) %>%
  dplyr::select(-"mean.flux")

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetone_flux_proc <- acetone_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  #mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.h > 0 & Time.h < 2) ~ "0",
    (Time.h > 5 & Time.h < 7) ~ "6",
    (Time.h > 46 & Time.h < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise(mean.flux = mean(Flux)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetone = mean.flux) %>%
  dplyr::select(-"mean.flux")

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
diacetyl_flux_proc <- diacetyl_flux %>%
    filter (Pyr == "C2") %>%
     mutate(Site = case_when(
    startsWith(Label, "S1") ~ "Site1",
    startsWith(Label, "S2") ~ "Site2",
    startsWith(Label, "S3") ~ "Site3"
  )) %>%
  #mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.h > 0 & Time.h < 2) ~ "0",
    (Time.h > 5 & Time.h < 7) ~ "6",
    (Time.h > 46 & Time.h < 48) ~ "48")) %>%
  group_by(Condition, Site, Label, Time.d) %>%
  summarise(mean.flux = mean(Flux)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.diacetyl = mean.flux) %>%
  dplyr::select(-"mean.flux")

#sample names are different in the flux and expression tables, so need to make them the same
KO_VOC.t.f[,28] <- rownames(KO_VOC.t.f)
KO_VOC.t.f_new_id <- KO_VOC.t.f %>%
  mutate(Sample = case_when(
  startsWith(V28, "X3300044942") ~ "S3P4",
  startsWith(V28, "X3300044943") ~ "S3P4",
  startsWith(V28, "X3300044944") ~ "S3P4",
  startsWith(V28, "X3300044945") ~ "S3P4",
  startsWith(V28, "X3300045458") ~ "S1P5",
  startsWith(V28, "X3300045459") ~ "S1P5",
  startsWith(V28, "X3300045460") ~ "S1P5",
  startsWith(V28, "X3300045461") ~ "S2P4",
  startsWith(V28, "X3300045462") ~ "S2P4",
  startsWith(V28, "X3300045463") ~ "S2P4",
  startsWith(V28, "X3300045464") ~ "S2P4",
  startsWith(V28, "X3300045465") ~ "S2P4",
  startsWith(V28, "X3300045466") ~ "S2P6",
  startsWith(V28, "X3300045467") ~ "S2P6",
  startsWith(V28, "X3300045468") ~ "S2P6",
  startsWith(V28, "X3300045469") ~ "S2P6",
  startsWith(V28, "X3300045470") ~ "S2P6",
  startsWith(V28, "X3300045471") ~ "S3P5",
  startsWith(V28, "X3300045472") ~ "S3P5",
  startsWith(V28, "X3300045473") ~ "S2P6",
  startsWith(V28, "X3300048051") ~ "S1P5",
  startsWith(V28, "X3300048052") ~ "S3P5",
  startsWith(V28, "X3300048053") ~ "S3P5",
  startsWith(V28, "X3300048054") ~ "S3P5",
  startsWith(V28, "X3300048092") ~ "S1P1",
  startsWith(V28, "X3300048093") ~ "S1P1",
  startsWith(V28, "X3300048094") ~ "S1P1",
  startsWith(V28, "X3300048095") ~ "S1P1",
  startsWith(V28, "X3300048096") ~ "S1P1",
  startsWith(V28, "X3300048099") ~ "S1P5",
  startsWith(V28, "X3300048100") ~ "S1P5",
  startsWith(V28, "X3300049511") ~ "S2P4",
  startsWith(V28, "X3300049512") ~ "S3P4",
  startsWith(V28, "X3300049557") ~ "S3P4",
  startsWith(V28, "X3300049628") ~ "S3P5"
  )) %>%
  mutate(Time.d = case_when(
    endsWith(V28, "6hr") ~ "6",
    endsWith(V28, "0hr") ~ "0",
    endsWith(V28, "48hr") ~ "48",
    endsWith(V28, "48_hr") ~ "48"
  )) %>%
  mutate(Condition = case_when(
    grepl("PreDrought", V28) ~ "Pre-Drought",
    grepl("Drought", V28) ~ "Drought",
    grepl("Pre_drought", V28) ~ "Pre-Drought"
  )) %>%
  mutate(Site = case_when(
    grepl("Site1", V28) ~ "Site1",
    grepl("Site2", V28) ~ "Site2",
    grepl("Site3", V28) ~ "Site3"
  )) %>%
   dplyr::select(-V28)

#merge VOC and KO tables
merged_subset <- KO_VOC.t.f_new_id %>%
  merge(., acetate_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label","Time.d", "Site"),all.x = FALSE, all.y = FALSE) %>%
  mutate_at(c(5:32), as.numeric) %>%
  mutate(flux.acetate = as.numeric(flux.acetate)) %>%
  mutate(log.flux.acetate = log(flux.acetate+abs(min(flux.acetate))+1, base = exp(10))) %>%
  merge(.,acetone_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label", "Time.d", "Site"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.acetone = as.numeric(flux.acetone)) %>%
  mutate(log.flux.acetone = log(flux.acetone+abs(min(flux.acetone))+1, base = exp(10))) %>%
  merge(.,diacetyl_flux_proc, by.x = c("Condition", "Sample", "Time.d", "Site"), by.y = c("Condition", "Label", "Time.d", "Site"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.diacetyl = as.numeric(flux.diacetyl)) %>%
  mutate(log.flux.diacetyl = log(flux.diacetyl+abs(min(flux.diacetyl))+1, base = exp(2))) 
  
merged_subset

#merge with soil moisture data, which is loaded in the 'soil_moisture.Rmd' script
merged_subset_SM <- merged_subset %>%
  merge(., data.sm, by.x = c("Condition", "Site"), by.y = c("Condition", "Site")) %>%
  mutate(log.SM = log(SM, base = exp(10)))


merged_subet_no_0 <- merged_subset %>%
  filter(Time.d != "0")

merged_subet_no_0_SM <- merged_subset %>%
  filter(Time.d != "0")%>%
  merge(., data.sm, by = c("Condition", "Site")) %>%
  mutate(log.SM = log(SM))
```

## Filter to diacetyl and diacetyl cycling genes
```{r}
KO_diacetyl <- KO_interest_subset %>%
  filter(VOC == "Diacetyl") 

KO_diacetyl_list <- KO_diacetyl$KO
flux_list <- c("Condition", "Site", "log.flux.diacetyl")

idx <- match(c(KO_diacetyl_list,flux_list), names(merged_subset))
idx <- idx[!is.na(idx)]

sub_diacetyl_pls <- merged_subset[,idx]

sub_diacetyl_pls2 <- sub_diacetyl_pls %>%
  #mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)

sub_diacetyl_pls2


```
### diacetyl w/ diacetyl genes
```{r}

pls.reg1.diacetyl <- plsreg1(sub_diacetyl_pls2[,1:3], sub_diacetyl_pls2[,4], comps=2, crosval = TRUE)

pls.reg1.diacetyl$x.scores
pls.reg1.diacetyl$x.loads
pls.reg1.diacetyl$y.scores
pls.reg1.diacetyl$y.loads
pls.reg1.diacetyl$cor.xyt
pls.reg1.diacetyl$R2Xy
pls.reg1.diacetyl$raw.wgs
pls.reg1.diacetyl$mod.wgs
pls.reg1.diacetyl$y.pred
pls.reg1.diacetyl$std.coefs
pls.reg1.diacetyl$reg.coefs
pls.reg1.diacetyl$residn
pls.reg1.diacetyl$R2
pls.reg1.diacetyl$Q2

plot(pls.reg1.diacetyl)
```




## Filter to acetate and acetone flux and acetate cycling genes with soil moisture
```{r}
KO_acetate <- KO_interest_subset %>%
  filter(VOC == "Acetate") 

KO_acetate_list <- KO_acetate$KO
flux_list.log <- c(  "log.SM", "Condition", "Site", "log.flux.acetone", "log.flux.acetate")
flux_list <- c( "log.SM", "Condition", "Site", "flux.acetone","flux.acetate")

idx <- match(c(KO_acetate_list,flux_list.log), names(merged_subset_SM))
idx <- idx[!is.na(idx)]

sub_acetate_pls2_SM <- merged_subset_SM[,idx]

sub_acetate_pls2_SM <- sub_acetate_pls2_SM %>%
  #mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)


sub_acetate_pls2_SM
```

### acetate w/ acetate genes and acetone flux with SM
```{r}

pls.reg1.acetate <- plsreg1(sub_acetate_pls2_SM[,1:11], sub_acetate_pls2_SM[,12], comps=3, crosval = TRUE)

pls.reg1.acetate$x.scores
pls.reg1.acetate$x.loads
pls.reg1.acetate$y.scores
pls.reg1.acetate$y.loads
pls.reg1.acetate$cor.xyt
pls.reg1.acetate$R2Xy
pls.reg1.acetate$raw.wgs
pls.reg1.acetate$mod.wgs
pls.reg1.acetate$y.pred
pls.reg1.acetate$std.coefs
pls.reg1.acetate$reg.coefs
pls.reg1.acetate$resid
pls.reg1.acetate$R2
pls.reg1.acetate$Q2

plot(pls.reg1.acetate)
```
### acetate w/ acetate genes and acetone flux with SM without low drivers - develop plsr model with plsdepot
```{r}
res.col.acetate <- c(1,2,3,4,5,6,7,8,10,11)
pls.reg1.acetate <- plsreg1(sub_acetate_pls2_SM[,res.col.acetate], sub_acetate_pls2_SM[,12], comps=3, crosval = TRUE)

pls.reg1.acetate$x.scores
pls.reg1.acetate$x.loads
pls.reg1.acetate$y.scores
pls.reg1.acetate$y.loads
pls.reg1.acetate$cor.xyt
pls.reg1.acetate$R2Xy
pls.reg1.acetate$raw.wgs
pls.reg1.acetate$mod.wgs
pls.reg1.acetate$y.pred
pls.reg1.acetate$std.coefs
pls.reg1.acetate$reg.coefs
pls.reg1.acetate$residn
pls.reg1.acetate$R2
pls.reg1.acetate$Q2
pls.reg1.acetate$VIP

plot(pls.reg1.acetate)
```
set up model data frame using Johnny's scripts and the plsr function of the pls package
```{r}
model_data <- sub_acetate_pls2_SM[,c(1,2,3,4,5,6,7, 8,10,11,12)]

set.seed(123)
m_full <- plsr(log.flux.acetate ~ ., data = model_data, scale = FALSE, validation = "LOO")
summary(m_full)

par(mfrow = c(3,1))
validationplot(m_full)
validationplot(m_full, val.type = "MSEP")
validationplot(m_full, val.type = "R2")

```

```{r}

m1 <- plsr(log.flux.acetate ~ ., data = model_data, 
     scale = FALSE, ncomp = 3) 
summary(m1)
plot(m1)
```

create a plsr and null model to compare and calculate residuals of both
```{r}
model_vals <- data.frame(
  preds = numeric(length = nrow(model_data)), 
  res0 = numeric(length = nrow(model_data)),
  res1 = numeric(length = nrow(model_data))
)


for(i in 1:nrow(model_data)){
  y <- model_data$log.flux.acetate
  
  m <- plsr(log.flux.acetate ~ ., data = model_data[-i,], 
     scale = FALSE, ncomp = 3)
  model_vals$preds[i] <- predict(m, newdata = model_data[i,], 
                                 ncomp = 3, type = "response")
  # predicted residual of the model
  model_vals$res1[i] <-y[i] - model_vals$preds[i]
  
  # predicted residual using the null model
  y_mean = mean(y[-i])
  model_vals$res0[i] = y[i] - y_mean
}

plot(model_data$log.flux.acetate, model_vals$preds)

```
Use F-Test to compare to null model

This is done analogues to comparing a regression model to a null model, 
i.e. by comparing the Residual some of squares of model and null model, calculating the F value using the respective degrees of freedom of the two models and getting the two-sided p-value via the cumulative distribution function of the F distribution. 

```{r}
RSS1 = sum((model_vals$res1)^2)
RSS0 = sum((model_vals$res0)^2)

df1 = nrow(model_data) - (ncol(model_data) - 1) - 1
df0 = nrow(model_data) - 1

F_value <- ((RSS0 - RSS1) / (df0 - df1)) / (RSS1 / df1)
c(F = F_value, df0 = df0, df1 = df1, RSS0 = RSS0, RSS1 = RSS1)

p.value <- 2 * pf(F_value, df0, df1, lower.tail = FALSE)
p.value

```


#Plsr on acetate with 100 iterations to calculate average q2 with SM
```{r plsr-acetate-100-iterations}
#create empty data frame 
cross.val <- data.frame(matrix(ncol = 5, nrow = 1))
colnames(cross.val) <- c("CV.PRESS", "CV.RSS", "CV.Q2", "CV.LimQ2", "CV.Q2cum")



#create for loop to repeat plsr 100 times and output r2 and q2 to dataframe
for (n in 1:54) {
   if (n %% 3 == 1 | n == 1) {
      pls.reg1.acetate <- plsreg1(sub_acetate_pls2_SM[,res.col.acetate], sub_acetate_pls2_SM[,12], comps=3, crosval = TRUE)
      cross.val[n:(n+2),1:5] <- pls.reg1.acetate$Q2
   }
  }

cross.val

q2.scores.v <- as.vector(cross.val$CV.Q2)
q2.scores <- matrix(q2.scores.v, nrow = 18, byrow = TRUE)
q2.scores.df <- as.data.frame(q2.scores)
colnames(q2.scores.df) <- c("q2.1" , "q2.2", "q2.3")


q2.scores

q2.1 <- mean(as.numeric(q2.scores.df$q2.1))
q2.2 <- mean(as.numeric(q2.scores.df$q2.2))
q2.3 <- mean(as.numeric(q2.scores.df$q2.3))

#show averaged q2 values for 1st and 2nd component
q2.1
q2.2
q2.3

#ifelse(as.integer(rownames(q2.scores)) %% 2, )
```

#attempt anova on residuals of 18 models, each with one left out
```{r plsr-acetate-100-iterations}
#define the subset of columns for plsr that contribute most to the model
#res.col.acetate <- c(3,4,5,10,11,12)
#res.col.acetate <- c(1,2,3,4,5,6,7,8,10,11,12)

#subset model data to just these columns
#sub_acetate_pls2_for_anova <- sub_acetate_pls2_SM[,res.col.acetate]

sub_acetate_pls2_for_anova <- as.data.frame(sub_acetate_pls2_SM)

#calculate mean y-value
mean.y <- mean(sub_acetate_pls2_for_anova$flux.acetate)

#create empty data frames for residuals of regression and residuals of means
cross.val.res.reg <- data.frame(matrix(ncol = (nrow(sub_acetate_pls2_for_anova)-1), nrow = nrow(sub_acetate_pls2_for_anova)))

cross.val.res.mean <- data.frame(matrix(ncol = (nrow(sub_acetate_pls2_for_anova)-1), nrow = nrow(sub_acetate_pls2_for_anova)))

#create for loop to repeat plsr 18 times, leaving one out for each iteration. and output results to data frames
for (n in 1:nrow(sub_acetate_pls2_for_anova)) {
      plsr_minus_one <- sub_acetate_pls2_for_anova[-n,]
      pls.reg1.acetate <- plsreg1(plsr_minus_one[,1:(ncol(sub_acetate_pls2_for_anova)-1)], plsr_minus_one[,ncol(sub_acetate_pls2_for_anova)], comps=3, crosval = FALSE)
      cross.val.res.reg[n,] <- pls.reg1.acetate$resid
      cross.val.res.mean[n,] <- pls.reg1.acetate$y.pred - mean.y
  }

    


cross.val.res.reg
cross.val.res.mean


cross.val.df.num <- as.numeric(unlist(cross.val.res.reg))
cross.val.df.num

cross.val.mean.num <- as.numeric(unlist(cross.val.res.mean))
cross.val.mean.num


cross.val.for.anova <- matrix(cross.val.df.num, nrow = 18, byrow = FALSE)
cross.val.mean.for.anova <- matrix(cross.val.mean.num, nrow = 18, byrow = FALSE)


cross.val.for.anova.t <- as.data.frame(t(cross.val.for.anova))
cross.val.mean.for.anova.t <- as.data.frame(t(cross.val.mean.for.anova))
#cross.val.qcum.for.anova.t <- as.data.frame(t(cross.val.qcum.for.anova))

colnames(cross.val.for.anova.t) <- c("it.1", "it.2", "it.3", "it.4", "it.5", "it.6", "it.7", "it.8", "it.9", "it.10",
                         "it.11", "it.12", "it.13", "it.14", "it.15", "it.16", "it.17", "it.18")

colnames(cross.val.mean.for.anova.t) <- c("it.1", "it.2", "it.3", "it.4", "it.5", "it.6", "it.7", "it.8", "it.9", "it.10",
                         "it.11", "it.12", "it.13", "it.14", "it.15", "it.16", "it.17", "it.18")


cross.val.for.anova.t.long <- cross.val.for.anova.t %>%
  gather(key = "iteration", value = "residuals.reg")

cross.val.mean.for.anova.t.long <- cross.val.mean.for.anova.t %>%
  gather(key = "iteration", value = "residuals.mean")

cross.val.combined <- cross.val.for.anova.t.long
cross.val.combined$residuals.mean <- cross.val.mean.for.anova.t.long$residuals.mean

cross.val.summary <- cross.val.combined %>%
  mutate(residuals.mean.sq = residuals.mean^2) %>%
  mutate(residuals.reg.sq = residuals.reg^2) %>%
  group_by(iteration) %>%
  summarise(across(everything(), list(sum))) %>%
  select(-residuals.reg_1, -residuals.mean_1) %>%
  mutate(residuals.sum = residuals.mean.sq_1 + residuals.reg.sq_1) 


cross.val.combined.short <- cross.val.combined %>%
  gather(key = "residuals.type", value = "residual.value", -iteration)

cross.val.summary <- cross.val.combined.short %>%
  group_by(iteration, residuals.type) %>%
  summarize(sum.sq = sum(residual.value^2))


#ifelse(as.integer(rownames(q2.scores)) %% 2, )
```

```{r calculate p-value}
RSS1 = sum((cross.val.combined$residuals.reg)^2)
RSS0 = sum((cross.val.combined$residuals.mean)^2)

df1 = nrow(sub_acetate_pls2_for_anova) - (ncol(sub_acetate_pls2_for_anova) - 1) + 1
df0 = nrow(sub_acetate_pls2_for_anova) - 1

F_value <- ((RSS0 - RSS1) / (df0 - df1)) / (RSS1 / df1)
c(F = F_value, df0 = df0, df1 = df1, RSS0 = RSS0, RSS1 = RSS1)

p.value <- 2 * pf(F_value, df0, df1, lower.tail = FALSE)
p.value

```
```{r Johnny's method}
model_vals <- data.frame(
  preds = numeric(length = nrow(sub_acetate_pls2_SM)), 
  res0 = numeric(length = nrow(sub_acetate_pls2_SM)),
  res1 = numeric(length = nrow(sub_acetate_pls2_SM))
)

exp_var <- as.matrix(sub_acetate_pls2_SM[,1:11])
res_var <- as.matrix(sub_acetate_pls2_SM[,12])
sub_acetate_pls2_SM.t <- as.data.frame(t(sub_acetate_pls2_SM))

for(i in 1:nrow(sub_acetate_pls2_SM)){
  y <- res_var
  m <- plsr(res_var[-i,] ~ exp_var[-i,], scale = FALSE, ncomp = 3)
  model_vals$preds[i] <- predict(m, newdata = sub_acetate_pls2_SM.t[,i], 
                                 ncomp = 3, type = "response")
  # predicted residual of the model
  model_vals$res1[i] <-y[i] - model_vals$preds[i]
  
  # predicted residual using the null model
  y_mean = mean(y[-i])
  model_vals$res0[i] = y[i] - y_mean
}

plot(sub_acetate_pls2_SM$flux.acetate, model_vals$preds)

```

```{r}
RSS1 = sum((model_vals$res1)^2)
RSS0 = sum((model_vals$res0)^2)

df1 = nrow(sub_acetate_pls2_SM) - ncol(sub_acetate_pls2_SM)
df0 = nrow(sub_acetate_pls2_SM) - 1

F_value <- ((RSS0 - RSS1) / (df0 - df1)) / (RSS1 / df1)
c(F = F_value, df0 = df0, df1 = df1, RSS0 = RSS0, RSS1 = RSS1)

p.value <- 2 * pf(F_value, df0, df1, lower.tail = FALSE)
p.value

```
### filter to acetonee cycling genes with acetate flux with soil moisture
```{r}
KO_acetone <- KO_interest_subset %>%
  filter(VOC == "Acetone") 

KO_acetone_list <- KO_acetone$KO
flux_list.log <- c("log.flux.acetate", "log.flux.acetone", "log.SM", "Condition", "Site")
flux_list <- c("flux.acetate", "flux.acetone", "log.SM", "Condition", "Site")

idx <- match(c(KO_acetone_list,flux_list), names(merged_subet_SM))
idx <- idx[!is.na(idx)]

sub_acetone_pls <- merged_subet_SM[,idx]

sub_acetone_pls2_SM <- sub_acetone_pls %>%
  #mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)

sub_acetone_pls2_SM <- sub_acetone_pls2_SM %>%
  #mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
   mutate(flux.acetone2 = flux.acetone) %>%
  select(-flux.acetone) %>%
  rename(flux.acetone = flux.acetone2)

sub_acetone_pls2_SM
```

### acetone w/ acetone genes and acetate flux with SM
```{r}
pls.reg1.acetone <- plsreg1(sub_acetone_pls2_SM[,1:9], sub_acetone_pls2_SM[,10], comps=3, crosval = TRUE)

pls.reg1.acetone$x.scores
pls.reg1.acetone$x.loads
pls.reg1.acetone$y.scores
pls.reg1.acetone$y.loads
pls.reg1.acetone$cor.xyt
pls.reg1.acetone$R2Xy
pls.reg1.acetone$raw.wgs
pls.reg1.acetone$mod.wgs
pls.reg1.acetone$y.pred
pls.reg1.acetone$std.coefs
pls.reg1.acetone$reg.coefs
pls.reg1.acetone$residn
pls.reg1.acetone$R2
pls.reg1.acetone$Q2
pls.reg1.acetone$VIP

plot(pls.reg1.acetone)
```

### acetone w/ acetone genes and acetate flux with SM - take out low drivers
```{r}
res.col.acetone <- c(1,2,3,4,5,6,7,8,10)
pls.reg1.acetone <- plsreg1(sub_acetone_pls2_SM[,res.col.acetone], sub_acetone_pls2_SM[,9], comps=2, crosval = TRUE)

pls.reg1.acetone$x.scores
pls.reg1.acetone$x.loads
pls.reg1.acetone$y.scores
pls.reg1.acetone$y.loads
pls.reg1.acetone$cor.xyt
pls.reg1.acetone$R2Xy
pls.reg1.acetone$raw.wgs
pls.reg1.acetone$mod.wgs
pls.reg1.acetone$y.pred
pls.reg1.acetone$std.coefs
pls.reg1.acetone$reg.coefs
pls.reg1.acetone$residn
pls.reg1.acetone$R2
pls.reg1.acetone$Q2
pls.reg1.acetone$VIP

plot(pls.reg1.acetone)
```

#Plsr on acetone with 1000 iterations to calculate  q2 with SM
```{r plsr-acetate-100-iterations}
#create empty data frame 
cross.val <- data.frame(matrix(ncol = 5, nrow = 1))
colnames(cross.val) <- c("CV.PRESS", "CV.RSS", "CV.Q2", "CV.LimQ2", "CV.Q2cum")



#create for loop to repeat plsr 100 times and output r2 and q2 to dataframe
for (n in 1:54) {
   if (n %% 3 == 1 | n == 1) {
      pls.reg1.acetone <- plsreg1(sub_acetone_pls2_SM[,1:9], sub_acetone_pls2_SM[,10], comps=3, crosval = TRUE)
      cross.val[n:(n+2),1:5] <- pls.reg1.acetone$Q2
   }
  }

cross.val

q2.scores.v <- as.vector(cross.val$CV.Q2)
q2.scores <- matrix(q2.scores.v, nrow = 18, byrow = TRUE)
q2.scores.df <- as.data.frame(q2.scores)
colnames(q2.scores.df) <- c("q2.1" , "q2.2", "q2.3")


q2.scores

q2.1 <- mean(as.numeric(q2.scores.df$q2.1))
q2.2 <- mean(as.numeric(q2.scores.df$q2.2))
q2.3 <- mean(as.numeric(q2.scores.df$q2.3))

#show averaged q2 values for 1st and 2nd component
q2.1
q2.2
q2.3

#ifelse(as.integer(rownames(q2.scores)) %% 2, )
```

#attempt anova on residuals of 18 models, each with one left out
```{r plsr-acetate-100-iterations}

#define the subset of columns for plsr that contribute most to the model
#res.col.acetate <- c(3,4,5,10,11,12)
#res.col.acetate <- c(1,2,3,4,5,6,7,8,10,11,12)

#subset model data to just these columns
#sub_acetate_pls2_for_anova <- sub_acetate_pls2_SM[,res.col.acetate]

sub_acetone_pls2_for_anova <- as.data.frame(sub_acetone_pls2_SM)

#calculate mean y-value
mean.y <- mean(sub_acetone_pls2_for_anova$flux.acetate)

#create empty data frames for residuals of regression and residuals of means
cross.val.res.reg <- data.frame(matrix(ncol = (nrow(sub_acetone_pls2_for_anova)-1), nrow = nrow(sub_acetone_pls2_for_anova)))

cross.val.res.mean <- data.frame(matrix(ncol = (nrow(sub_acetone_pls2_for_anova)-1), nrow = nrow(sub_acetone_pls2_for_anova)))

#create for loop to repeat plsr 18 times, leaving one out for each iteration. and output results to data frames
for (n in 1:nrow(sub_acetone_pls2_for_anova)) {
      plsr_minus_one <- sub_acetone_pls2_for_anova[-n,]
      mean.y <- mean(plsr_minus_one$flux.acetate)
      pls.reg1.acetone <- plsreg1(plsr_minus_one[,1:(ncol(sub_acetone_pls2_for_anova)-1)], plsr_minus_one[,ncol(sub_acetone_pls2_for_anova)], comps=3, crosval = FALSE)
      cross.val.res.reg[n,] <- pls.reg1.acetone$resid
      cross.val.res.mean[n,] <- pls.reg1.acetone$y.pred - mean.y
  }

    


cross.val.res.reg
cross.val.res.mean


cross.val.df.num <- as.numeric(unlist(cross.val.res.reg))
cross.val.df.num

cross.val.mean.num <- as.numeric(unlist(cross.val.res.mean))
cross.val.mean.num


cross.val.for.anova <- matrix(cross.val.df.num, nrow = 18, byrow = FALSE)
cross.val.mean.for.anova <- matrix(cross.val.mean.num, nrow = 18, byrow = FALSE)


cross.val.for.anova.t <- as.data.frame(t(cross.val.for.anova))
cross.val.mean.for.anova.t <- as.data.frame(t(cross.val.mean.for.anova))
#cross.val.qcum.for.anova.t <- as.data.frame(t(cross.val.qcum.for.anova))

colnames(cross.val.for.anova.t) <- c("it.1", "it.2", "it.3", "it.4", "it.5", "it.6", "it.7", "it.8", "it.9", "it.10",
                         "it.11", "it.12", "it.13", "it.14", "it.15", "it.16", "it.17", "it.18")

colnames(cross.val.mean.for.anova.t) <- c("it.1", "it.2", "it.3", "it.4", "it.5", "it.6", "it.7", "it.8", "it.9", "it.10",
                         "it.11", "it.12", "it.13", "it.14", "it.15", "it.16", "it.17", "it.18")


cross.val.for.anova.t.long <- cross.val.for.anova.t %>%
  gather(key = "iteration", value = "residuals.reg")

cross.val.mean.for.anova.t.long <- cross.val.mean.for.anova.t %>%
  gather(key = "iteration", value = "residuals.mean")

cross.val.combined <- cross.val.for.anova.t.long
cross.val.combined$residuals.mean <- cross.val.mean.for.anova.t.long$residuals.mean

cross.val.summary <- cross.val.combined %>%
  mutate(residuals.mean.sq = residuals.mean^2) %>%
  mutate(residuals.reg.sq = residuals.reg^2) %>%
  group_by(iteration) %>%
  summarise(across(everything(), list(sum))) %>%
  select(-residuals.reg_1, -residuals.mean_1) %>%
  mutate(residuals.sum = residuals.mean.sq_1 + residuals.reg.sq_1) 


cross.val.combined.short <- cross.val.combined %>%
  gather(key = "residuals.type", value = "residual.value", -iteration)

cross.val.summary <- cross.val.combined.short %>%
  group_by(iteration, residuals.type) %>%
  summarize(sum.sq = sum(residual.value^2))


#ifelse(as.integer(rownames(q2.scores)) %% 2, )
```
```{r Johnny's method}
model_vals <- data.frame(
  preds = numeric(length = nrow(sub_acetone_pls2_SM)), 
  res0 = numeric(length = nrow(sub_acetone_pls2_SM)),
  res1 = numeric(length = nrow(sub_acetone_pls2_SM))
)

exp_var <- sub_acetone_pls2_SM[,-10]
flux.acetone <- sub_acetone_pls2_SM[,10]

for(i in 1:nrow(sub_acetone_pls2_SM)){
  y <- sub_acetone_pls2_SM$flux.acetone
  
  m <- plsr(flux.acetone[-i,] ~ ., data = exp_var[-i,], 
     scale = FALSE, ncomp = 3)
  model_vals$preds[i] <- predict(m, newdata = exp_var[i,], 
                                 ncomp = 3, type = "response")
  # predicted residual of the model
  model_vals$res1[i] <-y[i] - model_vals$preds[i]
  
  # predicted residual using the null model
  y_mean = mean(y[-i])
  model_vals$res0[i] = y[i] - y_mean
}

plot(sub_acetone_pls2_SM$flux.acetate, model_vals$preds)

```

```{r calculate p-value}
RSS1 = sum((model_vals$res1)^2)
RSS0 = sum((model_vals$res0)^2)

df1 = nrow(sub_acetone_pls2_for_anova) - (ncol(sub_acetone_pls2_for_anova) - 1) + 1
df0 = nrow(sub_acetone_pls2_for_anova) - 1

F_value <- ((RSS0 - RSS1) / (df0 - df1)) / (RSS1 / df1)
c(F = F_value, df0 = df0, df1 = df1, RSS0 = RSS0, RSS1 = RSS1)

p.value <- 2 * pf(F_value, df0, df1, lower.tail = FALSE)
p.value

```
#attempt anova on residuals of 18 models, each with one left out - with q2 values
```{r plsr-acetate-100-iterations}

#move acetate column to the end
sub_acetate_pls2_for_anova <- sub_acetate_pls2_SM %>%
  mutate(log.flux.acetate2 = log.flux.acetate) %>%
  select(-log.flux.acetate, -tmt) 

sub_acetate_pls2_for_anova <- as.data.frame(sub_acetate_pls2_for_anova)

#calculate mean y-value
mean.y <- mean(sub_acetate_pls2_for_anova$log.flux.acetate)

#create empty data frames for residuals of regression and residuals of means
cross.val.res.reg <- data.frame(matrix(ncol = 17, nrow = 18))

cross.val.res.mean <- data.frame(matrix(ncol = 17, nrow = 18))
cross.val <- data.frame(matrix(ncol = 5, nrow = 1))
colnames(cross.val) <- c("CV.PRESS", "CV.RSS", "CV.Q2", "CV.LimQ2", "CV.Q2cum")


#set first iteration to 1
n = 1

#create for loop to repeat plsr 18 times, leaving one out for each iteration. and output results to data frames
for (n in 1:18) {
  if(n == 1) {
      plsr_minus_one <- sub_acetate_pls2_for_anova[-n,]
      pls.reg1.acetate <- plsreg1(plsr_minus_one[,1:11], plsr_minus_one[,12], comps=2, crosval = TRUE)
      q2.line <- n
      cross.val[q2.line:(q2.line+1),1:5] <- pls.reg1.acetate$Q2
      cross.val.res.reg[n,] <- pls.reg1.acetate$resid
      cross.val.res.mean[n,] <- pls.reg1.acetate$y.pred - mean.y
      #n = n + 1 
      }
   plsr_minus_one <- sub_acetate_pls2_for_anova[-n,]
   pls.reg1.acetate <- plsreg1(plsr_minus_one[,1:11], plsr_minus_one[,12], comps=2, crosval = TRUE)
   q2.line <- n + (n-1)
   cross.val[q2.line:(q2.line+1),1:5] <- pls.reg1.acetate$Q2
   cross.val.res.reg[n,] <- pls.reg1.acetate$resid
   cross.val.res.mean[n,] <- pls.reg1.acetate$y.pred - mean.y
  }

cross.val.res.reg
cross.val.res.mean
cross.val

toDelete <- seq(2, nrow(cross.val), 2)
cross.val.qcum <- cross.val[-toDelete,]



cross.val.df.num <- as.numeric(unlist(cross.val.res.reg))
cross.val.df.num

cross.val.mean.num <- as.numeric(unlist(cross.val.res.mean))
cross.val.mean.num

cross.val.qcum <- as.numeric(unlist(cross.val.qcum))
cross.val.qcum

cross.val.for.anova <- matrix(cross.val.df.num, nrow = 18, byrow = FALSE)
cross.val.mean.for.anova <- matrix(cross.val.mean.num, nrow = 18, byrow = FALSE)
cross.val.qcum.for.anova <- matrix(cross.val.qcum, nrow = 18, byrow = FALSE)

colnames(cross.val.qcum.for.anova) <- c("PRESS", "RSS", "Q2", "LimQ2", "Q2cum")
rownames(cross.val.qcum.for.anova) <- c("it.1", "it.2", "it.3", "it.4", "it.5", "it.6", "it.7", "it.8", "it.9", "it.10",
                         "it.11", "it.12", "it.13", "it.14", "it.15", "it.16", "it.17", "it.18")

cross.val.qcum.for.anova <- as.data.frame(cross.val.qcum.for.anova)

cross.val.qcum.for.anova.arr <- cross.val.qcum.for.anova %>%
  rownames_to_column(var = "iteration") %>%
  arrange(iteration) %>%
  column_to_rownames(var = "iteration")

cross.val.for.anova.t <- as.data.frame(t(cross.val.for.anova))
cross.val.mean.for.anova.t <- as.data.frame(t(cross.val.mean.for.anova))
#cross.val.qcum.for.anova.t <- as.data.frame(t(cross.val.qcum.for.anova))

colnames(cross.val.for.anova.t) <- c("it.1", "it.2", "it.3", "it.4", "it.5", "it.6", "it.7", "it.8", "it.9", "it.10",
                         "it.11", "it.12", "it.13", "it.14", "it.15", "it.16", "it.17", "it.18")

colnames(cross.val.mean.for.anova.t) <- c("it.1", "it.2", "it.3", "it.4", "it.5", "it.6", "it.7", "it.8", "it.9", "it.10",
                         "it.11", "it.12", "it.13", "it.14", "it.15", "it.16", "it.17", "it.18")


cross.val.for.anova.t.long <- cross.val.for.anova.t %>%
  gather(key = "iteration", value = "residuals.reg")

cross.val.mean.for.anova.t.long <- cross.val.mean.for.anova.t %>%
  gather(key = "iteration", value = "residuals.mean")

cross.val.combined <- cross.val.for.anova.t.long
cross.val.combined$residuals.mean <- cross.val.mean.for.anova.t.long$residuals.mean

cross.val.summary <- cross.val.combined %>%
  mutate(residuals.mean.sq = residuals.mean^2) %>%
  mutate(residuals.reg.sq = residuals.reg^2) %>%
  group_by(iteration) %>%
  summarise(across(everything(), list(sum))) %>%
  select(-residuals.reg_1, -residuals.mean_1) %>%
  mutate(residuals.sum = residuals.mean.sq_1 + residuals.reg.sq_1) 

cross.val.summary$Q2 <- cross.val.qcum.for.anova.arr$Q2cum

cross.val.summary.total <- cross.val.summary %>%
  mutate(q2.ssm = residuals.mean.sq_1 * Q2) %>%
  select(-iteration) %>%
  summarise(across(everything(), list(sum)))


cross.val.combined.short <- cross.val.combined %>%
  gather(key = "residuals.type", value = "residual.value", -iteration)

cross.val.summary <- cross.val.combined.short %>%
  group_by(iteration, residuals.type) %>%
  summarize(sum.sq = sum(residual.value^2))



res_aov <- aov(sum.sq ~ iteration * residuals.type + iteration + residuals.type,
  data = cross.val.summary)
summary(res_aov)

ggplot(res_aov) +
  aes(x = iteration, y = sum.sq) +
  geom_boxplot() +
  facet_wrap(~residuals.type)
#ifelse(as.integer(rownames(q2.scores)) %% 2, )
```