---
title: "PLS"
author: "Linnea Honeker"
date: "11/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(phyloseq)
library(vegan)
library(ggplot2)
library(rms)
library(reshape2)
library(coefplot)
library(MASS)
library(dplyr)
library(lme4)
library(MASS)
library(leaps)
library(car)
library(relaimpo)
library(nlme)
library(lmerTest)
library(plsdepot)
library(pls)
library(caret)
library(DESeq2)

```

# load metaT data and do variance stabilization tranformation
```{r load-metaT, echo = FALSE}
#load metaT gene copy data and column metadata
datExpr0 <-read.csv("../../Data/Deseq2/Drought_vs_predrought_metaT_missing_1_no_feature.csv", header = TRUE)
rownames(datExpr0) <- datExpr0$Feature
datExpr0$Feature <- NULL

coldata <- read.csv("../../Data/Deseq2/metaT-metadata.csv", header = TRUE)
rownames(coldata) <- coldata$SampleID
coldata$SampleID <- NULL
row.names(coldata) [7] <- "X3300045460_Site1_Drought_48hr"

#filter features that are not in at least 5 samples
datExpr.f <- datExpr0 %>%
  mutate(count = rowSums(. > 0)) %>%
  filter(count > 5) %>%
  subset(select = -count )

#normalize filtered gene copy data using deseq2, variance stabilization transformation
dds <- DESeqDataSetFromMatrix(countData = datExpr.f,
                              colData = coldata,
                              design= ~ Condition_Time)

#vst transformation#
vstcounts <- vst(dds, blind=TRUE)
vst.export<- assay(vstcounts)
metaT<- as.data.frame(vst.export)
metaT$X <- rownames(metaT)
```
```{r all-KOs, echo = FALSE}
KO_interest <- read.csv("../../Data/Deseq2/VOC-KO-interest-from-ML.csv")
#metaT <- read.csv("../../Output/WGCNA/vst-metaT.csv")
#metaT <- read.csv("../../Output/WGCNA/vst-metaT.csv")
metadata <- read.csv("../../Data/Deseq2/metaT-metadata.csv")
acetate_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/acetic-acid-for-ggplot.csv")
acetone_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/acetone-for-ggplot.csv")
diacetyl_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/diacetyl-for-ggplot.csv")


#remove "KO:"
metaT$X <- sub("KO:", "", metaT$X)
diacetyl_flux$Sample <- sub("_C1", "", diacetyl_flux$Sample)
diacetyl_flux$Sample <- sub("_C2", "", diacetyl_flux$Sample)

#merge  all VOC
KO_VOC <- KO_interest %>%
  merge(., metaT, by.x = "KO", by.y="X", all.x = FALSE, all.y = FALSE)

#transpose for glm
KO_VOC.t <- t(KO_VOC)

#further format KO_acetate.t for glm. Assign column headers and remove rows containing metadata
colnames(KO_VOC.t) <- KO_VOC.t[1,]
KO_VOC.t.f <- KO_VOC.t[-(1:3),]
KO_VOC.t.f <- as.data.frame(KO_VOC.t.f)

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetate_flux_proc <- acetate_flux %>%
    filter (Pyruv == "C2") %>%
     mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetate = Flux) %>%
  dplyr::select(-c("Pyruv", "Flux", "Time", "Time.hr"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetone_flux_proc <- acetone_flux %>%
    filter (Pyruv == "C2") %>%
     mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetone = Flux) %>%
  dplyr::select(-c("Pyruv", "Flux", "Time", "Time.hr"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
diacetyl_flux_proc <- diacetyl_flux %>%
    filter (Pyruv == "C2") %>%
     mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.diacetyl = Flux) %>%
  dplyr::select(-c("Pyruv", "Flux", "Time", "Time.hr"))

#sample names are different in the flux and expression tables, so need to make them the same
KO_VOC.t.f[,40] <- rownames(KO_VOC.t.f)
KO_VOC.t.f_new_id <- KO_VOC.t.f %>%
  mutate(Sample = case_when(
  startsWith(V40, "X3300044942") ~ "S3P4",
  startsWith(V40, "X3300044943") ~ "S3P4",
  startsWith(V40, "X3300044944") ~ "S3P4",
  startsWith(V40, "X3300044945") ~ "S3P4",
  startsWith(V40, "X3300045458") ~ "S1P5",
  startsWith(V40, "X3300045459") ~ "S1P5",
  startsWith(V40, "X3300045460") ~ "S1P5",
  startsWith(V40, "X3300045461") ~ "S2P4",
  startsWith(V40, "X3300045462") ~ "S2P4",
  startsWith(V40, "X3300045463") ~ "S2P4",
  startsWith(V40, "X3300045464") ~ "S2P4",
  startsWith(V40, "X3300045465") ~ "S2P4",
  startsWith(V40, "X3300045466") ~ "S2P6",
  startsWith(V40, "X3300045467") ~ "S2P6",
  startsWith(V40, "X3300045468") ~ "S2P6",
  startsWith(V40, "X3300045469") ~ "S2P6",
  startsWith(V40, "X3300045470") ~ "S2P6",
  startsWith(V40, "X3300045471") ~ "S3P5",
  startsWith(V40, "X3300045472") ~ "S3P5",
  startsWith(V40, "X3300045473") ~ "S2P6",
  startsWith(V40, "X3300048051") ~ "S1P5",
  startsWith(V40, "X3300048052") ~ "S3P5",
  startsWith(V40, "X3300048053") ~ "S3P5",
  startsWith(V40, "X3300048054") ~ "S3P5",
  startsWith(V40, "X3300048092") ~ "S1P1",
  startsWith(V40, "X3300048093") ~ "S1P1",
  startsWith(V40, "X3300048094") ~ "S1P1",
  startsWith(V40, "X3300048095") ~ "S1P1",
  startsWith(V40, "X3300048096") ~ "S1P1",
  startsWith(V40, "X3300048099") ~ "S1P5",
  startsWith(V40, "X3300048100") ~ "S1P5",
  startsWith(V40, "X3300049511") ~ "S2P4",
  startsWith(V40, "X3300049512") ~ "S3P4",
  startsWith(V40, "X3300049557") ~ "S3P4",
  startsWith(V40, "X3300049628") ~ "S3P5"
  )) %>%
  mutate(Time.d = case_when(
    endsWith(V40, "6hr") ~ "6",
    endsWith(V40, "0hr") ~ "0",
    endsWith(V40, "48hr") ~ "48",
    endsWith(V40, "48_hr") ~ "48"
  )) %>%
  mutate(Condition = case_when(
    grepl("PreDrought", V40) ~ "pre-drought",
    grepl("Drought", V40) ~ "drought",
    grepl("Pre_drought", V40) ~ "pre-drought"
  )) %>%
   dplyr::select(-V40)

#merge VOC and KO tables
merged <- KO_VOC.t.f_new_id %>%
  merge(., acectate_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample","Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate_at(c(4:42), as.numeric) %>%
  mutate(flux.acetate = as.numeric(flux.acetate)) %>%
  mutate(log.flux.acetate = log(flux.acetate+abs(min(flux.acetate))+1, base = exp(2))) %>%
  merge(.,acetone_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.acetone = as.numeric(flux.acetone)) %>%
  mutate(log.flux.acetone = log(flux.acetone+abs(min(flux.acetone))+1, base = exp(2))) %>%
  merge(.,diacetyl_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.diacetyl = as.numeric(flux.diacetyl)) %>%
  mutate(log.flux.diacetyl = log(flux.diacetyl+abs(min(flux.diacetyl))+1, base = exp(2)))  %>%
  select(c(-Type.x, -Type.y, -Site.x, -Site.y))
  
merged
```
## Subset of KOs originally included in the first manuscript submission

```{r KOs-subset, echo=FALSE}
KO_interest_subset <- read.csv("../../Data/Deseq2/KO_interest.csv")
#metaT <- read.csv("../../Output/WGCNA/vst-metaT.csv")
metadata <- read.csv("../../Data/Deseq2/metaT-metadata.csv")
acetate_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/acetic-acid-for-ggplot.csv")
acetone_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/acetone-for-ggplot.csv")
diacetyl_flux <- read.csv("../../Data/CO2-VOCs/voc_raw/diacetyl-for-ggplot.csv")

#remove "KO:"
metaT$X <- sub("KO:", "", metaT$X)
diacetyl_flux$Sample <- sub("_C1", "", diacetyl_flux$Sample)
diacetyl_flux$Sample <- sub("_C2", "", diacetyl_flux$Sample)

#merge  all VOC
KO_VOC <- KO_interest_subset %>%
  merge(., metaT, by.x = "KO", by.y="X", all.x = FALSE, all.y = FALSE)

#transpose for glm
KO_VOC.t <- t(KO_VOC)

#further format KO_acetate.t for glm. Assign column headers and remove rows containing metadata
colnames(KO_VOC.t) <- KO_VOC.t[1,]
KO_VOC.t.f <- KO_VOC.t[-(1:6),]
KO_VOC.t.f <- as.data.frame(KO_VOC.t.f)

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acectate_flux_proc <- acetate_flux %>%
  filter (Type == "sample") %>%
    filter (Pyruv == "C2") %>%
     mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetate = Flux) %>%
  dplyr::select(-c("Pyruv", "Type", "Flux", "Time", "Time.hr"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
acetone_flux_proc <- acetone_flux %>%
  filter (Type == "sample") %>%
  filter (Pyruv == "C2") %>%
  mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.acetone = Flux)%>%
   dplyr::select(-c("Pyruv", "Type", "Flux", "Time", "Time.hr", "Site"))

#process the voc flux data so that it is binned based on time to just 0 + 2 hours, 6 + or - 1 hour, and 48 - 2 hours.
diacetyl_flux_proc <- diacetyl_flux %>%
  filter (Type == "sample") %>%
  filter (Pyruv == "C2") %>%
  mutate(Site = case_when(
    startsWith(Sample, "S1") ~ "Site1",
    startsWith(Sample, "S2") ~ "Site2",
    startsWith(Sample, "S3") ~ "Site3"
  )) %>%
  mutate(Time.hr = Time * 24) %>%
  mutate (Time.d = case_when(
    (Time.hr > 0 & Time.hr < 2) ~ "0",
    (Time.hr > 5 & Time.hr < 7) ~ "6",
    (Time.hr > 46 & Time.hr < 48) ~ "48")) %>%
  group_by(Condition, Site, Sample, Time.d) %>%
  summarise_all(funs(mean)) %>%
  filter(!is.na(Time.d)) %>%
  mutate(flux.diacetyl = Flux) %>%
   dplyr::select(-c("Pyruv", "Type", "Flux", "Time", "Time.hr", "Site"))

#sample names are different in the flux and expression tables, so need to make them the same
KO_VOC.t.f[,28] <- rownames(KO_VOC.t.f)
KO_VOC.t.f_new_id <- KO_VOC.t.f %>%
  mutate(Sample = case_when(
  startsWith(V28, "X3300044942") ~ "S3P4",
  startsWith(V28, "X3300044943") ~ "S3P4",
  startsWith(V28, "X3300044944") ~ "S3P4",
  startsWith(V28, "X3300044945") ~ "S3P4",
  startsWith(V28, "X3300045458") ~ "S1P5",
  startsWith(V28, "X3300045459") ~ "S1P5",
  startsWith(V28, "X3300045460") ~ "S1P5",
  startsWith(V28, "X3300045461") ~ "S2P4",
  startsWith(V28, "X3300045462") ~ "S2P4",
  startsWith(V28, "X3300045463") ~ "S2P4",
  startsWith(V28, "X3300045464") ~ "S2P4",
  startsWith(V28, "X3300045465") ~ "S2P4",
  startsWith(V28, "X3300045466") ~ "S2P6",
  startsWith(V28, "X3300045467") ~ "S2P6",
  startsWith(V28, "X3300045468") ~ "S2P6",
  startsWith(V28, "X3300045469") ~ "S2P6",
  startsWith(V28, "X3300045470") ~ "S2P6",
  startsWith(V28, "X3300045471") ~ "S3P5",
  startsWith(V28, "X3300045472") ~ "S3P5",
  startsWith(V28, "X3300045473") ~ "S2P6",
  startsWith(V28, "X3300048051") ~ "S1P5",
  startsWith(V28, "X3300048052") ~ "S3P5",
  startsWith(V28, "X3300048053") ~ "S3P5",
  startsWith(V28, "X3300048054") ~ "S3P5",
  startsWith(V28, "X3300048092") ~ "S1P1",
  startsWith(V28, "X3300048093") ~ "S1P1",
  startsWith(V28, "X3300048094") ~ "S1P1",
  startsWith(V28, "X3300048095") ~ "S1P1",
  startsWith(V28, "X3300048096") ~ "S1P1",
  startsWith(V28, "X3300048099") ~ "S1P5",
  startsWith(V28, "X3300048100") ~ "S1P5",
  startsWith(V28, "X3300049511") ~ "S2P4",
  startsWith(V28, "X3300049512") ~ "S3P4",
  startsWith(V28, "X3300049557") ~ "S3P4",
  startsWith(V28, "X3300049628") ~ "S3P5"
  )) %>%
  mutate(Time.d = case_when(
    endsWith(V28, "6hr") ~ "6",
    endsWith(V28, "0hr") ~ "0",
    endsWith(V28, "48hr") ~ "48",
    endsWith(V28, "48_hr") ~ "48"
  )) %>%
  mutate(Condition = case_when(
    grepl("PreDrought", V28) ~ "pre-drought",
    grepl("Drought", V28) ~ "drought",
    grepl("Pre_drought", V28) ~ "pre-drought"
  )) %>%
   dplyr::select(-V28)

#merge VOC and KO tables
merged_subset <- KO_VOC.t.f_new_id %>%
  merge(., acectate_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample","Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate_at(c(4:31), as.numeric) %>%
  mutate(flux.acetate = as.numeric(flux.acetate)) %>%
  mutate(log.flux.acetate = log(flux.acetate+abs(min(flux.acetate))+1, base = exp(2))) %>%
  merge(.,acetone_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.acetone = as.numeric(flux.acetone)) %>%
  mutate(log.flux.acetone = log(flux.acetone+abs(min(flux.acetone))+1, base = exp(2))) %>%
  merge(.,diacetyl_flux_proc, by.x = c("Condition", "Sample", "Time.d"), by.y = c("Condition", "Sample", "Time.d"),all.x = FALSE, all.y = FALSE) %>%
  mutate(flux.diacetyl = as.numeric(flux.diacetyl)) %>%
  mutate(log.flux.diacetyl = log(flux.diacetyl+abs(min(flux.diacetyl))+1, base = exp(2))) 
  
merged_subset
```

## Filter to acetate and acetone flux and acetate cycling genes
```{r}
KO_acetate <- KO_interest_subset %>%
  filter(VOC == "Acetate") 

KO_acetate_list <- KO_acetate$KO
flux_list <- c("log.flux.acetate", "log.flux.acetone", "Condition", "Site")

idx <- match(c(KO_acetate_list,flux_list), names(merged_subset))
idx <- idx[!is.na(idx)]

sub_acetate_pls <- merged_subset[,idx]

sub_acetate_pls2 <- sub_acetate_pls %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)

sub_acetate_pls2
```
## Filter to acetone and acetate flux and acetone cycling genes
```{r}
KO_acetone <- KO_interest_subset %>%
  filter(VOC == "Acetone") 

KO_acetone_list <- KO_acetone$KO
flux_list <- c("log.flux.acetate", "log.flux.acetone", "Condition", "Site")

idx <- match(c(KO_acetone_list,flux_list), names(merged_subset))
idx <- idx[!is.na(idx)]

sub_acetone_pls <- merged_subset[,idx]

sub_acetone_pls2 <- sub_acetone_pls %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)

sub_acetone_pls2
```
## Filter to diacetyl flux and diacetyl genes
```{r}
KO_diacetyl <- KO_interest_subset %>%
  filter(VOC == "Diacetyl") 

KO_diacetyl_list <- KO_diacetyl$KO
flux_list <- c("log.flux.diacetyl", "Condition", "Site")

idx <- match(c(KO_diacetyl_list,flux_list), names(merged_subset))
idx <- idx[!is.na(idx)]

sub_diacetyl_pls <- merged_subset[,idx]

sub_diacetyl_pls2 <- sub_diacetyl_pls %>%
  mutate(tmt = ifelse(Condition == "drought", 1, 0)) %>%
  dplyr::select(-Condition, -Site)

sub_diacetyl_pls2
```


### acetate w/ acetate genes and acetone flux
```{r}
res.col <- c(1:9, 11)
pls.reg1 <- plsreg1(sub_acetate_pls2[,res.col], sub_acetate_pls2[,10], comps=2, crosval = TRUE)

pls.reg1$x.scores
pls.reg1$x.loads
pls.reg1$y.scores
pls.reg1$y.loads
pls.reg1$R2
pls.reg1$cor.xyt
pls.reg1$R2Xy
pls.reg1$raw.wgs
pls.reg1$mod.wgs
pls.reg1$y.pred
pls.reg1$std.coefs
pls.reg1$reg.coefs
pls.reg1$residn
pls.reg1$Q2
pls.reg1$VIP

plot(pls.reg1)
```


### acetone w/ acetone genes and acetate flux
```{r}
res.col <- c(1:7, 8)
pls.reg1 <- plsreg1(sub_acetone_pls2[,res.col], sub_acetone_pls2[,9], comps=2, crosval = TRUE)

pls.reg1$x.scores
pls.reg1$x.loads
pls.reg1$y.scores
pls.reg1$y.loads
pls.reg1$R2
pls.reg1$cor.xyt
pls.reg1$R2Xy
pls.reg1$raw.wgs
pls.reg1$mod.wgs
pls.reg1$y.pred
pls.reg1$std.coefs
pls.reg1$reg.coefs
pls.reg1$residn
pls.reg1$Q2

plot(pls.reg1)
```


### diacetyl w/ diacetyl genes
```{r}
res.col <- c(1:3)
pls.reg1 <- plsreg1(sub_diacetyl_pls2[,res.col], sub_diacetyl_pls2[,4], comps=2, crosval = TRUE)

pls.reg1$x.scores
pls.reg1$x.loads
pls.reg1$y.scores
pls.reg1$y.loads
pls.reg1$R2
pls.reg1$cor.xyt
pls.reg1$R2Xy
pls.reg1$raw.wgs
pls.reg1$mod.wgs
pls.reg1$y.pred
pls.reg1$std.coefs
pls.reg1$reg.coefs
pls.reg1$residn
pls.reg1$Q2

plot(pls.reg1)
```


