---
title: "Cumulative flux"
author: "Linnea Honeker"
date: "6/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(plotly)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(flux)
library(dplyr)
library(data.table)
library(janitor)
library(performance)
library(lme4)
library(MASS)
library(leaps)
library(car)
library(relaimpo)
library(nlme)
library(lmerTest)
library(see)

theme_custom <- theme_few()

theme_set(theme_custom)


colors = c("#2ECC71" ,"#F39C12")

```

## Cmulative efflux: CO2

```{r CO2, echo = FALSE}

#load data
data <- read.csv("../../Data/CO2-VOCs/01_raw/S3_PyruvateLabeling_qc.csv") %>% 
  dplyr::select(-X)

#add timestanp, however, this hasn't been working and not needed
data <- data %>% 
  mutate(Timestamp = lubridate::ymd_hms(Timestamp_MST, tz = "America/Phoenix"), 
         Timestamp_Label_MST = lubridate::ymd_hms(Timestamp_Label_MST, tz = "America/Phoenix")) %>% 
  dplyr::select(-Timestamp_MST)

# Quickview
data <- data %>% 
  dplyr::select(Label, Condition, Pyruv, Trel, Flux, D13C, Qc_iso, Qc_flux, AtF)

# quality control filtering, multiply Flux (which is in umol m-2 s-1) by 3600 to get units to umol m-2 h-1)
data_sub <- data %>% 
  filter(Qc_flux == 1, Qc_iso == 1)  %>% 
  dplyr::select(-Qc_iso, -Qc_flux) %>% 
  filter(Trel > -10, Trel < 50, Flux > 0, D13C > -100) %>%
  mutate(Flux.h = Flux * 3600) 

##### calculate average AtF, below are equations from Johnny
# atF = Flux13/(flux13 + Flux12);
# d13C = (((flux13 / flux12)/Rvpdb)-1) * 1000
# F_13CO2 = Flux * AtF
# F_12CO2 = Flux - F_13CO2 = Flux * (1-AtF)
#  Rvpdb = 0.0111802 (from Werner and Brand, 2001)
# solving for AtF, Atf = d13/11.1802

# alculate AtF natural abundance for pre labeling time points
data_sub_pre <- data_sub %>%
  filter(Trel <= 0) 

#calculate mean AtF natural abundance across all pre-labeling time points
AtFna_mean = mean(data_sub_pre$AtF)

# filter to all post-labeling time points
data_sub_post <- data_sub %>%
  filter(Trel > 0)

# Calculate AtF, flux13, flux12 (flux units: umol m-2 s-1)
data_sub_13C_flux <- data_sub_post %>%
  mutate(fCO2_13 = AtF * Flux.h) %>%
  mutate(AtFe = AtF - AtFna_mean) %>% #calculate AtF from 13C pyruvate
  mutate(fCO2_13_py = AtFe * Flux.h) #calculate 13C flux attributed to 13C pyruvate

########Calculate AUC separately for each collar, then average######################
#separate out drought and pre-drought
data_sub_13C_flux_drought <- data_sub_13C_flux %>%
  filter(Condition == "drought") #%>%
  #mutate(Trel.s = Trel * 360) %>%
  #select(-Trel)

data_sub_13C_flux_predrought <- data_sub_13C_flux %>%
  filter(Condition == "pre_drought") #%>%
 # mutate(Trel.s = Trel * 360) %>%
  #select(-Trel)


#Define samples that are either C1 or C2 labeled
Label.C1a = c("S1.P1", "S1.P3", "S1.P5", "S2.P1", "S2.P3", "S2.P7", "S3.P1", "S3.P3", "S3.P5")
Label.C1b = c("S1.P1", "S1.P3", "S1.P5", "S2.P1", "S2.P3", "S2.P5", "S3.P1", "S3.P3", "S3.P5")
Label.C2a = c("S1.P2", "S1.P4", "S1.P6", "S2.P2", "S2.P4", "S2.P6", "S3.P2", "S3.P4", "S3.P6")
Label.C2b = c("S1.P2", "S1.P4", "S1.P6",  "S2.P4", "S2.P6","S2.P8", "S3.P2", "S3.P4", "S3.P6")

#create empty lists for each category
AUC_list_PD_C1 <- list()
AUC_list_PD_C2 <- list()
AUC_list_D_C1 <- list()
AUC_list_D_C2 <- list()

for (i in Label.C1a) {
  x <- data_sub_13C_flux_predrought %>%
  filter(Label == i)
  i.AUC <- auc(x$Trel, x$fCO2_13_py) 
  AUC_list_PD_C1[[i]] <- i.AUC
}
  
for (i in Label.C2b) {
  x <- data_sub_13C_flux_predrought %>%
    filter(Label == i)
  i.AUC <- auc(x$Trel, x$fCO2_13_py) 
  AUC_list_PD_C2[[i]] <- i.AUC
}

for (i in Label.C1b) {
  x <- data_sub_13C_flux_drought %>%
    filter(Label == i)
  i.AUC <- auc(x$Trel, x$fCO2_13_py) 
  AUC_list_D_C1[[i]] <- i.AUC
}

for (i in Label.C2a) {
  x <- data_sub_13C_flux_drought %>%
    filter(Label == i)
  i.AUC <- auc(x$Trel, x$fCO2_13_py) 
  AUC_list_D_C2[[i]] <- i.AUC
}

# Create a data frame with area under the curves for each sample for statistical analysis and plotting
# start by converting lists to data frames, and renaming columns in order to merge

AUC.D.C1 <- as.data.frame(do.call(rbind,AUC_list_D_C1))
AUC.D.C1$Label <- rownames(AUC.D.C1)
AUC.D.C1$AUC <- AUC.D.C1$V1
AUC.D.C1$Pyr <- "C1"
AUC.D.C1$Cond <- "Drought"
AUC.D.C1$V1 <- NULL
  
AUC.D.C2 <- as.data.frame(do.call(rbind,AUC_list_D_C2))
AUC.D.C2$Label <- rownames(AUC.D.C2)
AUC.D.C2$AUC <- AUC.D.C2$V1
AUC.D.C2$Pyr <- "C2"
AUC.D.C2$Cond <- "Drought"
AUC.D.C2$V1 <- NULL

AUC.PD.C1 <- as.data.frame(do.call(rbind, AUC_list_PD_C1))
AUC.PD.C1$Label <- rownames(AUC.PD.C1)
AUC.PD.C1$AUC <- AUC.PD.C1$V1
AUC.PD.C1$Pyr <- "C1"
AUC.PD.C1$Cond <- "Pre_Drought"
AUC.PD.C1$V1 <- NULL

AUC.PD.C2 <- as.data.frame(do.call(rbind, AUC_list_PD_C2))
AUC.PD.C2$Label <- rownames(AUC.PD.C2)
AUC.PD.C2$AUC <- AUC.PD.C2$V1
AUC.PD.C2$Pyr <- "C2"
AUC.PD.C2$Cond <- "Pre_Drought"
AUC.PD.C2$V1 <- NULL

# merge dataframes, AUC is in umol/cm2
AUC.C1 <- rbind(AUC.D.C1, AUC.PD.C1)
AUC.C2 <- rbind(AUC.D.C2, AUC.PD.C2)
AUC.0 <- rbind(AUC.C1, AUC.C2)


# calculate total mmol of C (note, AUC is in umol/m2, diameter of collar = 20 cm) 
# mmol: multiply AUC by area of collar (314.16cm2) to get umol, then divide by 10000 to convert from cm2 to m2 (100X100)
# mol: divide umol by 1000000 to convert to mol (10^6 umol in mol)
AUC <- AUC.0 %>%
  mutate(umol = ((AUC * 314.16) /10000)) %>%
  mutate(mol = umol / 1000000) %>%
  mutate(mass.13C.g = mol * 13) #add column with total mass of 13C (molar mass of 13C is 13 g/mol)

# calculate % of 13C pyruvate in 13C-CO2 (note, 0.10g of 13C-pyruvate added, which = 0.0011344894 mol(molar mass of 13C-pyruvate = 89.06g/mol),
# Since only one carbon atom labeled, 13C is also 0.0011344894 mol, and total mass of 13C can by calculated)
mass.13C.pyr.g = 0.0011344894 * 13

AUC <- AUC %>%
  mutate(perc.mass = (mass.13C.g/mass.13C.pyr.g) *100) %>%
  mutate(perc.mol = (mol/ 0.0011344894) * 100)

# calculate average aucs and umol of C
AUC_summary.CO2 <- AUC %>%
  group_by(Cond, Pyr) %>%
  summarise_at(vars(c(AUC, umol, mass.13C.g, perc.mass)), list(name = mean)) %>%
  mutate(compound = "CO2")

# total C for biosynthesis during drought (units in umol)
AUC_summary_wide <- AUC_summary.CO2 %>%
  dplyr::select(-c(AUC_name, perc.mass_name,mass.13C.g_name )) %>%
  pivot_wider(names_from = c(Pyr, Cond), values_from = umol_name) %>%
  mutate(Biosynthesis.D = C1_Drought - C2_Drought) %>%
  mutate(Biosynthesis.PD = C1_Pre_Drought - C2_Pre_Drought) %>%
  mutate(TCA_cycle.D = C1_Drought - Biosynthesis.D) %>%
  mutate(TCA_cycle.PD = C1_Pre_Drought - Biosynthesis.PD) 

# plot umol of 13C pyruvate (cummulative 13C-CO2)
plot <- AUC %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  ggplot(aes(x= Cond, y = umol, fill = Cond)) +
  geom_boxplot() +
  facet_grid(.~Pyr) +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  scale_x_discrete(labels = c("P", "D")) +
  labs(y = expression(""^13*"C-CO"[2]*" ("*mu*"mol)")) +
  theme(text = element_text(size = 12,
                          family = "Arial",
                          color = "black"),
      legend.position = "bottom",
      axis.title.x = element_blank(),
      legend.title = element_blank())
plot
#ggsave("../../Figures/Fig2-S1-CO2-VOCs/Fig2b-total-C-CO2-flux.png",units=c('in'),width=4, height = 4,dpi=300,plot)

# plot % of 13C pyruvate
plot <- AUC %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  ggplot(aes(x= Cond, y = perc.mass, fill = Cond)) +
  geom_boxplot() +
  facet_grid(.~Pyr) +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  #labs(y = expression("% of "^13*"C-pyruvate in "^13*"C-CO"[2])) + 
  labs(y= expression (""^13*"C-CO"[2]*"/ "^13*"C-pyruvate (%)")) +
  theme(text = element_text(size = 12,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot
#ggsave("../../Figures/Fig2-S1-CO2-VOCs/total-C-CO2-per-flux.png",units=c('in'),width=6, height = 4,dpi=300,plot)


```

# CO2 cummulative efflux statistics (wilcoxon and lme)
```{r CO2 cumulative efflux statistics, echo = FALSE}
# data summary

AUC_sumary <- AUC %>%
  dplyr::group_by(Cond, Pyr) %>%
  dplyr::summarize(mean_c_flux = mean(mass.13C.g),
                   sd = sd(mass.13C.g))

# statistics - Wilcoxon texts
# C1 vs C2 during pre-drought
AUC.PD <- AUC %>%
  filter(Cond == "Pre_Drought")

wilcox.test(mass.13C.g ~ Pyr, data = AUC.PD, paired = TRUE) #4.11e-5, paired 0.003906

# C1 vs C2 during drought
AUC.D <- AUC %>%
  filter(Cond == "Drought")

wilcox.test(mass.13C.g ~ Pyr, data = AUC.D, paired = TRUE) #0.00399, paired 0.007812

# pre-drought vs drought (C1)
AUC.C1 <- AUC %>%
  filter(Pyr == "C1")

wilcox.test(mass.13C.g ~ Cond, data = AUC.C1, paired = TRUE) #4.94e-4, paired .007812

# pre-drought vs drought (C2)
AUC.C2 <- AUC %>%
  filter(Pyr == "C2")

wilcox.test(mass.13C.g ~ Cond, data = AUC.C2, paired = TRUE) #2.76e-3, paired 0.02734

## linear mixed effect model
# subset to C1 and CO2.
statdat <- AUC %>%
  filter(Pyr == "C1") %>%
  separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
co2.c1 <- lme(mass.13C.g ~ Cond,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Cond)
            )
summary(co2.c1)

# check model assumptions
check_model(co2.c1)

# subset to C2 and CO2.
statdat <- AUC %>%
  filter(Pyr == "C2") %>%
   separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
co2.c2 <- lme(mass.13C.g ~ Cond,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Cond)
            )
summary(co2.c2)

# check model assumptions
check_model(co2.c2)

# subset to pre-drought and CO2.
statdat <- AUC %>%
  filter(Cond == "Pre_Drought") %>%
   separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
co2.preD <- lme(mass.13C.g ~ Pyr,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Pyr)
            )
summary(co2.preD)

# check model assumptions
check_model(co2.preD)

# subset to drought and CO2.
statdat <- AUC %>%
  filter(Cond == "Drought") %>%
   separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
co2.D <- lme(mass.13C.g ~ Pyr,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Pyr)
            )
summary(co2.D)

# check model assumptions
check_model(co2.D)


```


## cummulative efflux: Acetone
```{r acetone, echo=FALSE}
#Load 13C data. units for fluxes are umol/(m2*h)
df.acetone <- read.csv("../../Output/CO2-VOCs/acetone_flux-13C-all.csv", header = TRUE)

# calculate "pre" and "water" labeling 13C flux, so that we can calcualte 13C flux from pyruvate
data_sub_pre.aton <- df.acetone %>%
  filter(Time <= 0)

NP_pre.aton <- mean(data_sub_pre.aton$Flux)

data_sub_water.aton <- df.acetone %>%
  filter(Pyr == "W") 

NP_water.aton <- mean(data_sub_water.aton$Flux)

NP.aton <- mean(NP_pre.aton, NP_water.aton)

# filter time points to between 0 and 48 hrs
df.acetone$Time <- as.numeric(df.acetone$Time)


df.acetone.f <- df.acetone %>%
  mutate(Time.h = Time * 24) %>%
  dplyr::select(-Time) %>%
  filter(Time.h < 48 &
           Time.h > 0)

df.acetone.f$Time.h
df.acetone.f$Pyr
df.acetone.f$Condition
df.acetone.f$Label
df.acetone.f$Flux


###calculate area under the curve for C1 and C2 during pre-drought and drought
#Calculate AUC separately for each collar
df.acetone.f.d <- df.acetone.f %>%
  filter(Condition == "Drought")
df.acetone.f.p <- df.acetone.f %>%
  filter(Condition == "Pre-Drought") 



#Define samples that are either C1 or C2 labeled
Label.C1a = c("S1P1", "S1P3", "S1P5", "S2P1", "S2P3","S2P7", "S3P1","S3P3", "S3P5")
Label.C1b = c("S1P1", "S1P3", "S1P5", "S2P1", "S2P3", "S2P5", "S3P1", "S3P3", "S3P5")
Label.C2a = c("S1P2", "S1P4", "S1P6", "S2P2", "S2P4", "S2P6", "S3P2", "S3P4", "S3P6")
Label.C2b = c("S1P2", "S1P4", "S1P6", "S2P4", "S2P6", "S2P8", "S3P2", "S3P4", "S3P6")
Label.W = c("S1P2", "S1P4")

#create empty lists for each category
AUC.aton_list_PD_C1 <- list()
AUC.aton_list_PD_C2 <- list()
AUC.aton_list_D_C1 <- list()
AUC.aton_list_D_C2 <- list()
AUC.aton_list_D_W <- list()

df.acetone.f.d$Label

for (i in Label.C1a) {
  x <- df.acetone.f.p %>%
    filter(Label == i &
             Pyr != "W")
  i.AUC.aton <- auc(x$Time.h, x$Flux) 
  AUC.aton_list_PD_C1[[i]] <- i.AUC.aton
}

for (i in Label.C2b) {
  x <- df.acetone.f.p %>%
    filter(Label == i &
      Pyr != "W")
  i.AUC.aton <- auc(x$Time.h, x$Flux) 
  AUC.aton_list_PD_C2[[i]] <- i.AUC.aton
}


for (i in Label.C1b) {
  x <- df.acetone.f.d %>%
    filter(Label == i &
      Pyr != "W")
  i.AUC.aton <- auc(x$Time.h, x$Flux) 
  AUC.aton_list_D_C1[[i]] <- i.AUC.aton
}


for (i in Label.C2a) {
  x <- df.acetone.f.d %>%
    filter(Label == i &
             Pyr != "W")
  i.AUC.aton <- auc(x$Time.h, x$Flux) 
  AUC.aton_list_D_C2[[i]] <- i.AUC.aton
}

for (i in Label.W) {
  x <- df.acetone.f.d %>%
    filter(Label == i &
             Pyr == "W")
  i.AUC.aton <- auc(x$Time.h, x$Flux) 
  AUC.aton_list_D_W[[i]] <- i.AUC.aton
}

#Create a data frame with area under the curves for each sample for statistical analysis and plotting
#start by converting lists to data frames, and renaming columns in order to merge
#units for area under the curve are umol/m2

AUC.aton.D.C1 <- as.data.frame(do.call(rbind,AUC.aton_list_D_C1))
AUC.aton.D.C1$Label <- rownames(AUC.aton.D.C1)
AUC.aton.D.C1$AUC <- AUC.aton.D.C1$V1
AUC.aton.D.C1$Pyr <- "C1"
AUC.aton.D.C1$Cond <- "Drought"
AUC.aton.D.C1$V1 <- NULL

AUC.aton.D.C2 <- as.data.frame(do.call(rbind,AUC.aton_list_D_C2))
AUC.aton.D.C2$Label <- rownames(AUC.aton.D.C2)
AUC.aton.D.C2$AUC <- AUC.aton.D.C2$V1
AUC.aton.D.C2$Pyr <- "C2"
AUC.aton.D.C2$Cond <- "Drought"
AUC.aton.D.C2$V1 <- NULL

AUC.aton.PD.C1 <- as.data.frame(do.call(rbind, AUC.aton_list_PD_C1))
AUC.aton.PD.C1$Label <- rownames(AUC.aton.PD.C1)
AUC.aton.PD.C1$AUC <- AUC.aton.PD.C1$V1
AUC.aton.PD.C1$Pyr <- "C1"
AUC.aton.PD.C1$Cond <- "Pre_Drought"
AUC.aton.PD.C1$V1 <- NULL

AUC.aton.PD.C2 <- as.data.frame(do.call(rbind, AUC.aton_list_PD_C2))
AUC.aton.PD.C2$Label <- rownames(AUC.aton.PD.C2)
AUC.aton.PD.C2$AUC <- AUC.aton.PD.C2$V1
AUC.aton.PD.C2$Pyr <- "C2"
AUC.aton.PD.C2$Cond <- "Pre_Drought"
AUC.aton.PD.C2$V1 <- NULL

AUC.aton.D.W <- as.data.frame(do.call(rbind, AUC.aton_list_D_W))
AUC.aton.D.W$Label <- rownames(AUC.aton.D.W)
AUC.aton.D.W$AUC <- AUC.aton.D.W$V1
AUC.aton.D.W$Pyr <- "W"
AUC.aton.D.W$Cond <- "Drought"
AUC.aton.D.W$V1 <- NULL

AUC.aton.W.avg <- mean(AUC.aton.D.W$AUC)

# merge dataframes, AUC.aton is in mmol/m2
AUC.aton.C1 <- rbind(AUC.aton.D.C1, AUC.aton.PD.C1)
AUC.aton.C2 <- rbind(AUC.aton.D.C2, AUC.aton.PD.C2)
AUC.aton.0 <- rbind(AUC.aton.C1, AUC.aton.C2)
AUC.aton.W <- rbind(AUC.aton.C1, AUC.aton.C2, AUC.aton.D.W)


# calculate total mol of C (note, AUC.aton is in umol/m2, diameter of collar = 20 cm which means area is 314.16 cm2) 
AUC.aton <- AUC.aton.0 %>%
  mutate(AUC.a = AUC) %>%
  mutate(umol = ((AUC.a * 314.16) / 10000) ) %>% 
  mutate(mol = umol / 1000000) %>%
  mutate(mass.13C.g = mol * 13) #add column with total mass of 13C (molar mass of 13C is 13 g/mol)

# calculate % of 13C pyruvate in 13C-acetone (note, 0.0011228385 mol of 13C-pyruvate added,
# 0.0034067682 mol of C, but only one carbon atom labeled, so 0.0011344894 mol of labeled C)
mass.13C.pyr.g = 0.0011344894 * 13

AUC.aton <- AUC.aton %>%
  mutate(perc.mol = (mol/0.0011344894) *100) %>%
  mutate(perc.mass = ((mass.13C.g / mass.13C.pyr.g) * 100))


# calculate average AUC.atons and umol of C
AUC_summary.acetone <- AUC.aton %>%
  group_by(Cond, Pyr) %>%
  summarise_at(vars(c(AUC, umol, perc.mass)), list(name = mean)) %>%
  mutate(compound = "Acetone")


# plot cummalative flux of 13C pyruvate (C1 and C2)
plot <- AUC.aton %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  ggplot(aes(x= Cond, y = umol, fill = Cond)) +
  geom_boxplot() +
  facet_grid(.~Pyr) +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  labs(y = expression(""^13*"C-acetone ("*mu*"mol)")) +
  theme(text = element_text(size = 14,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot
ggsave("../../Figures/Fig2-S1-CO2-VOCs/corrected_voc_data/Fig2H-total-C-acetone-flux.png",units=c('in'),width=4, height = 4,dpi=300,plot)

# plot cummalative percentage of of 13C pyruvate (c1 and C2)
plot <- AUC.aton %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  ggplot(aes(x= Cond, y = perc.mass, fill = Cond)) +
  geom_boxplot() +
  facet_grid(.~Pyr) +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  labs(y = expression(""^13*"C-acetone from pyruvate (% )")) +
  theme(text = element_text(size = 12,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot
#ggsave("../../Figures/Fig2-S1-CO2-VOCs/C-acetone-perc-flux.png",units=c('in'),width=6, height = 4,dpi=300,plot)

# plot cummalative flux of 13C pyruvate for just C2, because only C2 showed enrichment
plot <- AUC.aton %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  filter(Pyr == "C2") %>%
  ggplot(aes(x= Cond, y = umol, fill = Cond)) +
  geom_boxplot() + 
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  scale_x_discrete(labels = c("P", "D")) +
  labs(y = expression(""^13*"C-acetone ("*mu*"mol)")) +
  theme(text = element_text(size = 14,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot
ggsave("../../Figures/Fig2-S1-CO2-VOCs/corrected_voc_data/Fig2H-total-C-acetone-flux-C2.png",units=c('in'),width=2, height = 4,dpi=300,plot)

# plot cummalative percentage of of 13C pyruvate
plot <- AUC.aton %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  filter(Pyr == "C2") %>%
  ggplot(aes(x= Cond, y = perc.mass, fill = Cond)) +
  geom_boxplot() +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  labs(y = expression("% of "^13*"C-pyruvate in "^13*"C-acetone")) + 
  theme(text = element_text(size = 12,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot
#ggsave("../../Figures/Fig2-S1-CO2-VOCs/C-acetone-perc-flux-C2.png",units=c('in'),width=4, height = 4,dpi=300,plot)

```


```{r acetone cumulative efflux statistics, echo = FALSE}
# data summary

AUC.sumary.aton <- AUC.aton %>%
  dplyr::group_by(Cond, Pyr) %>%
  dplyr::summarize(mean_c_flux = mean(mass.13C.g),
                   sd = sd(mass.13C.g))
# statistics
# C1 vs C2 during pre-drought
AUC.aton.PD <- AUC.aton %>%
  filter(Cond == "Pre_Drought")

wilcox.test(mass.13C.g ~ Pyr, data = AUC.aton.PD, paired = TRUE) #0.0002879, paired 0.007812

# C1 vs C2 during drought
AUC.aton.D <- AUC.aton %>%
  filter(Cond == "Drought")

wilcox.test(mass.13C.g ~ Pyr, data = AUC.aton.D, paired = TRUE) #4.114e-05 paired = 0.003906

# pre-drought vs drought (C1)
AUC.aton.C1 <- AUC.aton %>%
  filter(Pyr == "C1")

wilcox.test(mass.13C.g ~ Cond, data = AUC.aton.C1, paired = TRUE) #0.5457, paired = 0.3008

# pre-drought vs drought (C2)
AUC.aton.C2 <- AUC.aton %>%
  filter(Pyr == "C2")

wilcox.test(mass.13C.g ~ Cond, data = AUC.aton.C2, paired = TRUE) #8.227e-05, paired = 0.003906

## linear mixed effect model
# subset to C1. Add column for site so we can add this as a random variable.
statdat <- AUC.aton %>%
  filter(Pyr == "C1") %>%
  separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
aton.c1 <- lme(mass.13C.g ~ Cond,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Cond)
            )
summary(aton.c1)

# check model assumptions
check_model(aton.c1)

# subset to C2 and CO2.
statdat <- AUC.aton %>%
  filter(Pyr == "C2") %>%
 separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
aton.c2 <- lme(mass.13C.g ~ Cond,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Cond)
            )
summary(aton.c2)

# check model assumptions
check_model(aton.c2)


```




## Cummulative Efflux: Diacetyl

```{r}
########C13 data###############
#Load 13C data
#Load 13C data. units for fluxes are umol/(m2*h)
df.diacetyl <- read.csv("../../Output/CO2-VOCs/diacetyl_flux-13C-all.csv", header = TRUE)

# calculate "pre" and "water" labeling 13C flux, so that we can calcualte 13C flux from pyruvate
data_sub_pre.diac <- df.diacetyl %>%
  filter(Time <= 0)

NP_pre.diac <- mean(data_sub_pre.diac$Flux)

data_sub_water.diac <- df.diacetyl %>%
  filter(Pyr == "W") 

NP_water.diac <- mean(data_sub_water.diac$Flux)

NP.diac <- mean(NP_pre.diac, NP_water.diac)

# filter time points to between 0 and 48 hrs
df.diacetyl.f <- df.diacetyl %>%
  mutate(Time.h = Time * 24) %>%
  dplyr::select(-Time) %>%
  filter(Time.h < 48 &
           Time.h > 0)

df.diacetyl.f$Time.h
df.diacetyl.f$Pyr
df.diacetyl.f$Condition
df.diacetyl.f$Label
df.diacetyl.f$Flux


###calculate area under the curve for C1 and C2 during pre-drought and drought
#Calculate AUC separately for each collar
df.diacetyl.f.d <- df.diacetyl.f %>%
  filter(Condition == "Drought")
df.diacetyl.f.p <- df.diacetyl.f %>%
  filter(Condition == "Pre-Drought") 

df.diacetyl.f.p$Label

#Define samples that are either C1 or C2 labeled
Label.C1a = c("S1P1", "S1P3", "S1P5", "S2P1", "S2P3","S2P7", "S3P1","S3P3", "S3P5")
Label.C1b = c("S1P1", "S1P3", "S1P5", "S2P1", "S2P3", "S2P5", "S3P1", "S3P3", "S3P5")
Label.C2a = c("S1P2", "S1P4", "S1P6", "S2P2", "S2P4", "S2P6", "S3P2", "S3P4", "S3P6")
Label.C2b = c("S1P2", "S1P4", "S1P6", "S2P4", "S2P6", "S2P8", "S3P2", "S3P4", "S3P6")
Label.W = c("S1P2", "S1P4")

#create empty lists for each category
AUC_list_PD_C1_Diac <- list()
AUC_list_PD_C2_Diac <- list()
AUC_list_D_C1_Diac <- list()
AUC_list_D_C2_Diac <- list()
AUC_list_D_W_Diac <- list()

df.diacetyl.f.p$Label

for (i in Label.C1a) {
  x <- df.diacetyl.f.p %>%
    filter(Label == i &
             Pyr != "W" &
           Flux > 0)
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC_list_PD_C1_Diac[[i]] <- i.AUC
}


for (i in Label.C2b) {
  x <- df.diacetyl.f.p %>%
    filter(Label == i &
             Pyr != "W"&
             Flux > 0)
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC_list_PD_C2_Diac[[i]] <- i.AUC
}


for (i in Label.C1b) {
  x <- df.diacetyl.f.d %>%
    filter(Label == i &
             Pyr != "W" &
             Flux > 0)
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC_list_D_C1_Diac[[i]] <- i.AUC
}


for (i in Label.C2a) {
  x <- df.diacetyl.f.d %>%
    filter(Label == i &
             Pyr != "W"&
             Flux > 0)
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC_list_D_C2_Diac[[i]] <- i.AUC
}

for (i in Label.W) {
  x <- df.diacetyl.f.d %>%
    filter(Label == i &
             Pyr == "W"&
             Flux > 0)
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC_list_D_W_Diac[[i]] <- i.AUC
}

#Create a data frame with area under the curves for each sample for statistical analysis and plotting
#start by converting lists to data frames, and renaming columns in order to merge

AUC.D.C1.Diac <- as.data.frame(do.call(rbind,AUC_list_D_C1_Diac))
AUC.D.C1.Diac$Label <- rownames(AUC.D.C1.Diac)
AUC.D.C1.Diac$AUC <- AUC.D.C1.Diac$V1
AUC.D.C1.Diac$Pyr <- "C1"
AUC.D.C1.Diac$Cond <- "Drought"
AUC.D.C1.Diac$V1 <- NULL

AUC.D.C2.Diac <- as.data.frame(do.call(rbind,AUC_list_D_C2_Diac))
AUC.D.C2.Diac$Label <- rownames(AUC.D.C2.Diac)
AUC.D.C2.Diac$AUC <- AUC.D.C2.Diac$V1
AUC.D.C2.Diac$Pyr <- "C2"
AUC.D.C2.Diac$Cond <- "Drought"
AUC.D.C2.Diac$V1 <- NULL

AUC.PD.C1.Diac <- as.data.frame(do.call(rbind, AUC_list_PD_C1_Diac))
AUC.PD.C1.Diac$Label <- rownames(AUC.PD.C1.Diac)
AUC.PD.C1.Diac$AUC <- AUC.PD.C1.Diac$V1
AUC.PD.C1.Diac$Pyr <- "C1"
AUC.PD.C1.Diac$Cond <- "Pre_Drought"
AUC.PD.C1.Diac$V1 <- NULL

AUC.PD.C2.Diac <- as.data.frame(do.call(rbind, AUC_list_PD_C2_Diac))
AUC.PD.C2.Diac$Label <- rownames(AUC.PD.C2.Diac)
AUC.PD.C2.Diac$AUC <- AUC.PD.C2.Diac$V1
AUC.PD.C2.Diac$Pyr <- "C2"
AUC.PD.C2.Diac$Cond <- "Pre_Drought"
AUC.PD.C2.Diac$V1 <- NULL

AUC.D.W.Diac <- as.data.frame(do.call(rbind, AUC_list_D_W_Diac))
AUC.D.W.Diac$Label <- rownames(AUC.D.W.Diac)
AUC.D.W.Diac$AUC <- AUC.D.W.Diac$V1
AUC.D.W.Diac$Pyr <- "W"
AUC.D.W.Diac$Cond <- "Drought"
AUC.D.W.Diac$V1 <- NULL

AUC.W.avg.Diac <- mean(AUC.D.W.Diac$AUC)


# merge dataframes, AUC is in umol/m2
AUC.C1.Diac <- rbind(AUC.D.C1.Diac, AUC.PD.C1.Diac)
AUC.C2.Diac <- rbind(AUC.D.C2.Diac, AUC.PD.C2.Diac)
AUC.0.Diac <- rbind(AUC.C1.Diac, AUC.C2.Diac)
AUC.W.Diac <- rbind(AUC.C1.Diac, AUC.C2.Diac, AUC.D.W.Diac)


# calculate total mol of C (note, AUC is in mmol/m2, diameter of collar = 20 cm which means area is 314.16 cm2) 
AUC.Diac <- AUC.0.Diac %>%
  mutate(AUC.a = AUC - AUC.W.avg.Diac) %>% #Subtract natural abundance of 13C
  mutate(umol = ((AUC.a * 314.16) / 10000)) %>% 
  mutate(mol = umol / 1000000) %>%
  mutate(mass.13C.g = mol * 13) #add column with total mass of 13C (molar mass of 13C is 13 g/mol)

# calculate % of 13C pyruvate in 13C-diacetyl (note, 0.0011228385 mol of 13C-pyruvate added,
# 0.0034067682 mol of C, but only one carbon atom labeled, so 0.0011344894 mol of labeled C)
mass.13C.pyr.g = 0.0011228385 * 13

AUC.Diac <- AUC.Diac %>%
  mutate(perc.mol = (mol/0.0011228385) *100) %>%
  mutate(perc.mass = ((mass.13C.g / mass.13C.pyr.g) * 100))

# calculate average aucs and nmol of C
AUC_summary.Diac <- AUC.Diac %>%
  group_by(Cond, Pyr) %>%
  summarise_at(vars(c(AUC, umol, perc.mass)), list(name = mean)) %>%
  mutate(compound = "Diacetyl")


# plot cummalative flux of 13C 
plot.Diac <- AUC.Diac %>%
 mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  ggplot(aes(x= Cond, y = umol, fill = Cond)) +
  geom_boxplot() +
  facet_grid(.~Pyr) +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("P", "D")) +
  scale_x_discrete(labels = c("P", "D")) +
  labs(y = expression(""^13*"C-diacetyl+ ("*mu*"mol)")) +
  theme(text = element_text(size = 14,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot.Diac
ggsave("../../Figures/Fig2-S1-CO2-VOCs/corrected_voc_data/Fig2F-total-C-diacetyl-flux-c1-and-c2.png",units=c('in'),width=4, height = 4,dpi=300,plot.Diac)

# plot cummalative percentage of of 13C pyruvate
plot.Diac <- AUC.Diac %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  ggplot(aes(x= Cond, y = perc.mass, fill = Cond)) +
  geom_boxplot() +
  facet_grid(.~Pyr) +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  labs(y = expression("Cummulative efflux of "^13*"C-diacetyl (% of 13C from pyruvate)")) +
  theme(text = element_text(size = 12,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot.Diac
#ggsave("../../Figures/Fig2-S1-CO2-VOCs/C-diacetyl-perc-flux.png",units=c('in'),width=6, height = 4,dpi=300,plot.Diac)

# plot cummalative flux of 13C diacetyl, only from C2 
plot.Diac <- AUC.Diac %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  filter(Pyr == "C2") %>%
  ggplot(aes(x= Cond, y = umol, fill = Cond)) +
  geom_boxplot() +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  scale_x_discrete(labels = c("P", "D")) +
  labs(y = expression(""^13*"C-diacetyl ("*mu*"mol)")) +
  theme(text = element_text(size = 14,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.text.x = element_text(size = 12),
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot.Diac
ggsave("../../Figures/Fig2-S1-CO2-VOCs/corrected_voc_data/Fig2I-total-C-diacetyl-flux-c2.png",units=c('in'),width=2, height = 4,dpi=300,plot.Diac)

# plot cummalative percentage of of 13C pyruvate
plot.Diac <- AUC.Diac %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  filter(Pyr == "C2") %>%
  ggplot(aes(x= Cond, y = perc.mass, fill = Cond)) +
  geom_boxplot() +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("Pre Drought", "Drought")) +
  labs(y = expression("% of "^13*"C-pyruvate in "^13*"C-diacetyl")) + 
  theme(text = element_text(size = 12,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot.Diac
#ggsave("../../Figures/Fig2-S1-CO2-VOCs/C-diacetyl-perc-flux-c2.png",units=c('in'),width=4, height = 4,dpi=300,plot.Diac)


```

```{r diacetyl statistics for AUC, echo = FALSE}
# data summary

AUC.sumary.diac <- AUC.Diac %>%
  dplyr::group_by(Cond, Pyr) %>%
  dplyr::summarize(mean_c_flux = mean(mass.13C.g),
                   sd = sd(mass.13C.g))
AUC.sumary.diac
# statistics
# C1 vs C2 during pre-drought
AUC.Diac.PD <- AUC.Diac %>%
  filter(Cond == "Pre_Drought")

wilcox.test(mass.13C.g ~ Pyr, data = AUC.Diac.PD, paired = TRUE) #0.2973, paired 0.25

# C1 vs C2 during drought
AUC.Diac.D <- AUC.Diac %>%
  filter(Cond == "Drought")

wilcox.test(mass.13C.g ~ Pyr, data = AUC.Diac.D, paired = TRUE) #0.7304, paired 0.6523

# pre-drought vs drought (C1)
AUC.Diac.C1 <- AUC.Diac %>%
  filter(Pyr == "C1")

wilcox.test(mass.13C.g ~ Cond, data = AUC.Diac.C1, paired = TRUE) #0.2224, apried 0.1289

# pre-drought vs drought (C2)
AUC.Diac.C2 <- AUC.Diac %>%
  filter(Pyr == "C2")

wilcox.test(mass.13C.g ~ Cond, data = AUC.Diac.C2, paired = TRUE) #0.05031, paired 0.03906

## linear mixed effect model
# subset to C1. Add column for site so we can add this as a random variable.
statdat <- AUC.Diac %>%
  filter(Pyr == "C1") %>%
  separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
diac.c1 <- lme(mass.13C.g ~ Cond,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Cond)
            )
summary(diac.c1)

# check model assumptions
check_model(diac.c1)

# subset to C2 and CO2.
statdat <- AUC.Diac %>%
  filter(Pyr == "C2") %>%
 separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
diac.c2 <- lme(mass.13C.g ~ Cond,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Cond)
            )
summary(diac.c2)

# check model assumptions
check_model(diac.c2)

```





## Cummulative Effluxes: Acetic acid
```{r acetic acid, echo = FALSE}
########C13 data###############
#Load 13C data
df.acetate <- read.csv("../../Output/CO2-VOCs/acetate_flux-13C-all.csv", header = TRUE)


# calculate "pre" and "water" labeling 13C flux, so that we can calcualte 13C flux from pyruvate
data_sub_pre <- df.acetate %>%
  filter(Time <= 0)

NP_pre <- mean(data_sub_pre$Flux)

data_sub_water <- df.acetate %>%
  filter(Pyr == "W")

NP_water <- mean(data_sub_water$Flux)

NP <- mean(NP_pre, NP_water)

# filter time points to between 0 and 48 hrs
df.acetate.f <- df.acetate %>%
  mutate(Time.h = Time * 24) %>%
  dplyr::select(-Time) %>%
  filter(Time.h < 48 &
           Time.h > 0)

df.acetate.f$Time.h
df.acetate.f$Pyr
df.acetate.f$Condition
df.acetate.f$Label


###calculate area under the curve for C1 and C2 during pre-drought and drought
#Calculate AUC separately for each collar
df.acetate.f.d <- df.acetate.f %>%
  filter(Condition == "Drought")
df.acetate.f.p <- df.acetate.f %>%
  filter(Condition == "Pre-Drought") 



#Define samples that are either C1 or C2 labeled
Label.C1a = c("S1P1", "S1P3", "S1P5", "S2P1", "S2P3", "S2P7", "S3P1", "S3P3", "S3P5")
Label.C1b = c("S1P1", "S1P3", "S1P5", "S2P1", "S2P3", "S2P5", "S3P1", "S3P3", "S3P5")
Label.C2a = c("S1P2", "S1P4", "S1P6", "S2P2", "S2P4", "S2P6", "S3P2", "S3P4", "S3P6")
Label.C2b = c("S1P2", "S1P4", "S1P6", "S2P4", "S2P6", "S2P8", "S3P2", "S3P4", "S3P6")
Label.W = c("S1P2", "S1P4")

#create empty lists for each category
AUC.acetate_list_PD_C1 <- list()
AUC.acetate_list_PD_C2 <- list()
AUC.acetate_list_D_C1 <- list()
AUC.acetate_list_D_C2 <- list()
AUC.acetate_list_D_W <- list()


for (i in Label.C1a) {
  x <- df.acetate.f.p %>%
    filter(Label == i &
             Pyr != "W")
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC.acetate_list_PD_C1[[i]] <- i.AUC
}


for (i in Label.C2b) {
  x <- df.acetate.f.p %>%
    filter(Label == i &
             Pyr != "W")
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC.acetate_list_PD_C2[[i]] <- i.AUC
}


for (i in Label.C1b) {
  x <- df.acetate.f.d %>%
    filter(Label == i &
             Pyr != "W")
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC.acetate_list_D_C1[[i]] <- i.AUC
}


for (i in Label.C2a) {
  x <- df.acetate.f.d %>%
    filter(Label == i &
             Pyr != "W")
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC.acetate_list_D_C2[[i]] <- i.AUC
}

for (i in Label.W) {
  x <- df.acetate.f.d %>%
    filter(Label == i &
             Pyr == "W")
  i.AUC <- auc(x$Time.h, x$Flux) 
  AUC.acetate_list_D_W[[i]] <- i.AUC
}

#Create a data frame with area under the curves for each sample for statistical analysis and plotting
#start by converting lists to data frames, and renaming columns in order to merge

AUC.acetate.D.C1 <- as.data.frame(do.call(rbind,AUC.acetate_list_D_C1))
AUC.acetate.D.C1$Label <- rownames(AUC.acetate.D.C1)
AUC.acetate.D.C1$AUC <- AUC.acetate.D.C1$V1
AUC.acetate.D.C1$Pyr <- "C1"
AUC.acetate.D.C1$Cond <- "Drought"
AUC.acetate.D.C1$V1 <- NULL

AUC.acetate.D.C2 <- as.data.frame(do.call(rbind,AUC.acetate_list_D_C2))
AUC.acetate.D.C2$Label <- rownames(AUC.acetate.D.C2)
AUC.acetate.D.C2$AUC <- AUC.acetate.D.C2$V1
AUC.acetate.D.C2$Pyr <- "C2"
AUC.acetate.D.C2$Cond <- "Drought"
AUC.acetate.D.C2$V1 <- NULL

AUC.acetate.PD.C1 <- as.data.frame(do.call(rbind, AUC.acetate_list_PD_C1))
AUC.acetate.PD.C1$Label <- rownames(AUC.acetate.PD.C1)
AUC.acetate.PD.C1$AUC <- AUC.acetate.PD.C1$V1
AUC.acetate.PD.C1$Pyr <- "C1"
AUC.acetate.PD.C1$Cond <- "Pre_Drought"
AUC.acetate.PD.C1$V1 <- NULL

AUC.acetate.PD.C2 <- as.data.frame(do.call(rbind, AUC.acetate_list_PD_C2))
AUC.acetate.PD.C2$Label <- rownames(AUC.acetate.PD.C2)
AUC.acetate.PD.C2$AUC <- AUC.acetate.PD.C2$V1
AUC.acetate.PD.C2$Pyr <- "C2"
AUC.acetate.PD.C2$Cond <- "Pre_Drought"
AUC.acetate.PD.C2$V1 <- NULL

AUC.acetate.D.W <- as.data.frame(do.call(rbind, AUC.acetate_list_D_W))
AUC.acetate.D.W$Label <- rownames(AUC.acetate.D.W)
AUC.acetate.D.W$AUC <- AUC.acetate.D.W$V1
AUC.acetate.D.W$Pyr <- "W"
AUC.acetate.D.W$Cond <- "Drought"
AUC.acetate.D.W$V1 <- NULL

AUC.acetate.W.avg <- mean(AUC.acetate.D.W$AUC)


# merge dataframes, AUC is in umol/m2
AUC.acetate.C1 <- rbind(AUC.acetate.D.C1, AUC.acetate.PD.C1)
AUC.acetate.C2 <- rbind(AUC.acetate.D.C2, AUC.acetate.PD.C2)
AUC.acetate.0 <- rbind(AUC.acetate.C1, AUC.acetate.C2)
AUC.acetate.W <- rbind(AUC.acetate.C1, AUC.acetate.C2, AUC.acetate.D.W)


# calculate total mol of C (note, AUC is in umol/m2, diameter of collar = 20 cm which means area is 314.16 cm2) 
#add column with total mass of 13C from pyruvate in aceate (molar mass of 13C is 13 g/mol)
AUC.acetate <- AUC.acetate.0 %>%
  mutate(AUC.a = AUC - AUC.acetate.W.avg) %>%
  mutate(umol = ((AUC.a * 314.16) /10000 ))  %>%
  mutate(mol = umol / 1000000) %>%
  mutate(mass.13C.g = (mol*13) )

# calculate % of 13C pyruvate in 13C-acetate (note, 0.0011355894 mol of 13C-pyruvate added,
# 0.0011355894 mol of 13C)
mass.13C.pyr.g = 0.0011355894 * 13

AUC.acetate <- AUC.acetate %>%
  mutate(perc.mol = (mol/0.0011355894) *100) %>%
  mutate(perc.mass = ((mass.13C.g / mass.13C.pyr.g) * 100))

# calculate average aucs and nmol of C
AUC_summary.acetate <- AUC.acetate %>%
  group_by(Cond, Pyr) %>%
  summarise_at(vars(c(AUC, umol, perc.mass)), list(name = mean)) %>%
  mutate(compound = "acetate")


# plot cummalative flux of 13C pyruvate
plot <- AUC.acetate %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  ggplot(aes(x= Cond, y = umol, fill = Cond)) +
  geom_boxplot() +
  facet_grid(.~Pyr) +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("P", "D")) +
  scale_x_discrete(labels = c("P", "D")) +
  labs(y = expression(""^13*"C-acetic acid ("*mu*"mol)")) +
  theme(text = element_text(size = 14,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot
ggsave("../../Figures/Fig2-S1-CO2-VOCs/corrected_voc_data/Fig2G-total-C-acetate-flux.png",units=c('in'),width=4, height = 4,dpi=300,plot)

# plot cummalative percentage of of 13C pyruvate
plot <- AUC.acetate %>%
  mutate(Cond=factor(Cond, levels = c("Pre_Drought", "Drought"))) %>%
  ggplot(aes(x= Cond, y = perc.mass, fill = Cond)) +
  geom_boxplot() +
  facet_grid(.~Pyr) +
  scale_fill_manual(values = colors,
                    breaks = c("Pre_Drought", "Drought"),
                    labels = c("P", "D")) +
  labs(y = expression("% of "^13*"C-pyruvate in "^13*"C-acetone")) + 
  theme(text = element_text(size = 12,
                            family = "Arial",
                            color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        legend.title = element_blank())
plot
#ggsave("../../Figures/Fig2-S1-CO2-VOCs/C-acetate-perc-flux.png",units=c('in'),width=6, height = 4,dpi=300,plot)
```



```{r statistics acetate for AUC, echo = FALSE}
# data summary
AUC.sumary.acetate<- AUC.acetate %>%
  dplyr::group_by(Cond, Pyr) %>%
  dplyr::summarize(mean_c_flux = mean(mass.13C.g),
                   sd = sd(mass.13C.g))
AUC.sumary.acetate

# statistics
# C1 vs C2 during pre-drought
AUC.acetate.PD <- AUC.acetate %>%
  filter(Cond == "Pre_Drought")

wilcox.test(mass.13C.g ~ Pyr, data = AUC.acetate.PD, paired = TRUE) #0.6665, paired 0.5703

# C1 vs C2 during drought
AUC.acetate.D <- AUC.acetate %>%
  filter(Cond == "Drought")

wilcox.test(mass.13C.g ~ Pyr, data = AUC.acetate.D, paired = TRUE) #0.7962, paired 0.8203

# pre-drought vs drought (C1)
AUC.acetate.C1 <- AUC.acetate %>%
  filter(Pyr == "C1")

wilcox.test(mass.13C.g ~ Cond, data = AUC.acetate.C1, paired = TRUE) #0.002756, paired = 0.007812

# pre-drought vs drought (C2)
AUC.acetate.C2 <- AUC.acetate %>%
  filter(Pyr == "C2")

wilcox.test(mass.13C.g ~ Cond, data = AUC.acetate.C2, paired = TRUE) #0.0001645, paired = 0.0039

## linear mixed effect model
# subset to C1. Add column for site so we can add this as a random variable.
statdat <- AUC.acetate %>%
  filter(Pyr == "C1") %>%
  separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
acetate.c1 <- lme(mass.13C.g ~ Cond,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Cond)
            )
summary(acetate.c1)

# check model assumptions
check_model(acetate.c1)

# subset to C2 and CO2.
statdat <- AUC.acetate %>%
  filter(Pyr == "C2") %>%
 separate(Label, into = c("Site", "ID"), sep = "P", remove = FALSE)

# lme model
acetate.c2 <- lme(mass.13C.g ~ Cond,
            random = list(Site = ~1,
                          ID =  ~1),
            data = statdat,
            weights =  varIdent(form = ~1|Cond)
            )
summary(acetate.c2)

# check model assumptions
check_model(acetate.c2)
```


## compare cummulative fluxes

```{r}
# combine summary tables to compare cummulative fluxes

# create summary of umol and perc
summary.1 <- rbind(AUC_summary.acetate, AUC_summary.acetone)
summary.2 <- AUC_summary.CO2 %>%
  dplyr::select(-mass.13C.g_name) %>% #remove column that doesn't match with diacetyl summary table
  rbind(AUC_summary.Diac)
summary.total <- rbind(summary.1, summary.2)
summary.total

# spread c1 and C2 across value columns (umol and perc)
summary.umol <- summary.total %>%
  dplyr::select(-c(AUC_name)) %>% #remove columns that we don't want to include in final summary table
  pivot_wider(names_from = Pyr, values_from = c("umol_name", "perc.mass_name")) %>%
  mutate(diff.umol = umol_name_C1 - umol_name_C2) %>% #calculate C1 - C2 so we know how much 13C within 13C-co2 goes towards biosynthesis 
  mutate(sum.umol = umol_name_C1 + umol_name_C2) %>%
  mutate(diff.perc = perc.mass_name_C1 - perc.mass_name_C2) %>%
  mutate(sum.perc = perc.mass_name_C1 + perc.mass_name_C2)  # calculate C1 + C2 for a sum of total labeled acetae
summary.umol


# make pie charts showing biosynthesis  and tca cycle
# drought
summary.allocation.d <- summary.umol %>%
  filter(compound == "CO2" &
           Cond == "Drought") %>%
  mutate(other = 100 - (perc.mass_name_C1)) %>%
  dplyr::select(perc.mass_name_C2, diff.perc, other)

summary.allocation.d.t <- t(summary.allocation.d)
colnames(summary.allocation.d.t) <- summary.allocation.d.t[1,]
summary.allocation.d.t <- summary.allocation.d.t[-1,]
summary.allocation.d.t <- as.data.frame(summary.allocation.d.t)
summary.allocation.d.t$perc.alloc <- summary.allocation.d.t$summary.allocation.d.t
summary.allocation.d.t$perc.alloc <- as.numeric(summary.allocation.d.t$perc.alloc)
summary.allocation.d.t$cat.alloc <- rownames(summary.allocation.d.t)

pie.d <- ggplot(summary.allocation.d.t, aes(x="", y=perc.alloc, fill =cat.alloc )) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y" , start = 0) +
 scale_fill_manual(breaks = c("diff.perc", "perc.mass_name_C2", "other"),
                    labels = c("Biosynthesis", "TCA cycle","Other"),
                    values = c("blue", "red","darkgray")) +
  theme_void() +
  theme(legend.title = element_blank())
pie.d
ggsave("../../Figures/Fig2-S1-CO2-VOCs/corrected_voc_data/Fig2J-pie_d_bio_tca.png",units=c('in'),width=3, height = 3,dpi=300,pie.d)


# pre-drought
summary.allocation.p <- summary.umol %>%
  filter(compound == "CO2" &
           Cond == "Pre_Drought") %>%
  mutate(other = 100 - (perc.mass_name_C1)) %>%
  dplyr::select(perc.mass_name_C2, diff.perc, other)

summary.allocation.p.t <- t(summary.allocation.p)
colnames(summary.allocation.p.t) <- summary.allocation.p.t[1,]
summary.allocation.p.t <- summary.allocation.p.t[-1,]
summary.allocation.p.t <- as.data.frame(summary.allocation.p.t)
summary.allocation.p.t$perc.alloc <- summary.allocation.p.t$summary.allocation.p.t
summary.allocation.p.t$perc.alloc <- as.numeric(summary.allocation.p.t$perc.alloc)
summary.allocation.p.t$cat.alloc <- rownames(summary.allocation.p.t)

pie.p <- ggplot(summary.allocation.p.t, aes(x="", y=perc.alloc, fill =cat.alloc )) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y" , start = 0) +
  scale_fill_manual(breaks = c("diff.perc", "perc.mass_name_C2", "other"),
                    labels = c("Biosynthesis", "TCA cycle","Other"),
                    values = c("blue", "red","darkgray")) +
  theme_void() +
  theme(legend.title = element_blank())
pie.p
ggsave("../../Figures/Fig2-S1-CO2-VOCs/corrected_voc_data/Fig2J-pie_p_bio_tca.png",units=c('in'),width=3, height = 3,dpi=300,pie.p)

# #####piechart showing proportion of biosynthesis that are VOCs
#first set up data table to plot
#first assign variable for biosynthesis during pre-drought
biosynthesis.p <- summary.umol %>%
  filter(compound == "CO2" & Cond == "Pre_Drought") %>%
  dplyr::select(diff.umol)
biosynthesis.p <- biosynthesis.p$diff.umol

#filter table to just include VOCs and calculate percent biosynthesis of total
summary.umol.voc.p <- summary.umol %>%
  filter(compound != "CO2") %>%
  filter(Cond == "Pre_Drought") %>%
  dplyr::select(compound, umol_name_C2) %>%
  mutate(perc.bio = (umol_name_C2/biosynthesis.p) * 100)

#transpose table to calculate 'other' category
summary.umol.voc.p.t <- t(summary.umol.voc.p)
summary.umol.voc.p.t <- as.data.frame(summary.umol.voc.p.t)
summary.umol.voc.p.t <- summary.umol.voc.p.t[-1,] 
colnames(summary.umol.voc.p.t) <- summary.umol.voc.p.t[1,]
summary.umol.voc.p.t <- summary.umol.voc.p.t[-1,]
summary.umol.voc.p.t$acetate <- as.numeric(summary.umol.voc.p.t$acetate)
summary.umol.voc.p.t$Acetone <- as.numeric(summary.umol.voc.p.t$Acetone)
summary.umol.voc.p.t$Diacetyl <- as.numeric(summary.umol.voc.p.t$Diacetyl)

summary.umol.voc.p.t <- summary.umol.voc.p.t %>%
  dplyr::select(-acetate)  %>% #because negative values (consumption)
  mutate(Other = (100 - (Acetone + Diacetyl))) 

#transpose table again so easy to graph
summary.umol.voc.p.final <- as.data.frame(t(summary.umol.voc.p.t))
summary.umol.voc.p.final$VOC <- rownames(summary.umol.voc.p.final)
summary.umol.voc.p.final

pie.p <- ggplot(summary.umol.voc.p.final, aes(x="", y=perc.bio, fill =VOC)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y" , start = 0) +
  scale_fill_manual(breaks = c("Acetone", "Diacetyl", "Other"),
                    values = c("lightblue", "yellow","darkgray")) +
  theme_void() +
  theme(legend.title = element_blank())
pie.p

# #####piechart showing proportion of biosynthesis that are VOCs
#first set up data table to plot
#first assign variable for biosynthesis during drought
biosynthesis.d <- summary.umol %>%
  filter(compound == "CO2" & Cond == "Drought") %>%
  dplyr::select(diff.umol)
biosynthesis.d <- biosynthesis.d$diff.umol

#filter table to just include VOCs and calculate percent biosynthesis of total
summary.umol.voc.d <- summary.umol %>%
  filter(compound != "CO2") %>%
  filter(Cond == "Drought") %>%
  dplyr::select(compound, umol_name_C2) %>%
  mutate(perc.bio = ((umol_name_C2/biosynthesis.d) * 100))

#transpose table to calculate 'other' category
summary.umol.voc.d.t <- t(summary.umol.voc.d)
summary.umol.voc.d.t <- as.data.frame(summary.umol.voc.d.t)
summary.umol.voc.d.t <- summary.umol.voc.d.t[-1,] 
colnames(summary.umol.voc.d.t) <- summary.umol.voc.d.t[1,]
summary.umol.voc.d.t <- summary.umol.voc.d.t[-1,]
summary.umol.voc.d.t$acetate <- as.numeric(summary.umol.voc.d.t$acetate)
summary.umol.voc.d.t$Acetone <- as.numeric(summary.umol.voc.d.t$Acetone)
summary.umol.voc.d.t$Diacetyl <- as.numeric(summary.umol.voc.d.t$Diacetyl)

summary.umol.voc.d.t <- summary.umol.voc.d.t %>%
  #dplyr::select(-acetate)  %>% #because negative values (consumption) (changing this because values are positive now)
  mutate(Other = (100 - (Acetone + Diacetyl + acetate))) 

#transpose table again so easy to graph
summary.umol.voc.d.final <- as.data.frame(t(summary.umol.voc.d.t))
summary.umol.voc.d.final$VOC <- rownames(summary.umol.voc.d.final)
summary.umol.voc.d.final
  
pie.d <- ggplot(summary.umol.voc.d.final, aes(x="", y=perc.bio, fill =VOC)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y" , start = 0) +
  scale_fill_manual(breaks = c("Acetone", "Diacetyl", "Acetate", "Other"),
                    values = c("lightblue", "yellow","lightgreen", "darkgray")) +
  theme_void() +
  theme(legend.title = element_blank())
pie.d


```

# notes on summary table (haven't updated with corrected results):
- diff in C1 - c2 for CO2 represents umol of 13C that go towards biosynthesis 
    - Pre-drought: 465.48 umol
    - Drought 196.59 umol  
- sum of acetate, acetone, and diaceytl (c2) represent 13C from biosynthesis pathways 
   
VOC biosynthesis represents 0.11 % of total biosynthesis during Drought, and only 0.0022 % during Pre-drought!! Really cool. 

ALSO.... Comparing microbially-mediated carbon emissions... 
- Pre-drought 465.48umol co2 + 10.4 umol voc = 475.88 umol, 
- Drought 196.59 umol co2 + 122.68 umol voc = 319.27 umol

Overall distribution of 13C-pyruvate incorporation:
- Pre-drought
  - 

```{r}
