---
title: "soil-moisture"
author: "Linnea Honeker"
date: "4/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("dplyr")
library("ggplot2")
library("sjPlot")
library("stargazer")

colors = c("#2ECC71" ,"#F39C12")

```

## Soil Moisture
Data was obtained from soil moisture collected along with cartridge sampling during dates that were close to pyruvate labeling. During pre-drought, there is soil moisture from 9/30 and 10/9/19, and for drought, there is soil moisture from 11/10/19 and 11/17/19.

```{r load_data}
#load soil moisture data
data <- read.csv("../Data/soil_moisture/soil_moisture.csv", header = TRUE)
data$Condition <- factor(data$Condition, c("PreDrought", "Drought"))

#add soil moisture means
data <- data %>%
  replace(is.na(.), 0) %>%
  rowwise() %>%
   mutate(SM = (SM.1 + SM.2 + SM.3)/3, SM.sd = sd(c(SM.1, SM.2, SM.3))) %>%
   filter(SM != 0)

#load pH data
data.pH <- read.csv("../Data/soil_moisture/pH.csv", header = TRUE)
data.pH$Condition <- factor(data$Condition, c("PreDrought", "Drought"))


```

## Plot soil moisture


```{r plot_soil_moisture}
#overall 
p1 <- ggplot(data, aes(x = Condition, y= SM, fill = Condition)) +
 geom_boxplot() +
  scale_fill_manual(values = colors) +
  labs(y = "Soil Moisture %")
  
p1


#by site
p2 <- ggplot(data, aes(x = Condition, y= SM, fill = Condition)) +
 geom_boxplot() +
  facet_grid(~Site) +
  scale_fill_manual(values = colors)
  
p2
```

## Plot soil temperature
```{r plot_soil_temp}
#overall 
p1 <- ggplot(data, aes(x = Condition, y= ST)) +
 geom_boxplot() 
p1

#by site
p2 <- ggplot(data, aes(x = Condition, y= ST)) +
 geom_boxplot() +
  facet_grid(~Site)
p2
```
```{r plot_pH}
#overall 
p1 <- ggplot(data.pH, aes(x = Condition, y= pH, fill = Condition)) +
 geom_boxplot() +
  scale_fill_manual(values = colors) +
  labs(y = "pH")
  
p1


#by site
p2 <- ggplot(data.pH, aes(x = Site, y= pH, fill = Site)) +
 geom_boxplot() +
  #facet_grid(~Site) +
  scale_fill_manual(values = c(colors, "blue"))
  
p2
```
## Return summary statistics for PreDrought and Drought
```{r}
#soil moisture statistics
data.PD <- data %>%
  filter(Condition == "PreDrought") 

data.PD.mean <- mean(data.PD$SM)
data.PD.sd <- sd(data.PD$SM)

print(c("Predrought mean SM =",data.PD.mean,"±",data.PD.sd))

data.D <- data %>%
  filter(Condition == "Drought") 

data.D.mean <- mean(data.D$SM)
data.D.sd <- sd(data.D$SM)

print(c("drought mean SM =",data.D.mean,"±",data.D.sd))

#soil temperature statistics
data.PD <- data %>%
  filter(Condition == "PreDrought") 

data.PD.mean <- mean(data.PD$ST)
data.PD.sd <- sd(data.PD$ST)

print(c("Predrought mean ST =",data.PD.mean,"±",data.PD.sd))

data.D <- data %>%
  filter(Condition == "Drought") 

data.D.mean <- mean(data.D$ST)
data.D.sd <- sd(data.D$ST)

print(c("drought mean ST =",data.D.mean,"±",data.D.sd))

```

##statistical test

```{r}
wilcox.test(SM ~  Condition, data)

wilcox.test(ST ~ Condition, data)

wilcox.test(pH ~ Condition, data.pH)


```

```{r}
#create table for publication
data.pub <- data %>%
  dplyr::select(-c(SM.1, SM.2, SM.3, Date)) %>%
  pivot_wider(id_cols = Site, names_from = Condition)
  

```
#get average soil moisture, temp, and pH for sites 1 and 2 to add to plsr model
```{r}
#add soil moisture means
data.sm <- data %>%
  replace(is.na(.), 0) %>%
  group_by(Site, Condition) %>%
  dplyr::summarize(SM = mean(SM), ST = mean(ST)) %>%
   mutate(Condition2 = case_when(
    grepl("PreDrought", Condition) ~ "Pre-Drought",
    grepl("Drought", Condition) ~ "Drought")) %>%
  dplyr::select(-Condition) %>%
  mutate(Condition = Condition2) %>%
  dplyr::select(-Condition2)

data.sm

data.pH.summary <- data.pH %>%
  group_by(Site, Condition) %>%
  dplyr::summarize(pH = mean(pH)) %>%
   mutate(Condition2 = case_when(
    grepl("PreDrought", Condition) ~ "Pre-Drought",
    grepl("Drought", Condition) ~ "Drought")) %>%
  dplyr::select(-Condition) %>%
  mutate(Condition = Condition2) %>%
  dplyr::select(-Condition2)

data.abiotic <- data.sm %>%
  merge(.,data.pH.summary, by = c("Site", "Condition"))
```

