---
title: "metat-phyloseq"
author: "Linnea Honeker"
date: "3/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(ggplot2)
library(ape)
library(magrittr)
library(dplyr)
library(vegan)
library (DESeq2)
library(devtools)
library(RColorBrewer)
library(microbiome)
library(knitr)
library(plyr)
library(cowplot)
library(genefilter)

#set colors

colors_very_short <- c("black", "red", "blue", "green", "orange")

colors = c("black","gray","coral4","coral",
"chartreuse3", "darkseagreen1","blue","lightblue",
"yellow4", "yellow","darkorchid", "plum2",
"darkred", "darksalmon","green4",
"greenyellow","orange",
"moccasin",   "hotpink4", "lightpink", "lightblue4", "lightcyan3",  "lightslateblue", "lightsteelblue1", "navy", "blue4",
"darkgray", "black","brown", "cornflowerblue","darkgoldenrod",
"brown3","white","aliceblue","aquamarine","azure",
"beige","bisque","blue2","blueviolet","cyan","darkblue",
"chocolate","aquamarine3","darkcyan","deeppink",
"darkred","darkslateblue", "red", "green", "blue", "purple", "yellow")

```

## MetaT phyloseq


```{r import_data}
#import biom table
otu=read.table('../../Data/Taxonomy/MetaT/feature-table.csv',sep = ",", header = TRUE)

rownames(otu) <- otu[,1]
otu$Tax_ID <- NULL

otu <- as.matrix(otu) 
is.matrix(otu)

otu <- otu_table(otu, taxa_are_rows = TRUE)

#import sample metadata
map = read.table('../../Data/Taxonomy/MetaT/metadata.csv',sep = ",", header = TRUE)
row.names(map) <- map[,1]
map$SampleID <- NULL
map <- as.data.frame(map)
map <- sample_data(map)

#import taxnomy
taxa.frame = read.csv(file="../../Data/Taxonomy/MetaT/metaT-taxonomy.csv", header = TRUE, sep=",", 
                      blank.lines.skip = TRUE)
taxmat = as.matrix(taxa.frame)
rownames(taxmat) <- taxmat[,1]
taxmat <- taxmat[,-1]
tax=tax_table(taxmat)


#merge otu-table, sample-metadata, taxonomy
SP= merge_phyloseq(otu, tax,map)
SP

#filter to taxa > 0.001% and in > or = to 2 samples
total_counts <-sum(taxa_sums(SP))
threshold <- 0.000001*total_counts
SP.f <- filter_taxa(SP, function(x) sum(x) > threshold, TRUE)

#SP.f <- filter_taxa(SP, flist, TRUE)

#filter to Bacteria and Archaea Kingdoms
SPb = subset_taxa(SP.f, Kingdom == "Bacteria" |
                    Kingdom == "Archaea")
SPb

#filter to leave fungi 
SPf = subset_taxa(SP.f, Phylum == "Chytridiomycota" |
                    Phylum == "Blastocladiomycota" |
                    Phylum == "Neocallimastigomycota" |
                    Phylum == "Microsporidia" |
                    #Phlyum == "Glomeromycota") |
                    Phylum == "Ascomycota" |
                    Phylum == "Basidiomycota")
SPf

#filter to leave eukaryota
SPe = subset_taxa(SP.f, Kingdom == "Eukaryota")

SPe

#taxonomy--------------------------------------------------------
SP.rel.abun <- transform_sample_counts(SP.f, function(x) x/sum(x)) #convert counts to relative abundances
k.glom <- tax_glom(SP.rel.abun, taxrank = "Kingdom") #agglomerate taxa to Phylum-level
k.dat <- psmelt(k.glom)
k.dat$Kingdom <- as.character(k.dat$Kingdom) #convert to character vector from factor

#specify fungal relative abundance from all kingdoms
SP.rel.abun <- transform_sample_counts(SP.f, function(x) x/sum(x)) #convert counts to relative abundances
p.glom <- tax_glom(SP.rel.abun, taxrank = "Phylum") #agglomerate taxa to Phylum-level
p.dat <- psmelt(p.glom)
p.dat$Phylum <- as.character(p.dat$Phylum) #convert to character vector from factor

#add sub_kingdom to all phya table
p.dat2 <- p.dat %>%
  dplyr::mutate(sub_kingdom = ifelse(Phylum == "Chytridiomycota" |
                    Phylum == "Blastocladiomycota" |
                    Phylum == "Neocallimastigomycota" |
                    Phylum == "Microsporidia" |
                    Phylum == "Ascomycota" |
                    Phylum == "Basidiomycota",
                    "fungi", "other"))
#calculate means and standard deviations for fungal abundances within eukaryota
p.dat2.p <- p.dat2 %>% 
  dplyr::filter(Condition == "PreDrought" &
           sub_kingdom == "fungi") 

p.dat2.p.m <- mean(p.dat2.p$Abundance)
p.dat2.p.se <- sd(p.dat2.p$Abundance)

p.dat2.d <- p.dat2 %>% 
  dplyr::filter(Condition == "Drought" &
           sub_kingdom == "fungi") 

p.dat2.d.m <- mean(p.dat2.d$Abundance)
p.dat2.d.se <- sd(p.dat2.d$Abundance)
p.dat2.p.m
p.dat2.p.se
p.dat2.d.m
p.dat2.d.se


#prune out phyla below 2% in each sample for all experiments (bacteria/archaea)
SPb.rel.abun <- transform_sample_counts(SPb, function(x) x/sum(x)) #convert counts to relative abundances
P.glomb <- tax_glom(SPb.rel.abun, taxrank = "Phylum") #agglomerate taxa to Phylum-level
P.datb <- psmelt(P.glom)
P.datb$Phylum <- as.character(P.datb$Phylum) #convert to character vector from factor
P.datb$Phylum[P.datb$Abundance < 0.02] <- "Phylum < 2%"

SPb.rel.abun <- transform_sample_counts(SPb, function(x) x/sum(x)) #convert counts to relative abundances
C.glom <- tax_glom(SPb.rel.abun, taxrank = "Class") #agglomerate taxa to Phylum-level
C.dat <- psmelt(C.glom)
C.dat$Class <- as.character(C.dat$Class) #convert to character vector from factor
C.dat$Class[C.dat$Abundance < 0.02] <- "Class < 2%"

SPb.rel.abun <- transform_sample_counts(SPb, function(x) x/sum(x)) #convert counts to relative abundances
F.glom <- tax_glom(SPb.rel.abun, taxrank = "Family") #agglomerate taxa to Phylum-level
F.dat <- psmelt(F.glom)
F.dat$Family <- as.character(F.dat$Family) #convert to character vector from factor
F.dat$Family[F.dat$Abundance < 0.035] <- "Family < 3.5%"

#prune out phyla below 2% in each sample for all experiments for fungi
SPf.rel.abun <- transform_sample_counts(SPf, function(x) x/sum(x)) #convert counts to relative abundances
Pf.glom <- tax_glom(SPf.rel.abun, taxrank = "Phylum") #agglomerate taxa to Phylum-level
Pf.dat <- psmelt(Pf.glom)
Pf.dat$Phylum <- as.character(Pf.dat$Phylum) #convert to character vector from factor
Pf.dat$Phylum[Pf.dat$Abundance < 0.02] <- "Phylum < 2%"

SPf.rel.abun <- transform_sample_counts(SPf, function(x) x/sum(x)) #convert counts to relative abundances
Cf.glom <- tax_glom(SPf.rel.abun, taxrank = "Class") #agglomerate taxa to Phylum-level
Cf.dat <- psmelt(Cf.glom)
Cf.dat$Class <- as.character(Cf.dat$Class) #convert to character vector from factor
Cf.dat$Class[Cf.dat$Abundance < 0.02] <- "Class < 2%"

SPf.rel.abun <- transform_sample_counts(SPf, function(x) x/sum(x)) #convert counts to relative abundances
Ff.glom <- tax_glom(SPf.rel.abun, taxrank = "Family") #agglomerate taxa to Phylum-level
Ff.dat <- psmelt(Ff.glom)
Ff.dat$Family <- as.character(Ff.dat$Family) #convert to character vector from factor
Ff.dat$Family[Ff.dat$Abundance < 0.035] <- "Family < 3.5%"

SPf.rel.abun <- transform_sample_counts(SPf, function(x) x/sum(x)) #convert counts to relative abundances
Gf.glom <- tax_glom(SPf.rel.abun, taxrank = "Genus") #agglomerate taxa to Phylum-level
Gf.dat <- psmelt(Gf.glom)
Gf.dat$Genus <- as.character(Gf.dat$Genus) #convert to character vector from factor
Gf.dat$Genus[Gf.dat$Abundance < 0.05] <- "Genus < 3.5%"

#prune out phyla below 2% in each sample for all experiments for eukaryota
SPe.rel.abun <- transform_sample_counts(SPe, function(x) x/sum(x)) #convert counts to relative abundances
Pe.glom <- tax_glom(SPe.rel.abun, taxrank = "Phylum") #agglomerate taxa to Phylum-level
Pe.dat <- psmelt(Pe.glom)
Pe.dat$Phylum <- as.character(Pe.dat$Phylum) #convert to character vector from factor
Pe.dat$Phylum[Pe.dat$Abundance < 0.02] <- "Phylum < 2%"

#add sub_kingdom to eukaryota table
Pe.dat2 <- Pe.dat %>%
  mutate(sub_kingdom = ifelse(Phylum == "Chytridiomycota" |
                    Phylum == "Blastocladiomycota" |
                    Phylum == "Neocallimastigomycota" |
                    Phylum == "Microsporidia" |
                    Phylum == "Ascomycota" |
                    Phylum == "Basidiomycota",
                    "fungi", "other"))

#calculate means and standard deviations for fungal abundances within eukaryota
Pe.dat2.p <- Pe.dat2 %>% 
  dplyr::filter(Condition == "PreDrought" &
           sub_kingdom == "fungi") 

Pe.dat2.p.m <- mean(Pe.dat2.p$Abundance)
Pe.dat2.p.se <- sd(Pe.dat2.p$Abundance)

Pe.dat2.d <- Pe.dat2 %>% 
  dplyr::filter(Condition == "Drought" &
           sub_kingdom == "fungi") 

Pe.dat2.d.m <- mean(Pe.dat2.d$Abundance)
Pe.dat2.d.se <- sd(Pe.dat2.d$Abundance)



```

# Taxonomic profiles

```{r echo=FALSE, comment=""}
##########plot kingdom######################################
k.dat$Condition <- factor(k.dat$Condition, c("PreDrought", "Drought"))
k.dat$Time <- factor(k.dat$Time, c("0hr", "6hr", "48hr"))


p <- ggplot(k.dat, aes(x=Time, y=Abundance, fill=Kingdom)) 

K_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors_very_short) +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x=element_text(hjust =0.5, vjust=0.2),
        legend.title = element_blank()
  )

filename <- paste0("../../Figures/FigS3-Taxonomy/FigS3-metaT-Kingdom.png")
ggsave(filename,width=4,height=5,dpi=1000,K_barplot)

#######plot Phylum bacteria######################################
P.datb$Condition <- factor(P.datb$Condition, c("PreDrought", "Drought"))
P.datb$Time <- factor(P.datb$Time, c("0hr", "6hr", "48hr"))


p <- ggplot(P.datb, aes(x=Time, y=Abundance, fill=Phylum)) 

p_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors) +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x=element_text(hjust =0.5, vjust=0.2),
        legend.title = element_blank()
  )

filename <- paste0("../../Figures/FigS3-Taxonomy/FigS3-metaT-Phylum.png")
ggsave(filename,width=5,height=6,dpi=1000,p_barplot)

#######plot Classes bacteria######################################
C.dat$Condition <- factor(C.dat$Condition, c("PreDrought", "Drought"))
C.dat$Time <- factor(C.dat$Time, c("0hr", "6hr", "48hr"))


p <- ggplot(C.dat, aes(x=Time, y=Abundance, fill=Class)) 

c_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors) +
  theme(text = element_text(size = 12),
        axis.text.x=element_text(size=12, hjust =0.5, vjust=0.2),
        legend.text = element_text(size = 10))

filename <- paste0("../../Figures/FigS3-Taxonomy/metaT-Class.png")
ggsave(filename,width=10,height=5,dpi=1000,c_barplot)

#######plot family bacteria######################################
F.dat$Condition <- factor(F.dat$Condition, c("PreDrought", "Drought"))


p <- ggplot(F.dat, aes(x=Time, y=Abundance, fill=Family)) 

f_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors) +
  theme(text = element_text(size = 12),
        axis.text.x=element_text(size=12,hjust =0.5, vjust=0.2),
        legend.text = element_text(size = 10),
  )

#######plot Phylum fungi######################################
Pf.dat$Condition <- factor(Pf.dat$Condition, c("PreDrought", "Drought"))
Pf.dat$Time <- factor(Pf.dat$Time, c("0hr", "6hr", "48hr"))


p <- ggplot(Pf.dat, aes(x=Time, y=Abundance, fill=Phylum)) 

pf_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors_very_short) +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x=element_text(hjust =0.5, vjust=0.2),
        legend.title = element_blank()
  )

filename <- paste0("../../Figures/FigS3-Taxonomy/FigS3-metaT-Phylum-fingi.png")
ggsave(filename,width=5,height=5,dpi=1000,pf_barplot)

#######plot Classes eukaryota######################################
Cf.dat$Condition <- factor(Cf.dat$Condition, c("PreDrought", "Drought"))
Cf.dat$Time <- factor(Cf.dat$Time, c("0hr", "6hr", "48hr"))


p <- ggplot(Cf.dat, aes(x=Time, y=Abundance, fill=Class)) 

cf_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors) +
  theme(text = element_text(size = 12),
        axis.text.x=element_text(size=12, hjust =0.5, vjust=0.2),
        legend.text = element_text(size = 10))

filename <- paste0("../../Figures/FigS3-Taxonomy/metaT-Class-fungi.png")
ggsave(filename,width=10,height=5,dpi=1000,cf_barplot)

#######plot family fungi######################################
Ff.dat$Condition <- factor(Ff.dat$Condition, c("PreDrought", "Drought"))


p <- ggplot(Ff.dat, aes(x=Time, y=Abundance, fill=Family)) 

ff_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors) +
  theme(text = element_text(size = 12),
        axis.text.x=element_text(size=12,hjust =0.5, vjust=0.2),
        legend.text = element_text(size = 10),
  )

#######plot genus fungi######################################
Gf.dat$Condition <- factor(Gf.dat$Condition, c("PreDrought", "Drought"))


p <- ggplot(Gf.dat, aes(x=Time, y=Abundance, fill=Genus)) 

Gf_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors) +
  theme(text = element_text(size = 12),
        axis.text.x=element_text(size=12,hjust =0.5, vjust=0.2),
        legend.text = element_text(size = 10),
  )

filename <- paste0("../../Figures/FigS3-Taxonomy/metaT-genus-fungi.png")
ggsave(filename,width=12,height=5,dpi=1000,Gf_barplot)

#######plot phyla eukaryota######################################
Pe.dat$Condition <- factor(Pe.dat$Condition, c("PreDrought", "Drought"))
Pe.dat$Time <- factor(Pe.dat$Time, c("0hr", "6hr", "48hr"))


p <- ggplot(Pe.dat, aes(x=Time, y=Abundance, fill=Phylum)) 

pf_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors_long) +
  theme_bw() +
  theme(text = element_text(size = 11),
        axis.text.x=element_text(hjust =0.5, vjust=0.2),
        legend.title = element_blank()
  )

filename <- paste0("../../Figures/FigS3-Taxonomy/FigS3-metaT-Phylum-eurkaryota.png")
ggsave(filename,width=5,height=5,dpi=1000,pf_barplot)

#######plot fungi vs other in eukaryota######################################
Pe.dat2$Condition <- factor(Pe.dat2$Condition, c("PreDrought", "Drought"))
Pe.dat2$Time <- factor(Pe.dat2$Time, c("0hr", "6hr", "48hr"))


p <- ggplot(Pe.dat2, aes(x=Condition, y=Abundance, fill=sub_kingdom)) 

pf_barplot <- p + geom_bar(aes(), stat = "identity", position = "fill") +
  #facet_wrap(~ Condition, scales="free") +
  scale_fill_manual(values=colors_short) +
  theme_bw() +
  theme(text = element_text(size = 6),
        axis.text.x=element_text(hjust =0.5, vjust=0.2),
        legend.title = element_blank()
  )

filename <- paste0("../../Figures/FigS3-Taxonomy/FigS3-metaT-Phylum-eurkaryota-fungi-other.png")
ggsave(filename,width=2,height=2,dpi=1000,pf_barplot)
```

## Kingdom     

```{r, echo = FALSE}
K_barplot
```

## Phylum 

```{r, echo = FALSE}
p_barplot
```

## Class    

```{r, echo = FALSE}
c_barplot
```

## Family    

```{r fig.height=8, fig.width=10, echo = FALSE}
f_barplot
```

## Beta diversity
PCoA: bray-curtis 
```{r "beta-diversity", fig.height=8, fig.width=10, echo=FALSE, comment=""}
SPb.ra = rarefy_even_depth(SPb,(min(sample_sums(SPb))))

sample_data(SPb.ra)$Condition <- factor(sample_data(SPb.ra)$Condition, levels = c("PreDrought","Drought"))
sample_data(SPb.ra)$Time <- factor(sample_data(SPb.ra)$Time, levels = c("0hr","6hr", "48hr"))
sample_data(SPb.ra)$Condition_Time <- factor(sample_data(SPb.ra)$Condition_Time, 
                                             levels = c("PreDrought_0hr","PreDrought_6hr", "PreDrought_48hr",
                                                        "Drought_0hr", "Drought_6hr", "Drought_48hr"))

my.ord <- ordinate(
  physeq = SPb.ra,
  method = "PCoA",
  distance = "bray")
  #weighted=TRUE)

#condition and time
ord.plot.time <- (
  plot_ordination(
    physeq = SPb.ra,
    ordination = my.ord,
    color = "Condition",
    #type = "split",
    shape = "Time",
  ) +
    geom_point(aes(color = Condition), size = 4) )#+
    #geom_point(colour = "grey90", size = 1) +
    #scale_color_manual(breaks = c("Clitoria", "Piper", "Hibiscus"),
      #                 values = c("green4", "red", "blue"))
#)

#condition and site
ord.plot.site <- (
  plot_ordination(
    physeq = SPb.ra,
    ordination = my.ord,
    color = "Condition",
    #type = "split",
    shape = "Site",
  ) +
    geom_point(aes(color = Condition), size = 4) )

plot_grid (ord.plot.time, ord.plot.site,
           labels = c("A", "B"),
           label_size = 12,
           ncol = 2)
```


```{r "alpha-diversity", echo=FALSE, comment=""}
#define and plot alpha diversity measurements on unrarefied samples
my_colors_st = c("#B6D7A8", "#6AA84F", "#274E13", "#EA9999", "#CC0000", "#660000")
alpha_meas = c("Observed", "Shannon")
p <- plot_richness(SPb.ra, "Condition_Time", scales = "free_y", measures="Observed")

#Add ggplot2 layer-boxplots
p1 <- p + scale_color_manual(values = my_colors_st) +
  geom_boxplot(data=p$data, aes(x=Condition_Time, y=value, fill=Condition_Time), 
                      alpha=20, size = 0.5) +
  scale_fill_manual(values = my_colors_st) +
  scale_x_discrete(name = element_blank(), labels = c("0hr", "6hr", "48hr", "0hr", "6hr", "48hr")) +
  theme_bw() +
  theme(text = element_text(size = 11),
        legend.position = "none",
        axis.text.x=element_text(size=11,hjust =0.5, vjust=0.8),)

#condition and site
#define and plot alpha diversity measurements on unrarefied samples
p <- plot_richness(SPb.ra, "Site", measures="Observed")

#Add ggplot2 layer-boxplots
p2 <- p + scale_color_manual(values = c("orange", "magenta", "turquoise")) +
  geom_boxplot(data=p$data, aes(x=Site, y=value, fill=Site), 
               alpha=20, size = 0.5) +
  scale_fill_manual(values = c("orange", "magenta", "turquoise")) +
  scale_x_discrete(name = element_blank()) +
  theme_bw() +
  theme(text = element_text(size = 11),
        legend.position = "none",
        axis.text.x=element_text(size=11,hjust =0.5, vjust=0.8))

plot_grid (p1, p2,
           labels = c("A", "B"),
           label_size = 12,
           ncol = 2)

```


## statistical analysis of taxonomy - bacteria/archaea
```{r}
####phlya across all kingdoms that are in tableS4 (removed some phyla that were in only 1 or 2 samples)
phyla = c("Proteobacteria", "Acidobacteria", "Actinobacteria", "Bacteroidetes",
          "Chloroflexi","Cyanobacteria",  "Firmicutes", 
          "Gemmatimonadetes",  "Planctomycetes",
          "Thaumarchaeota", "Verrucomicrobia")
for (p in phyla)
{
p.P.datb <- P.datb %>%
  dplyr::filter(Phylum == p)

lme.p.P.dat <- lme(Abundance ~ Condition,
                 random = list(Site = ~1),
                 data = p.P.datb,
                 weights = varIdent(form = ~1|Condition)
)

sum <- summary(lme.p.P.dat)
tabl = sum$tTable
cat("\nresults for ",p,":\n")
print(tabl)
}

# Condition - Drought
phyla = c("Proteobacteria", "Acidobacteria", "Actinobacteria", "Bacteroidetes",
          "Chloroflexi","Cyanobacteria",  "Firmicutes",
          "Planctomycetes",
          "Thaumarchaeota", "Verrucomicrobia")
for (p in phyla)
{
p.P.datb <- P.datb %>%
  dplyr::filter(Phylum == p) %>%
  dplyr::filter(Condition == "Drought")

lme.p.P.dat <- lme(Abundance ~ Time,
                 random = list(Site = ~1),
                 data = p.P.datb,
                 weights = varIdent(form = ~1|Time)
)
sum <- summary(lme.p.P.dat)
tabl = sum$tTable
cat("\nresults for ",p,":\n")
print(tabl)
}

# Condition - Pre-Drought
phyla = c("Proteobacteria", "Acidobacteria", "Actinobacteria", "Bacteroidetes",
          "Firmicutes", 
          "Gemmatimonadetes",  "Planctomycetes",
          "Verrucomicrobia")
for (p in phyla)
{
p.P.datb <- P.datb %>%
  dplyr::filter(Phylum == p) %>%
  dplyr::filter(Condition == "PreDrought")

lme.p.P.dat <- lme(Abundance ~ Time,
                 random = list(Site = ~1),
                 data = p.P.datb,
                 weights = varIdent(form = ~1|Time)
)
sum <- summary(lme.p.P.dat)
tabl = sum$tTable
cat("\nresults for ",p,":\n")
print(tabl)
}

```
## statistical analysis of taxonomy - fungi
```{r}
####phlya across all kingdoms that are in tableS4 (removed some phyla that were in only 1 or 2 samples)
phyla = c("Chytridiomycota", "Blastocladiomycota" ,"Ascomycota",     
          "Basidiomycota")
for (p in phyla)
{
p.Pf.dat <- Pf.dat %>%
  dplyr::filter(Phylum == p)

lme.p.Pf.dat <- lme(Abundance ~ Condition,
                 random = list(Site = ~1),
                 data = p.Pf.dat,
                 weights = varIdent(form = ~1|Condition)
)

sum <- summary(lme.p.Pf.dat)
tabl = sum$tTable
cat("\nresults for ",p,":\n")
print(tabl)
}

# Condition - Drought
phyla = c("Chytridiomycota", "Blastocladiomycota" ,"Ascomycota",     
          "Basidiomycota")
for (p in phyla)
{
p.Pf.dat <- Pf.dat %>%
  dplyr::filter(Phylum == p) %>%
  dplyr::filter(Condition == "Drought")

lme.p.Pf.dat <- lme(Abundance ~ Time,
                 random = list(Site = ~1),
                 data = p.Pf.dat,
                 weights = varIdent(form = ~1|Time)
)
sum <- summary(lme.p.Pf.dat)
tabl = sum$tTable
cat("\nresults for ",p,":\n")
print(tabl)
}

# Condition - Pre-Drought
phyla = c("Chytridiomycota", "Blastocladiomycota" ,"Ascomycota",     
          "Basidiomycota")
for (p in phyla)
{
p.Pf.dat <- Pf.dat %>%
  dplyr::filter(Phylum == p) %>%
  dplyr::filter(Condition == "PreDrought")

lme.p.Pf.dat <- lme(Abundance ~ Time,
                 random = list(Site = ~1),
                 data = p.Pf.datb,
                 weights = varIdent(form = ~1|Time)
)
sum <- summary(lme.p.Pf.dat)
tabl = sum$tTable
cat("\nresults for ",p,":\n")
print(tabl)
}

```

## statistical analysis of taxonomy - total fungi, archaea, bacteria
```{r}
####phlya across all kingdoms that are in tableS4 (removed some phyla that were in only 1 or 2 samples)
##fungi
f.p.dat2 <- p.dat2 %>%
  dplyr::filter(sub_kingdom == "fungi")
  
lme.f.p.dat2 <- lme(Abundance ~ Condition,
                 random = list(Site = ~1),
                 data = f.p.dat2,
                 weights = varIdent(form = ~1|Condition)
)


sum <- summary(lme.f.p.dat2)
tabl = sum$tTable
tabl

# archaea
a.p.dat2 <- p.dat2 %>%
  dplyr::filter(Kingdom == "Archaea")
  
lme.a.p.dat2 <- lme(Abundance ~ Condition,
                 random = list(Site = ~1),
                 data = a.p.dat2,
                 weights = varIdent(form = ~1|Condition)
)


sum <- summary(lme.a.p.dat2)
tabl = sum$tTable
tabl

